### 2019-11-05 paper/paper-v3.tex
started improving according to reviewers' response

\documentclass[12pt]{article}

\usepackage[margin=3cm]{geometry}

\renewcommand{\baselinestretch}{1.5}

\usepackage[hidelinks]{hyperref}

\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{shapes,positioning,arrows,backgrounds,fit,calc,automata}   
\usetikzlibrary{decorations.markings} 
\usepackage{xcolor}
\usepackage{mathtools} 
\usepackage{booktabs}  
\usepackage{xspace}
\usepackage{color}
\usepackage{amsfonts}	
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage[shortlabels]{enumitem}
\usepackage{etoolbox}  
\usepackage{cleveref}   
\usepackage{wrapfig}
\usepackage{caption} 
\usepackage{adjustbox}

\usepackage[framemethod=TikZ]{mdframed}
\usepackage{amsthm}
\usepackage{thmtools}
\usepackage{amsthm,thmtools,xcolor}

\usepackage{multirow}

\usepackage{palatino}
\usepackage[normalem]{ulem}
\usepackage{color}
\usepackage[utf8x]{inputenc}
\usepackage{default}
\usepackage{xcolor}
\usepackage{colortbl}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{array,amsmath}
\usepackage{booktabs}
\usepackage{ulem}
\usepackage{tikz}
\usepackage{graphicx,wrapfig,lipsum}
\usepackage{rotating}
\usepackage{xcolor}
\usepackage{scalerel}

\input{abbreviations}

% \newcommand{\commentem}[1]{
% \medskip \noindent
% {\sc COMMENT (EMMA):
% #1}
% }

\parskip 0.2in
\linespread{1.25}

\thispagestyle{empty}

\title{Cognitive Principles and Individual Differences \\ in Human Syllogistic Reasoning}
\author{
Emmanuelle-Anna Dietz Saldanha and Richard M{\"o}rbitz\thanks{The authors are mentioned in alphabetical order.} \\
\normalsize International Center for Computational Logic, TU
  Dresden, Germany
}  
% \institute{}
\date{}

\begin{document}
\maketitle

\section{Introduction}
% 
% 
% Technology occupies a large part of every day life in our society.
% Therefore, an effortless interaction between humans and technical devices
% is of crucial interest for everyone. As it is not likely that humans will make the effort to understand their devices but instead they will place high demands on them, their devices should be able to adapt to the human needs in the best possible way. For instance, a system should be able to recognize small changes and additionally identify which environmental parameters are relevant for the current communication with its owner. This in turn should allow the device to adapt flexibly and independently to the environment it is currently situated in. In order to develop such a system, it is necessary to understand how humans think and to recognize their behavioral patterns.

Originally, one of the objectives of using (not necessarily classical) logic in the field of AI and logic programming was to formalize human reasoning and commonsense reasoning~\cite{McCarthy59,McCarthy98}. 
However, the proposed approaches did not have the expected successes in this~(cf.~\cite{mcdermott:87}),
as proposed approaches were mostly theoretical and did not apply to real case studies.
On the other hand, cognitive scientists have extensively investigated human thinking and acquired a great expertise.
However, this expertise was not consulted, i.e. there was no communication among the relevant fields.

Cognitive theories that try to explain human reasoning need to be evaluated by their cognitive adequacy~\cite{strube:1992} and their performance should be compared to other theories.
Conceptual cognitive adequacy describes to what extent a system corresponds to human conceptual knowledge, whereas inferential cognitive adequacy indicates whether the reasoning process of a system is structured similarly to the way humans reason~\cite{knauff:rauh:renz:1997}.

Yet, the cognitive adequacy of a theory can only be measured, if that theory is evaluable. In this context, that means, if the theory can uniformally account for episodes of human reasoning systematically, according to its underlying assumptions. Only then, the cognitve adequacy of that theory can be compared to others.

Formal methods can offer the required tools to implement such cognitive theories. This in turn, can allow a new way of communication among the fields, by avoiding ambiguous understanding, showing their formal correspondences and finally providing an appropriate evaluation method with respect to their cognitive adequacy. In order to determine the required factors that are necessary for the evaluation of an adequate cognitive theory, we take~\cite{ragni:stolzenburg:2015} as a starting point:
We aim for a theory in which (i) all processes are comprehensible, (ii) the suggested principles represent general cognitive processes, and (iii) the predictions of our theory are evaluated with respect to a wide range of human reasoning tasks. 
% Even though the main logical formalism of Classical Logic (CL) has successfully been applied to our scientific reasoning,
% various experimental studies have shown that human reasoning systematically deviates 
% from classical logical conclusions (cf.~\cite{wason:68,byrne:89}).

The results of the meta-study on syllogistic reasoning in~\cite{khemlani:2012} show that the answers given by humans does not only systematically differ from the valid conclusions under classical logic, but from any other proposed cognitive theory (such as the theories proposed in~\cite{BJ89,johnsonlaird:1983,Rips1994,Polk1995,chater:oaksford:1999}).
% The approach suggested in~\cite{stolzenburg:luederitz:2017}, is based on venn diagrams, where moods are understood as conditional probabilities, and implemented 
% in constraint logic programming. Yet, another logic programming approach, the Weak Completion Semantics, is suggested in~\cite{cogsci:2017}.
Syllogisms originate from Aristotle~\cite{aristotle} and
they are interesting because on one hand they are easy to understand and on the other hand
they are complex enough to require actual reasoning. 
A syllogism consists of two
premises and one conclusion. 
The assertions are statements about properties over terms and can have one of the classical quantifiers in Table~\ref{table:moods}, which are called moods. The pair of syllogistic premises can have four different orders as shown in Table~\ref{t:2}. Based on these moods and orders, 64 different pairs of syllogistic premises can be constructed.
A conclusion then consists of an assertion about two properties over terms where each of these properties occurs in only one of the premises, together with a third middle property that does not appear in the conclusion but appears in both premises. 
\begin{table}[t]
    \begin{minipage}{0.6\textwidth}
        \centering
	\[	
\begin{array}{@{\hspace{0.0cm}}l@{\hspace{0.1cm}}G@{\hspace{0.1cm}}l@{\hspace{0.3cm}}c}\toprule
	\mbox{mood} & \mbox{natural language} & \mbox{FOL} &
\mbox{Short} \\\midrule
	\mbox{affirmative universal} & \mbox{\em all a are b} & \forall
	X (a(X) \rightarrow b(X)) & {\MA ab}\\
	\mbox{affirmative existential} & \mbox{\em some a are b} &
	\exists X (a(X) \wedge b(X)) & {\MI ab}\\
	\mbox{negative universal} & \mbox{\em no a are b} &
	\forall X (a(X) \rightarrow \neg b(X)) & {\ME ab}\\
	\mbox{negative existential} & \mbox{\em some a are not b} &
	\exists X (a(X) \wedge \neg b(X)) & {\MO ab} \\
	\bottomrule
	\end{array}
	\]
	\caption{The four moods and their
formalization\label{table:moods}.}
	\end{minipage}
    \begin{minipage}{0.45\textwidth}
        \centering
%  	\small
	\begin{center}
	\[
		\begin{array}{@{\hspace{0cm}}r@{\hspace{0.1cm}}c@{\hspace{0.1cm}}c@{\hspace{0.1cm}}c@{\hspace{0.1cm}}c}
		\toprule
		& \mbox{premise 1} & \mbox{premise 2} \\
		\midrule
		\mbox{figure 1} &  a-b & b-c \\
		\mbox{figure 2} & b-a &  c-b \\
		\mbox{figure 3} & a-b & c-b \\
		\mbox{figure 4} & b-a & b-c \\
		\bottomrule
		\end{array}
		\]
	\end{center}
	\caption{The four figures.\label{t:2}}
	\end{minipage}
\end{table}
Consider the following pair of syllogistic premises: 
\begin{align}
 \emph{Some bakers are artists. }\quad\quad  \quad\quad & \emph{No chemists are bakers.} \tag{IE2} 
\end{align}
Given these
two premises, which conclusion on the relation between \emph{artists} and \emph{chemists} necessarily follows?
According to~\cite{khemlani:2012}, the majority of participants in experimental studies
concluded \emph{No chemists are artists} (29\%) and \emph{No valid conclusion} (27\%). The classical logical valid conclusion \emph{Some artists are not chemists}, was concluded 
by only 16\%, which is exactly the threshold, that the answer had not been chosen randomly.\footnote{
 Given that there are nine different possible conclusions, the chance that a conclusion has been chosen randomly is 1/9 = 11.1 \%. 
 A binomial test shows that if a conclusion is drawn in
more than 16\% of the cases by the participants it is unlikely that is has been
chosen by just random guesses~\cite{khemlani:2012}. }
% Recently, a new cognitive theory has been proposed for syllogistic reasoning~\cite{cogsci:2017}.
% Surprisingly this approach is competitive with the other cognitive theories discussed in~\cite{khemlani:2012}.
% \cite{cogsci:2017} deals with the first step of modeling human reasoning towards an adequate representation, for which it identifies seven cognitive principles, i.e.\ assumptions that humans make while reasoning, motivated from the literature and own observations. For each of these principles a formal representation
% in logic program clauses is provided. 
% For the second step,  reasoning towards the least models of the weak completion of these programs under the three-valued
% {\L}ukasiewicz Logic, is suggested.
% One reason is because often only {aggregated data} of the experiments are accessible.
% i.e.\ the data summed up with respect to the individual answer possibilities.
% The experimental set-up was made such that participants would only choose one possible conclusion, which implies that 
% these answers excluded each other. In particular, it can be assumed that participants were at least minimally consistent in the sense that
% i.e.\ 
% The majority's conclusions exclude each other, as the test was made such that 
The ones who answered \emph{No valid conclusion}
were not the same ones who answered \emph{No chemists are artists} or \emph{Some artists are not chemists}, and vice versa.\footnote{Note that usually in these experiments, participants were supposed to give only one answer}

A major drawback of the suggested cognitive theories so far is that only the participants'
aggregated data was considered: These theories did not account for differences
among individuals, i.e.\ they only considered the data summed up with respect to the individual answer possibilities. In contrast, non-aggregated data records the individual participants' response patterns, i.e.\ the complete set of responses that are given by each participant.

As the above example and recent investigations on psychological experiments show (cf.~\cite{khemlani:2016, ragni:2017}), it seems that different humans represent and reason about conditional information differently.
This implies that \textit{the} human reasoner does not exist, instead there 
seem to exist various (human) reasoning clusters. The identification of these differences and the specification of these clusters is central 
for the development of an adequate cognitive theory. Only recently approaches have been proposed that account for individual differences (or clusters) in syllogistic reasoning (cf.~\cite{khemlani:2016,declare:2017}).

%  The results show that some but not all of the proposed clusters are well reflected in the individual reasoning patterns.
% More interestingly are the following two observations:
% The dataset in~\cite{khemlani:2016} provides an additional interesting information: 
% Interestingly, when comparing the changes of results of each individual participant
% from the first to the second week, it seems that their answers in the second week fits better the classical logic derivations. 
% We argue that these reasoners can be characterized by a set of cognitive principles, motivated from studies made in Cognitive Science.
% First attempts to propose a model that distinguishes between human reasoners have been proposed in~\cite{khemlani:2016}. 
% The authors identify three different clusters of reasoners labeled as intuitive, intermediate and deliberative. They propose an approaches based on the MReasoner and suggest to   ...
% We evaluate the non-aggregated data provided by~\cite{khemlani:2016} and taking the clusters suggested in~\cite{declare:2017} as starting point. The non-aggregated data consists of two dataset for 20 participants, who carried out the same syllogistic reasoning task twice with a interval of one week.
%  The evaluation shows that some but not all of the proposed clusters in~\cite{declare:2017} are well reflected in the individual reasoning patterns.
% Interestingly, when comparing the changes of results of each individual participant
% from the first to the second week, it seems that their answers in the second week fits better the classical logic derivations. 
% We formulate the hypothesis that the individual clusters do not represent different human reasoners, but different stages of reasoning. Depending on the persistence and the
% previous training of the individual participant, ...

Here, we focus on a qualitative assessment of the clustering approach in~\cite{declare:2017},
in the sense that we investigate the individual participants' conclusions with respect to the non-aggregated dataset provided by~\cite{khemlani:2016}.
In Section~\ref{sect:wcsclustering} and~\ref{sect:representing}, we introduce the underlying computational logic approach and the cognitive principles. Section~\ref{sect:aggregateddata} illustrates an example and briefly discusses the performance of the approach in~\cite{declare:2017} on non-aggregated data.
Section~\ref{sect:nonag} provides a new characterization of the individual's reasoning patterns that 
seems to be more adequate.

\section{WCS Clustering} \label{sect:wcsclustering}

The authors in~\cite{declare:2017} put forward the hypothesis that reasoning clusters can be expressed through cognitive principles and heuristic strategies. Cognitive principles are (not necessarily valid) assumptions made by humans while reasoning, 
whereas heuristic strategies describe the methods that are applied when
conclusions are not derived by reasoning but instead based on structural patterns or {biases}.
The approach suggested in~\cite{declare:2017} is based on a new computational logic approach for human syllogistic reasoning~\cite{cogsci:2017}, which seems to perform better than other state-of-the-art cognitive theories when evaluated with aggregated data provided in~\cite{khemlani:2012}. 
The approach presented in~\cite{declare:2017} extended~\cite{cogsci:2017} by specifying three reasoning clusters and two heuristic strategies and achieved a match of 92\%, which was the highest result compared to the presented approaches in~\cite{khemlani:2012}.
The system developed in~\cite{declare:2017}, was extended and participated at the Syllogism Challenge 2017.\footnote{\url{https://www.cc.uni-freiburg.de/modelingchallenge/challenge-2017}} Figure~\ref{overview:system:syllchal} illustrates the components of the system:
First, groups of reasoners were specified based on cognitive principles and heuristic strategies.
After that, 
a translation into logic program clauses was provided, such that for each reasoning clusters, the corresponding logic program
was constructed. Next, the desired models according to the Weak Completion Semantics~\cite{hk:2009a,hk:2009b} were computed with a corresponding semantic operator~\cite{stenning:vanlambalgen:2008}.
After that, the entailed relation between $a$ and $c$ were extracted. If no relation between $a$ and $c$ could be entailed, then \textit{No valid conclusion}
was derived. The final answer distribution was specified according to the number of reasoning clusters 
that entailed the conclusion.
Finally, an illustrative representation of these clusters with respect to their distribution was done by means of of multinomial process trees~\cite{Riefer1988}.

\begin{figure}\small
\begin{tikzpicture} 
\draw (0,8.5) rectangle (13.2,9.5);
\node (A1) at (2,8.4) {};

\node[anchor=west]  at (0,9.2) {\textbf{Cognitive Principles} presuppositions that humans make while reasoning};
\node[anchor=west]  at (0,8.7) {\textbf{Heuristic Strategies} reasoning based on structure in patterns or biases};

\node (A2) at (2,7.9) {};
\draw[thick, ->] (A1) to[m] node[right] {\tiny \gcolor{Description}} (A2);

\draw (0,7.3) rectangle (13.2,7.8);
\node[anchor=west] at (0,7.5) {\textbf{Reasoning Clusters} groups of reasoners};
\node (A3) at (2,7.3) {};

\node (A4) at (2,6.8) {};
\draw[thick, ->] (A3) to[m] node[right] {\tiny \gcolor{Translation}} (A4);
% 
\draw (0,6.3) rectangle (13.2,6.8);
\node[anchor=west]  at (0,6.5) {\textbf{Program $\CalP$} cognitive principles represented
by logic program clauses};
\node (A5) at (2,6.3) {};

\node (A6) at (2,5.8) {};
\draw[thick, ->] (A5) to[m] node[right] {\tiny \gcolor{Computation}} (A6);

\draw (0,5.3) rectangle (13.2,5.8);
\node[anchor=west]  at (0,5.5) {\textbf{Model $\CalM$} least fixed point of the semantic $\Phi_\CalP$ operator};
\node (A7) at (2,5.3) {};

\node (A8) at (2,4.8) {};
\draw[thick, ->] (A7) to[m] node[right] {\tiny \gcolor{Entailment}} (A8);

\node[anchor=west]  at (0,4.5) {\textbf{Conclusions} all artists are chemists (\MA ac), some artists are chemists (\MI ac), .... };
\draw (0,4.3) rectangle (13.2,4.8);
\node (A9) at (2,4.3) {};

\node (A10) at (2,3.8) {};
\draw[thick, ->] (A9) to[m] node[right] {\tiny \gcolor{Specification}} (A10);

\draw (0,3.2) rectangle (13.2,3.8);
\node[anchor=west]  at (0,3.5) {\textbf{Distribution} based on number of clusters that entail the conclusion};

\draw (13.4,5.3) rectangle (15.6,7.3);
\node (A11) at (14.5,6.7) {{\small Multinomial}};
\node at (14.5,6.3) {\small Process};
\node at (14.5,5.8) {\small Trees};
\draw[thick, ->] (13.2,9) -| node {}  (14.5,7.3);
\draw[thick, ->] (14.5,5.3) |- node {}  (13.2,3.55);

\end{tikzpicture}
\caption{The system that participated at the Syllogism Challenge 2017\label{overview:system:syllchal}}

\end{figure}

As environment for the formalization, we suggest an extension of the
Weak Completion Semantics (WCS). WCS has first been introduced in~\cite{hk:2009a,hk:2009b} 
and has shown to model a wide range of human reasoning tasks~\cite{btg:2015:h}. 
WCS is the non-monotonic computational logic approach to consider
weakly completed logic programs, to compute their least models
under the three-valued {\L}ukasiewicz logic~\cite{lukasiewicz:20}, and to reason with respect to these models or search for explanations by means of (skeptical) abduction~\cite{kakas:etal:93}.

\subsection{Contextual Logic Programs} \label{sect:wcs}

We introduce an extension of WCS first presented in~\cite{dietz:hoelldobler:pereira:2017}. 
\textit{Contextual logic programs} are (data) logic programs extended by the truth-functional operator~$\cxt$, called \textit{context}. \textit{Contextual (logic) program clauses}
are expressions of the forms $A  \leftarrow  L_1 \land \ldots \land L_m \land
\cxt(L_{m+1}) \land \ldots \land \cxt(L_{m+p})$ (called
\textit{rules}), $A \leftarrow \top$ (called \textit{facts}), $A \leftarrow \bot$ (called
\textit{negative assumptions})\footnote{Under WCS, the negative
assumption will become $A \leftrightarrow \bot$ and, hence, $A$ has to be false.
This can, however, be overwritten by other rules and facts (\textit{defeating} the assumption).} and $A \leftarrow \udf$ (called
\textit{unknown assumptions}), where
$A$ is an atom and the~$L_i$ with~$1 \leq i \leq m+p$ is a literal.
$A$ is called \textit{head}
and $L_1 \land \ldots \land L_m \land
\cxt(L_{m+1}) \land \ldots \land \cxt(L_{m+p})$ as well as~$\top, \bot$
and $\udf$,
standing for \textit{true}, \textit{false} and \textit{unknown} respectively, are
called \textit{body} of the corresponding clauses.  A
\textit{contextual program}, denoted by $\mathcal P$,
is a finite set of contextual program clauses.
As we restrict terms to be either constants or variables only, we consider so-called \defname{data logic programs}.
If the clause contains variables, then they are implicitly universally quantified within the scope of the entire clause. 
For instance, the clause $p(X) \leftarrow q(X)$ represents the formula $(\forall X) (p(X) \leftarrow q(X))$. Therefore, a clause never contains free variables.
A \defname{ground clause} is a clause that does not contain variables.
$\Ground \CalP$ denotes the set of all its ground clauses.
$A$ is \textit{defined in $\Ground \CalP$}  iff $\Ground \CalP$ contains a rule, a fact 
or an assumption with head $A$.
The \textit{definition of $A$ in $\Ground \CalP$} is defined as $\Def(A,\CalP)
= \{ A \leftarrow \Body \mid A \leftarrow \Body \mbox{ is a rule, a fact or an assumption
occurring in } \Ground \CalP\}$.
$A$ is \textit{undefined in $\Ground \CalP$} iff $A$ is not defined in $\Ground \CalP$.
The set of all atoms that are undefined in~$\Ground \CalP$ is denoted by~$\ud(\CalP)$.
% $\neg A$ is \textit{assumed in $\Ground \CalP$} iff $\Ground \CalP$
% contains a negative assumption with head $A$, $\Ground \CalP$ does not
% contain an unknown assumption with head $A$, and $\Def(A,\CalP) = \emptyset$.
We omit the word \textit{contextual} when we refer to programs, if not stated otherwise.

\subsection{Three-Valued {\L}ukasiewicz Logic Extended with $\cxt$} \label{sub:3valuedL}

We consider the three-valued {\L}ukasiewicz logic~\cite{lukasiewicz:20}
extended with the $\cxt$ connective, for which the
corresponding truth values are $\top$, $\bot$ and $\udf$, meaning \textit{true}, \textit{false}
and \textit{unknown}, respectively.
% As it seems,  $\cxt$ corresponds to {\L}ukasiewicz necessity operator~\cite{}, where the difference is, that $\cxt$ is restricted to be applicable to literals only.
A \defname{three-valued interpretation}~$I$ is a mapping from the set of
logical formulas to the set
of truth values~$\{ \true, \bot, \udf\}$, represented as a pair~$I~=~\langle I^\true,I^\bot \rangle$ of two disjoint sets of atoms:  $I^\true =  \{A \mid A \mbox{ is mapped to } \true \mbox{ under } I\}$ and $I^\bot = \{A \mid A \mbox{ is mapped to } \bot \mbox{ under } I \}$.
Atoms which do not occur in~$I^\true \cup I^\bot$ are mapped to~$\udf$.
The truth value of a given formula
under $I$ is determined according to the truth tables in
Table~\ref{tab:3vldluka}.
$I(F) = \top$ denotes that a formula $F$ is mapped to true under $I$.
A \defname{three-valued model} $\CalM$ of~$\CalP$ is a three-valued interpretation such that
$\CalM(A \leftarrow \Body) = \true$ for each $A \leftarrow \Body \in \Ground\CalP$.
Let $I = \langle I^\top, I^\bot \rangle$ and $J = \langle J^\top,
J^\bot \rangle$ be two interpretations.
$I \subseteq J$ iff $I^\top \subseteq J^\top$ and $I^\bot \subseteq J^\bot$. $I$ is the \textit{least model} of $\CalP$
iff for any other model~$J$ of $\CalP$ it holds that $I \subseteq J$.

\begin{table}[t]
\[\small
\begin{array}[c]{@{\hspace{0mm}}c|c}
F & \neg F\\ \midrule
\top & \bot \\
\bot & \top \\
\udf & \udf
\end{array}
\;\
\begin{array}{c|lll}
 \wedge & \top & \udf & \bot \\
\midrule
\top & \top & \udf & \bot \\
\udf & \udf & \udf & \bot \\
\bot & \bot & \bot & \bot
 \end{array}
\;\
\begin{array}{c|lll}
 \vee & \top & \udf & \bot \\
\midrule
\top & \top & \top & \top \\
\udf & \top & \udf & \udf \\
\bot & \top & \udf & \bot
 \end{array}
\;\
\begin{array}{c|lll}
\leftarrow & \top & \udf & \bot \\
\midrule
\top & \top & \top & \top \\
\udf & \udf & \top & \top \\
\bot & \bot & \udf & \top
 \end{array}
\;\
\begin{array}{c|ccc}
  \leftrightarrow  & \top & \udf & \bot \\
  \midrule
\top & \top & \udf & \bot \\
\udf & \udf & \top & \udf\\
\bot & \bot & \udf & \top
 \end{array}
\;\
\begin{array}{c|c}
 L & \cxt(L)\\
\midrule
\top & \top \\
\bot & \bot  \\
\udf & \bot
\end{array}
\]
\caption{The truth tables for the connectives under the three-valued {\L}ukasiewicz logic and for $\cxt(L)$. $L$ is a literal, $\top$, $\bot$, and $\udf$ denote \textit{true}, \textit{false},
and \textit{unknown}, respectively.\label{tab:3vldluka}}
\end{table}

A set of \textit{integrity constraints} $\CalIC$ consists of clauses of
the form $\udf \leftarrow \Body$, where $\Body$ is a conjunction of
literals and $\udf$ denotes unknown. An interpretation maps
an integrity constraint to $\top$ iff $\Body$ is either mapped to
$\udf$ or $\bot$.  
Given an interpretation~$I$
and a set of integrity constraints~$\CalIC$,~$I$ \textit{satisfies}~$\CalIC$
iff all clauses in~$\CalIC$ are true under~$I$.

\subsection{Reasoning with Respect to the Least Fixed Point of $\Svlp$}

For a given $\CalP$, consider the following transformation:
1.\ For each ground atom~$A$ which occurs as head of a clause in $\Ground \CalP$, replace
  all clauses of the form $A \leftarrow \Body_1, \ldots,$ $A \leftarrow
  \Body_m$ occurring in $\Ground \CalP$ by $ A \leftarrow \Body_1 \vee
  \ldots \vee \Body_m$. 2.\ Replace all occurrences of $\leftarrow$ by $\leftrightarrow$.
The obtained ground set of equivalences is called the \textit{weak completion}
of~$\CalP$ or~$\WC\CalP$.
Consider the following semantic operator, which is due to Stenning and van Lambalgen~\cite{stenning:vanlambalgen:2008}: Let $I = \langle I^\top,I^\bot
\rangle$ be an interpretation. $\Phi_\CalP (I) =
\langle J^\top,J^\bot\rangle$, where
\[
\begin{array}{lcl}
J^\top & = & \{ A \mid A \leftarrow \Body \in \Def(A,\CalP)
\mbox{ and
} \Body \mbox{ is \True under } \langle I^\top,I^\bot \rangle \} \\
J^\bot & = & \{ A \mid \Def(A,\CalP) \neq \emptyset \mbox{ and }\\
&& \rule{7.5mm}{0mm}\Body \mbox{ is \False under } \langle
I^\top,I^\bot \rangle \mbox{ for all }  A \leftarrow \Body \in \Def(A,\CalP)\}.
\end{array}\Shrink
\]
\cite{hk:2009a} showed that the
weak completion of non-contextual programs\footnote{Non-contextual programs refer programs that do not contain rules, in which $\cxt$ appears in the body of any rule, i.e.\ $p = 0$.} 
always has a least model under  {\L}ukasiewicz logic,
which can be computed by the least fixed point of~$\Svlp$. The correspondence to other well-known logic programming approaches, in particular 
to the Well-founded semantics has been shown in~\cite{dietz:hoelldobler:wernhard:2014}.

The weak completion of contextual
programs might have more than one minimal model, and it is not guaranteed that 
the least fixed point of $\Svlp$ always exists. However,
if these programs do not contain cycles, then there always exists a supported model, which can be 
computed by iterating $\Svlp$: A least fixed point of $\Svlp$ always exists for contextual programs
that do not contain cycles~\cite{dietz:hoelldobler:pereira:2017}.
Let us denote the least fixed point of $\Svlp$ as $\CalM_\CalP$:
As the fixed point of $\Svlp$ is unique and only computes the supported minimal model of $\WComp\CalP$, we define
$\CalP \ModelsWCS F$ iff $\CalM_\CalP(F) = \top$.
Note that for all programs that are considered in the following, a least fixed point of $\Svlp$ exists.

% % \subsection{Backward Reasoning with respect to Skeptical Abduction} \label{sect:abduction}
% % 
% % An \textit{abductive framework} $\langle \CalP, \CalA, \CalIC, \ModelsWCS \rangle$
% % consists of a program $\CalP$, a set~$\CalA$ of abducibles, a set $\CalIC$ of
% % integrity constraints, and the entailment relation $\ModelsWCS$.  The set of
% % abducibles is~$\CalA =\{ A \leftarrow \top \mid A \in \ud(\CalP) \} \ \cup \ \{ A
% % \leftarrow \bot \mid A \in \ud(\CalP) \mbox{ and $\neg A$ is not assumed in
% % $\Ground \CalP$}\}$.
% % Let $\langle \CalP, \CalA, \CalIC, \ModelsWCS \rangle$ be
% % an abductive framework and the observation~$\CalO$ a set of literals.
% % $\CalO$ is \emph{explainable} in $\langle \CalP, \CalA, \CalIC, \ModelsWCS
% % \rangle$ iff there exists an $\CalE \subseteq \CalA$, such that
% % $\mathcal{P} \cup \mathcal{E} \models L$ for all $L \in \mathcal{O}$ and
% % $\mathcal{P} \cup \mathcal{E}$ satisfies $\CalIC$.  $\mathcal{E}$ is then
% % called \emph{explanation} for $\CalO$ given $\CalP$ and $\CalIC$.  We restrict
% % $\CalE$ to be \emph{minimal}, i.e.\ there does not exist any other explanation
% % $\CalE' \subseteq \CalA$ for $\CalO$ such that $\CalE' \subseteq \CalE$.
% % 
% % Among the minimal explanations, it is possible that some of them entail a certain formula $F$ while others do not. There exist two strategies to determine whether $F$ is a valid conclusion in such cases. $F$ follows \emph{credulously}, if it is entailed by at least one explanation. $F$ follows \emph{skeptically}, if it is entailed by all explanations.
% % Due to previous results on modeling human reasoning~\cite{btg:2015:h},
% % skeptical abduction seems to be adequate.
% % 
% % Here, observations, are specified as
% % $\mathcal{O}_\mathcal{P} = \{ A \mid A \leftarrow \top \in \Def(A, \CalP) \}$. Usually, this set is further restricted by
% % considering only facts that result from the application of certain principles.
% % %\leftarrow \top \in def(A, \mathcal{P}) \land (A \leftarrow B_1 \land \dots \land B_n) \i%n def(A, \mathcal{P}) \},
% % The idea is to find an explanation for each observation $A \in
% % \CalO_\CalP$ after the fact $A \leftarrow \top$ has been removed from
% % $\Ground\CalP$.

\section{Cognitive Principles}\label{sect:representing}

Eight principles for quantified statements have been developed in~\cite{cogsci:2017,costa:dietz:Hoelldobler:2017,declare:2017}.
In the following, the programs are
specified using the predicates $y$ and $z$, where~$yz$ can be replaced by $ab$, $ba$, $cb$, or $bc$ and depend on the figures shown in Table~\ref{t:2}. First, we introduce the basic principles that are assumed for all premises when applicable. Thus, we assume them to be assumed by all (human) reasoners. 
In Section~\ref{sect:advanced}, we present the advanced principles, which will 
later be used to distinguish between the clusters of reasoners.
% We indicate the cognitive principles in the text by specifying them in brackets.
A complete specification of the logic programs with respect to the moods according to all principles can be found in~\cite{declare:2017}.

\subsection{Basic Principles}   \label{sect:basic}

\subsubsection{Conditional representation, license for inference and existential import}
Consider \textit{All y are z} ({\MA yz}), which in CL, is represented as $\forall X (y(X) \rightarrow z(X))$.
Similarly, we assume that humans also understand this statement as a conditional sentence (\conditionals).
However, it seems that humans do not understand it as \textit{strict}, but defeasible~\cite{stenning:vanlambalgen:2008},
i.e.\ they implicitly assume a license for the inference (\licenses). For (\MA yz) this can be 
explicitly expressed by 
\textit{all $y$ that are not abnormal, are $z$}.
% \\
% Accordingly, $z(X)$ is conjoint with an abnormality predicate: 
% \[\forall X (y(X) \wedge \neg \Ab(X) \rightarrow z(X)).\] 
By default, we assume that \textit{nothing is abnormal}. % with respect to $X$}:
% \[
% \Ab(X) \leftarrow \bot.
% \]
Furthermore, humans have a pragmatic understanding of quantifiers in the language~\cite{grice1975}.
For instance, in natural language, we normally do not quantify over things that do not exist.
Consequently, \textit{All $z$} implies \textit{Some $z$ exist}.
This is known as the \textit{Gricean implicature} or the \textit{existential import} (\import). 
It also corresponds to the Aristotelian interpretation~\cite{parry:1991} and is assumed by several cognitive theories (cf.~\cite{johnsonlaird:1983,Rips1994}). 
 Likewise,~\cite{stenning:vanlambalgen:2008} have shown that humans
require existential import for a conditional to be true.
$\CalP_{Ayz}$, the logic program representing (\MA yz) is as follows:
\begin{align} %{rll@{\hspace{\hfill}}r}
z(X) & \leftarrow  y(X) \wedge \neg \Ab_{yz}(X). 
\tag{\conditionals\ \&\ \licenses} \\
\Ab_{yz}(X) & \leftarrow \cxt(z'(X)). \tag{\licenses}\\
y(o) & \leftarrow \top. \tag{\import}
\end{align}
Note that $z'(X)$ appears within the $\cxt$ operator in the body of the second clause, where $z'$ represents the negation of $z$:\footnote{$z'(X)$ can be introduced by the negation by transformation principle in the following.}
Nothing is abnormal, unless the negation of $z$ is true.
% , then $\Ab_{yz}(X)$ is false, for which then, given the first clause in $\CalP_{Ayx}$, $z(o)$ is true, if $y(o)$ is true. Note that, if $z'(X)$ would be written instead of $\cxt(z'(X))$, in the case where $\cxt(z'(o))$ is unknown, $z(o)$ could not be concluded, even if $y(o)$ would be true. This is due to the fact that negation in WCS is understood as strong negation.\footnote{This is different in other logic programming approaches, i.e.\ well-founded semantics.}
% \end{align}
% \]
We obtain \[\CalM_{\CalP_{Ayz}} = \langle \{y(o),z(o)\}, \{\Ab_{yz}(o)\} \rangle.\]

\subsubsection{Negation by transformation and no double negation}

Consider the statement \textit{No y is z.} ({\ME yz}):
Similarly to the previous case, (\licenses) and (\import) apply here
as well.
The logic programs specified in Subsection~\ref{sect:wcs} 
do not allow heads of clauses to be negative literals. Therefore, a negative conclusion~$\neg p(X)$ is represented
by introducing an auxiliary formula $p'(X)$ together with the clause $p(X) \leftarrow \neg p'(X)$ and the integrity constraint $\udf \leftarrow p(X) \wedge p'(X)$ (\transformation). This is a widely used technique in Logic Programming. 
Another principle that we introduce is the no double negation principle: Under the Weak Completion Semantics, a positive conclusion can be derived from double negation using two conditionals. However, it appears to be the case that humans do not reason in such a way~\cite{khemlani:2012}. Hence, we block them with the help of abnormalities (\dnegation). $\CalP_{Eyz}$, the logic program representing (\ME yz) is as follows:
% \footnote{For simplicity, we omitted to list the clauses that are introduced by the (\converse) principle in $\CalP_{Eyz}$: The same clauses just switching $y$ and $z$ are added and thus also the fact $z(o_2) \leftarrow \top$ for another object $o_2$.
% The entire program can be found in~\cite{declare:2017}.}
\begin{align}
 z'(X) & \leftarrow y(X) \wedge \neg \Ab_{ynz}(X).  \tag{\conditionals\ \&\ \transformation\ \&\  \licenses} \\
\Ab_{ynz}(X)  & \leftarrow \cxt(z(X)).\tag{\licenses} \\
z(X) & \leftarrow \neg z'(X) \wedge \neg \Ab_{nzz}(X). \tag{\transformation\ \&\ \licenses} \\
y(o_1) & \leftarrow \top. \tag{\import} \\
\Ab_{nzz}(o_1) & \leftarrow \bot. \tag{\licenses\ \&\  \dnegation}
\end{align}
As we now have both $z(X)$ and $z'(X)$ in the head of some clause, we need to add the 
following integrity constraints: $\udf \leftarrow z(X) \wedge
z'(X)$. We obtain 
\[\CalM_{\CalP_{Eyz}} = \langle \{y(o_1), z'(o_1)\},\{\Ab_{ynz}(o_1), \Ab_{nzz}(o_1),  z(o_1)\}
\rangle.
\]

\subsubsection{Converse implication and unknown generalization}
Consider the statement \textit{Some y are z.} ({\MI yz}), which in CL, is represented as $\exists X (y(X) \wedge z(X))$
and which is semantically equivalent to $\exists X(z(X) \wedge y(X))$ (\converse).
As~\cite{khemlani:2012}
shows, humans seem to
distinguish between \textit{some y are z} and \textit{all y are z} (\unknownGen).
Accordingly, if we only observe that an object $o$ belongs to $y$ and $z$ then we do not want to conclude both, \textit{some y are z} and \textit{all y are z}.
In order to distinguish between \textit{some y are z} and \textit{all y are z}, we introduce the following principle:  If we know that \textit{some y are z}, then there must not only be an object~$o_1$, which belongs to $y$ and $z$ but there must be another object $o_2$, which belongs to
$y$ and for which it is unknown whether it belongs to~$z$.
$\CalP_{Iyz}$, the logic program representing (\MI yz), is as follows:\footnote{For simplicity, here we omit the clauses introduced by ({\small\converse}). They can be found in~\cite{declare:2017}.}
\begin{align}
z(X) & \leftarrow y(X) \wedge \neg \Ab_{yz}(X). & \tag{\conditionals\ \&\ \licenses}\\
\Ab_{yz}(o_1) & \leftarrow  \bot. & \tag{\unknownGen\ \& \ \licenses}\\
y(o_1) & \leftarrow  \top. & \tag{\import} \\
y(o_2) & \leftarrow  \top. & \tag{\unknownGen} 
\end{align}
Consider the second clause: Different to the universal moods, \ME\ and \MA, $\Ab_{yz}$ is only false for $o_1$, which in turn implies that 
$\Ab_{yz}$ is unknown for all other objects in the program.
We obtain $\CalM_{\CalP_{Iyz}}= \langle \{y(o_1), y(o_2), z(o_1)\}, \{\Ab_{yz}(o_1)\} \rangle$.
One should observe that  $\Ab_{yz}(o_2)$ is an unknown assumption in
$\CalP_{Iyz}$ and,
hence, $\CalM_{\CalP_{Iyz}}(z(o_2)) = \udf$.

Consider the statement \textit{Some y are not z} ({\MO yz}), which in CL is represented as $\exists X (y(X) \wedge \neg z(X)$.
As in the previous cases, (\conditionals), (\licenses) and (\unknownGen) apply.
Furthermore, as the conclusion is again a negation of a claim, we need to assume (\transformation) and
(\dnegation). $\CalP_{Oyz}$, the logic program representing (\MO yz) is as follows:
\begin{align}
z'(X) & \leftarrow y(X) \wedge \neg \Ab_{ynz}(X). & \tag{\conditionals\ \&\ \transformation \ \&\ \licenses} \\
\Ab_{ynz}(o_1) & \leftarrow \bot. & \tag{\licenses\ \& \unknownGen} \\
z(X) & \leftarrow  \neg z'(X) \wedge \neg \Ab_{nzz}(X). &
\tag{\licenses\ \&\ \transformation} \\
y(o_1) & \leftarrow \top. & \tag{\import} \\
y(o_2) & \leftarrow \top. & \tag{\unknownGen} \\
\Ab_{nzz}(o_1)& \leftarrow \bot. & \tag{\licenses\ \&\ \dnegation}\\
\Ab_{nzz}(o_2)& \leftarrow \bot. & \tag{\licenses\ \&\ \dnegation}
\end{align}
We have to add the integrity constraint $U \leftarrow z(X) \wedge
z'(X)$
and obtain \[\CalM_{\CalP_{Oyz}} = \langle \{y(o_1), y(o_2), z'(o_1)\},\{\Ab_{ynz}(o_1),
\Ab_{nzz}(o_1), \Ab_{nzz}(o_2), z(o_1) \} \rangle.\]

\subsection{Advanced Principles}\label{sect:advanced}

The principles introduced here, are assumed not to be applied by all humans, 
but characterize the reasoning clusters, that will be specified in the following section.

\subsubsection{Contraposition}

In FOL, the conditional statement $\forall (X) (a(X) \rightarrow b(X))$, is logically
equivalent to its contrapositive, $\forall (X) (\neg b(X) \rightarrow \neg a(X))$.
% This
% contraposition also holds for the syllogistic moods~\MA\ and~\ME. 
There is
evidence that some of the participants make use of this
equivalence when solving syllogistic reasoning tasks~\cite{khemlani:2012}. We believe that when they
encounter a premise with the mood \MA\ (e.g., \textit{all~a are~b}),
then might reason with the
contrapositive conditional as well (\contraposition).
Consider again $\CalP_{Ayz}$, the logic program representing (\MA yz)
from Section~\ref{sect:basic}. The clauses that represent the contraposition principles are as follows:
\begin{align}
y'(X) & \leftarrow  \neg z(X) \wedge \neg \Ab_{zy}(X). & \tag{\conditionals\ \&\ \licenses\ \&\ \contraposition\ } \\
\Ab_{zy}(X) & \leftarrow \cxt(y(X)). & \tag{ \licenses\ \&\ \contraposition}\\
y(X) & \leftarrow  \neg y'(X) \wedge \neg \Ab_{nyy}(X). & \tag{\transformation\ \&\ \licenses\ \&\ \contraposition} 
\end{align}
The application of (\contraposition) for the \ME\ mood,
\textit{No y is z}, yields \textit{No z is y}, and consists of the following clauses:
\begin{align}
y'(X) & \leftarrow z(X) \wedge \neg \Ab_{zny}(X). & \tag{\conditionals\ \&\ \licenses\ \&\ \contraposition\ } \\
\Ab_{zny}(X) & \leftarrow \cxt(y(X)). &\tag{ \licenses\ \&\ \contraposition}\\
y(X) & \leftarrow \neg y'(X) \wedge \neg \Ab_{nyy}(X). & \tag{\transformation\ \&\ \licenses\ \&\ \contraposition} 
% z(o_2) \leftarrow \top. & (\contraposition\  \&\ \import) \\
% \Ab_{nyy}(o_2) \leftarrow \bot. & (\contraposition\  \&\ \licenses\ \&\ \dnegation)
\end{align}

% Ignoring unknown information in conditionals is known as
% negation as failure~\cite{clark:1978} in logic programming.

\subsubsection{Generalization}  %\label{sect:naf}

If all of the principles introduced so far are applied to an existential
premise, the only object about which an inference can be made is the one
resulting from (\import). This is because the
abnormality introduced by (\licenses) and according to (\unknownGen)
 has to be false only for the object introduced by (\import), but unknown
for other objects.
There is, however, evidence that some humans still draw conclusions in such
circumstances~\cite{khemlani:2012}, i.e.\ they generalize (universally quantify) the existential 
premise with respect to other objects (\negFailure).
Consider again $\CalP_{I yz}$ of Section~\ref{sect:basic}, where $\Ab_{yz}$ is only false for 
$o_1$. In case (\negFailure) applies, then $\Ab_{yz}(o_1) \leftarrow \bot$
in $\CalP_{I yz}$ is replaced by the following two clauses:
\begin{align}
\Ab_{yz}(X) & \leftarrow \cxt(z'(X)). & \tag{\licenses\ \&\ \negFailure} \\
\Ab_{yz}(o_2) & \leftarrow \udf. & \tag{\licenses\ \&\ \negFailure} 
\end{align}
Similar to the case of $\CalP_{Ayz}$ for the representation of the basic principles of (\MA yz),
the $\cxt$ operator in the body of the first clause is false for all $o$
for which $z'(o)$ is unknown. This clause transforms the representation for the existential quantification over \MI yz in $\CalP_{Iyz}$
into a universal one. In order to maintain (\unknownGen), the second clause is added:
$\Ab_{yz}(o_2) \leftarrow \udf$ guarantees that $\Ab_{yz}(o_2)$ stays unknown, even though $\cxt(z'(o_2))$ might false.
Recall, programs are considered under their weak completion, i.e.\ if $\cxt(z'(o_2))$ is false, then
\[\Ab_{yz}(o_2) \leftrightarrow (\cxt(z'(o_2)) \vee \udf)
 \equiv \Ab_{yz}(o_2) \leftrightarrow \udf.\]
%  is semantically equivalent to $\Ab_{yz}(o_2) \leftrightarrow \udf$ if $\cxt(z'(o_2))$ is false.}

\subsubsection{Search for Alternative Conclusions}

When participants are faced with a \NVC\ conclusion (\textit{No valid conclusion}), they might not want to accept this conclusion and proceed to check whether there exists unknown information that might be relevant (\abduction). This information may be explanations about
facts introduced either from an existential import or from unknown generalization. 
We use only the first as source for observations, as they are used directly to infer new information. 
Consider the following pair of syllogistic premises: 
\begin{align}
 \emph{All artists are bakers. } & \emph{Some bakers are not chemists.} \tag{AO1} 
\end{align}
Classical logically \textit{No valid conclusion} follows. However, 62\% of the participants concluded \textit{Some artists are not chemists}.
This conclusion can be modeled by abductive reasoning, where the system searches for explanations
for a given observation. The second premise states that some bakers exist, i.e.\ the observation is $\CalO = \{b(o_2) \}$. $\CalO$ can be explained by the first premise, namely that
these bakers are artists. Under WCS, we formalize this reasoning procedure by means of skeptical abduction for logic programming~\cite{kakas:etal:93}.
Due to lack of space we will not introduce the formal abductive approach for human syllogistic reasoning in this paper.
The interested reader is referred to~\cite{cogsci:2017}.

\subsection{Entailment of Conclusions}\label{sect:entail}

We specify when $\MP$ entails a conclusion with respect to $yz$:\footnote{$yz$ can be replaced by $ac$ or~$ca$.}
\begin{description}
\item[\MA yz (all)] $\CalP \models Ayz$ iff there exists an object $o$ such that $\CalP \ModelsWCS
y(o)$ and for all objects $o$ we find that if $\CalP \ModelsWCS y(o)$ then $\CalP \ModelsWCS z(o)$.
\item [\ME yz (no)]
$\CalP \models Eyz$ iff there exists an object $o_1$ such that $\CalP \ModelsWCS
y(o_1)$ and for all objects $o_1$ we find that if $\CalP \ModelsWCS y(o_1)$ then
$\CalP \ModelsWCS \neg z(o_1)$.\footnote{
When the contraposition principles applies, then additionally 
the following condition needs to hold: There exists an object $o_2$ such that $\CalP \ModelsWCS
z(o_2)$ and for all objects $o_2$ we find that if $\CalP \ModelsWCS z(o_2)$ then
$\CalP \ModelsWCS \neg y(o_2)$.}
\item[\MI yz (some)]
$\CalP \models Iyz$ iff there exists an object $o_1$ such that $\CalP \ModelsWCS
y(o_1) \wedge z(o_1)$ and there exists an
object $o_2$ such that $\CalP \ModelsWCS y(o_2)$ and
$\CalP \not\ModelsWCS z(o_2)$ and
% there exists an object $o_3$ such that $\CalP \ModelsWCS
% z(o_3) \wedge y(o_3)$ and 
there exists an
object $o_3$ such that $\CalP \ModelsWCS z(o_3)$ and
$\CalP \not\ModelsWCS y(o_3)$.
\item[\MO yz (some are not)]
$\CalP \models Oyz$ iff there exists an object $o_1$ such that $\CalP \ModelsWCS
y(o_1) \wedge \neg z(o_1)$ and there exists an
object $o_2$ such that $\CalP \ModelsWCS y(o_2)$ and
$\CalP \not\ModelsWCS \neg z(o_2)$.
\item[\NVC] When no previous conclusion can be derived, no valid conclusion holds.
\end{description}
For \MI yz (some) the conversion is explicitly specified, motivated by it (\converse) principle.
By requiring for \MI\ and \MO, the existence of an object with the first property,
for which the second property has to be either either false or unknown, \MA\ and \MI, and, \ME\ and \MO, exclude each other, respectively.

\section{Preliminary Evaluation} \label{sect:aggregateddata}

Let us again consider \mbox{\MI\ME 2} from the introduction: 
\begin{align}
 \emph{Some bakers are artists. } & \emph{No chemists are bakers.} \tag{\MI\ME 2} 
\end{align}
The majority of participants in experimental studies
concluded \emph{No chemists are artists} (29\%), \emph{No valid conclusion} (27\%)
and \emph{Some artists are not chemists} (16\%): Participants who concluded the first answer, did not conclude the other two answers, and vice versa.
In order to account for these individual differences, three clusters on principles and two on heuristic strategies have been suggested
in~\cite{declare:2017}. The two heuristic strategies are the atmosphere bias~\cite{woodworth:sells:1935} and the matching bias~\cite{Wetherick1995}.\footnote{
The interested reader is referred to~\cite{declare:2017} for details on their application on the syllogistic reasoning task. 
}
The characterization of the three clusters specified through the cognitive principles is shown Table~\ref{tab:cluster}, together with the conclusions for \MI\ME 2 according to each cluster. 
\begin{table}
\begin{center}
\begin{tabular}{@{\hspace{0cm}}l@{\hspace{0.9cm}}c@{\hspace{0.9cm}}c@{\hspace{0.9cm}}cGG}
Principle & Basic & Contraposition & Generalization & Heuristic 1 & Heuristic 2 \medskip \\ \midrule  
\conditionals & $\checkmark$   & $\checkmark$ & $\checkmark$ & - & - \smallskip\\
\licenses & $\checkmark$ & $\checkmark$       & $\checkmark$ & - & -   \smallskip \\
\import  & $\checkmark$  & $\checkmark$       & $\checkmark$ & - & - \smallskip\\
\unknownGen & $\checkmark$  & $\checkmark$    & $\checkmark$ & - & -     \smallskip\\
\abduction &  - & -             & -            & - & - \smallskip\\
\converse\ for \MI & $\checkmark$  & $\checkmark$      & $\checkmark$ & - & -   \smallskip\\
\converse\ for \ME & -  & $\checkmark$      & -  & - & -   \smallskip\\
{\small\textsf{contraposition}} & - & $\checkmark$            & -            & - & - \smallskip\\\midrule
% matching   &  -            & -             & -            & $\checkmark$ & -   \smallskip\\
% biased fig 1 & -     & -             & -            & -  & $\checkmark$ \medskip \\ \midrule
 &  $\Downarrow$ & $\Downarrow$ &  $\Downarrow$ & $\Downarrow$  &  $\Downarrow$ \\
 &  \NVC &  \textit{Some artists are not chemists}  &   \textit{No chemists are artists} \\
%            &       & \textit{} & \textit{}
\end{tabular}
\caption{Cognitive principles and the corresponding entailments for \MI\ME 2.
\label{tab:cluster}}
\end{center}
\end{table}
The corresponding logic program $\Pie$ that represents the basic cluster, consists of the following clauses:
\begin{align}
a(X) & \leftarrow b(X) \wedge \neg \Ab_{ba}(X). & \tag{\conditionals\ \&\ \licenses}\\
\Ab_{ba}(o_1) & \leftarrow  \bot. & \tag{\licenses\ \& \unknownGen}\\
b(o_1) & \leftarrow  \top. & \tag{\import} \\
b(o_2) & \leftarrow  \top. & \tag{\unknownGen}
% \medskip \\
\end{align}
\begin{align}
b(X) & \leftarrow a(X) \wedge \neg \Ab_{ab}(X). & \tag{\conditionals\ \&\ \licenses\ \&\ \converse}\\
\Ab_{ab}(o_3) & \leftarrow  \bot. & \tag{\unknownGen\ \& \ \licenses\ \&\ \converse}\\
a(o_3) & \leftarrow  \top. & \tag{\import\ \&\ \converse} \\
a(o_4) & \leftarrow  \top. & \tag{\unknownGen\ \& \converse} 
\medskip \\
% \end{align}
% \begin{align}
b'(X) & \leftarrow c(X) \wedge \neg \Ab_{cnb}(X).  \tag{\licenses\ \&\ \transformation} \\
\Ab_{cnb}(X)  & \leftarrow \cxt(b(X)).\tag{\licenses} \\
b(X) & \leftarrow \neg b'(X) \wedge \neg \Ab_{nbb}(X). \tag{\licenses\ \&\ \transformation} \\
c(o_5) & \leftarrow \top. \tag{\import} \\
\Ab_{nbb}(o_5) & \leftarrow \bot. \tag{\licenses\ \&\  \dnegation}
\end{align}
We obtain $\CalM_\Pie = \langle I^\top, I^\bot \rangle$, where 
\[
\begin{array}{lll}
 I^\top & = & \{ b(o_1), a(o_1), b(o_2), a(o_3), b(o_3),  a(o_4), c(o_5), b'(o_5), \Ab_{cnb}(o_1), \Ab_{cnb}(o_2), \Ab_{cnb}(o_3)  \}, \\
I^\bot & = & \{ \Ab_{ba}(o_1), \Ab_{ab}(o_3), \Ab_{cnb}(o_5), \Ab_{nbb}(o_5), \Ab_{cnb}(o_4) \}, 
\end{array}
\]
for which according to the entailment of conclusions as defined in Section~\ref{sect:entail}, \textit{No valid conclusion}
follows.

Applying the contraposition principle to \textit{No chemists are bakers} yields to \textit{No bakers are chemists}.
$\Piecontra$ consists of the following clauses:
\begin{align*}
\Pie \cup \{ c'(X) & \leftarrow b(X) \wedge \neg \Ab_{bnc}(X), & \tag{\basic\ \&\ \licenses\ \&\ \transformation\  \&\  \contraposition} \\
\Ab_{bnc}(X) & \leftarrow \cxt(c(X)), & \tag{\contraposition\  \&\ \licenses} \\
c(X) & \leftarrow \neg c'(X) \wedge \neg \Ab_{ncc}(X), &  \tag{\licenses\ \&\ \transformation\  \&\  \contraposition} \\
b(o_6) & \leftarrow \top, & \tag{\import\ \&\ \contraposition} \\
\Ab_{ncc}(o_6) & \leftarrow \bot  \} & \tag{\licenses\ \& \ \dnegation\ \&\ \contraposition}
\end{align*}
We obtain $\CalM_{\Piecontra} = \langle I^{\top'}, I^{\bot'} \rangle$ where 
\[
\begin{array}{lllll}
 I^{\top'} & = & I^\top & \cup & \{ b(o_6), c'(o_3), c'(o_1), c'(o_2), c'(o_6), \Ab_{bnc}(o_5), \Ab_{cnb}(o_6) \}, \\
I^{\bot'} & =  & I^\bot  & \cup & \{  \Ab_{bnc}(o_i) \mid i \in \{1,2,3,4,6 \} \} 
\\ &&& \cup  &
\{ c(o_i), b'(o_1) \mid i \in \{ 1, 2, 3, 6 \} \} \cup \Ab_{cnb}(o_3), \Ab_{ncc}(o_6) \}.
\end{array}
\]
\textit{Some artists are not chemists} is entailed by $\CalM_{\Piecontra}$ because 
$a(o_1)$ is true and $c(o_1)$ is false and $a(o_4)$ is true, but $c(o_4)$ is unknown.

Applying the generalization principle to \textit{Some bakers are artists} and its converse,
yields to \textit{All bakers are artists} and \textit{All artists are bakers}.
$\Piegen$ consists of the clauses in $\CalP_{IE2}$ and the following ones:
\begin{align}
\Pie \cup \{  \Ab_{ba}(X) & \leftarrow \cxt(a'(X)),  \Ab_{ba}(o_2) \leftarrow \udf, & \tag{\basic\ \&\ \licenses\ \&\ \negFailure} \\
  \Ab_{ab}(X) & \leftarrow \cxt(b'(X)),   \Ab_{ab}(o_4) \leftarrow \udf \}  & \tag{\licenses\ \&\ \negFailure\ \&\ \converse}
\end{align}
We obtain $\CalM_{\Piegen}  = \langle I^{\top''}, I^{\bot''} \rangle$ where 
\[
\begin{array}{lllll}
 I^{\top''} & = & I^\top & \cup & \{\Ab_{ab}(o_5)  \}, \\
I^{\bot''} & = & I^\bot & \cup & \{ a(o_5),  b(o5) \cup \{ \Ab_{ab}(o_i) \mid i \in \{ 1,2\}\} \\
& & & \cup &
\{ \Ab_{ba}(o_i) \mid i \in \{3, 4, 5 \} \} \cup \{ \Ab_{cnb}(o_3) \mid i \in \{1,2,3 \} \}.
\end{array}
\]
\textit{No chemists are artists} is entailed by $\CalM_{\Piegen}$ because 
for all $o$ for which $c(o)$ is true,
which only applies for $c(o_5)$, we find that $a(o_5)$ is false.

WCS clustering~\cite{declare:2017} suggests three reasoning clusters characterized by the following principles:\footnote{\textsf{basic} refers to the set of basic principles introduced in Subsection~\ref{sect:basic}} (1) \basic, \abduction, and \converse\ for \MI,
    (2) \basic, \converse\ for \MI\ and \negFailure, and
    (3) \basic, \converse\ for \MI, \ME, and \contraposition\ for~\MA.\footnote{\cite{declare:2017} also included two heuristic clusters based on
         atmosphere bias~\cite{Woodworth1935} and matching bias~\cite{Wetherick1995}.}
This approach achieved a match of 92\% with the aggregated data reported in~\cite{khemlani:2012}, and was the highest result compared to the presented approaches in~\cite{khemlani:2012}.

We were interested whether these clusters were also cognitively adequate in the sense that they 
would also correspond to the reasoning patterns of individuals. Therefore, we investigated the non-aggregated dataset provided by~\cite{khemlani:2016} consisting of the responses of 20 participants for two weeks,
i.e. the participants carried out the task twice, with a break of one week. The results with respect to WCS clustering is two-fold:
\begin{description}
 \item[Aggregated evaluation] Most of the majority's answers can be explained by the clusters' predictions. In the first week 
 the majority's conclusion for 63 out of 64 pair of syllogistic premises and in the second week
 all majority's conclusions were predicted by WCS clustering.
\item[Non-aggregated evaluation] None of the clusters fitted any individual participants' response pattern, i.e.\ no participant did belong to a single cluster.
\end{description}
Given this sobering outcome of the non-aggregated evaluation, 
we investigated whether patterns could be found on the level of the
cognitive principles: \basic, \contraposition, \abduction\ and \negFailure.
Table~\ref{tab:cogprinc} provides an overview of the predicted conclusions according to these principles.\footnote{Note that any advanced principle also includes the logic program clauses specified for \textsf{basic}.}
The predictions of the principles naturally raises a few questions on the plausibility of 
the underlying assumptions:
(i) Do participants use the same reasoning strategy throughout the same reasoning task, i.e.\ 
can their conclusions be 
explained by  the same principles throughout the whole reasoning task, or do they switch strategies within one reasoning task?
(ii) Is the list of principles in Section~\ref{sect:representing} complete, i.e.\ can the conclusions of all reasoners be covered
 by these principles? and
(iii) Do participants learn, i.e.\ do they improve their logical ability by doing the same reasoning task again.
We address these questions in the following section.

\clearpage
\begin{table}[t]\betweentinyandsmall\centering
\newcolumntype{a}{>{\columncolor{white}}c@{\hspace{2mm}}}
\newcolumntype{b}{>{\columncolor{gray!40}}c@{\hspace{2mm}}}  
\captionsetup{font=footnotesize}
 \begin{tabular}{llllllllll}
 \toprule
	 & \MA ac & \ME ac & \MI ac & \MO ac & \MA ca & \ME ca & \MI ca & \MO ca & \NVC \\ \midrule

\rowcolor{lightgray} 
\MA\MA1  & \basicc&        &        &        &        &        &        &        &      \\ 
\MA\MA2  &        &        &        &        & \basicc&        &        &        & \\ \rowcolor{lightgray} 
\MA\MA3  &        &        &        &        &        &        &        &        & \basicc \\
\MA\MA4  & \basicc&        &        &        &\basicc &        &        &        & \\  \rowcolor{lightgray}  %\midrule

\MA\MI1  &        &        & \abdc  &        &        &        & \abdc  &        & \basicc \\
\MA\MI2  &        &        & \basicc&        &        &        & \basicc \\  \rowcolor{lightgray} 
\MA\MI3  &        &        & \abdc  &        &        &        & \abdc  &        & \basicc \\
\MA\MI4  &        &        & \basicc&        &        &        & \basicc&        & \\%\midrule

\rowcolor{lightgray} 
\MA\ME1  &        & \basicc&        &        &        &        &        &        & \\ %\rowcolor{lightgray} 
\MA\ME2  &        &\contrac&        &        &        & \basicc&        &        & \\ \rowcolor{lightgray} 
\MA\ME3  &        &\contrac&        &        &        &\contrac&        &        & \basicc \\ %\rowcolor{lightgray} 
\MA\ME4  &        & \basicc&        &        &        &\contrac&        &        &  \\  \rowcolor{lightgray}  %\midrule

\MA\MO1  &        &        &        &\abdc   &        &        &        &        & \basicc \\
\MA\MO2  &        &        &        &        &        &        &        &\basicc \\  \rowcolor{lightgray} 	
\MA\MO3  &        &        &        &        &        &        &        &\contrac&\basicc \\
\MA\MO4  &        &        &        &\basicc \\ %\midrule

\rowcolor{lightgray} 
\MI\MA1  &        &        &\basicc &        &        &        &\basicc &        & \\ % \rowcolor{lightgray} 
\MI\MA2  &        &        &\abdc   &        &\genc   &        & \abdc  &        & \basicc \\ \rowcolor{lightgray} 
\MI\MA3  &        &        & \abdc  &        &        &        & \abdc  &        & \basicc \\ % \rowcolor{lightgray} 
\MI\MA4  &        &        &\basicc &        &        &        &\basicc &        &  \\  %\midrule
 \rowcolor{lightgray} 
\MI\MI1  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\
\MI\MI2  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\ \rowcolor{lightgray} 
\MI\MI3  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\
\MI\MI4  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\ %\midrule

\rowcolor{lightgray} 
\MI\ME1  &        &        &        &\basicc &        &        &        &        & \\ %\rowcolor{lightgray} 
\MI\ME2  &        &        &        &\contrac&        & \genc  &        &        & \basicc \\\rowcolor{lightgray} 
\MI\ME3  &        &        &        &\contrac&        & \genc  &        &        & \basicc  \\% \rowcolor{lightgray} 
\MI\ME4  &        &        &        & \basicc&        &        &        &        & \\ %\midrule 
 \rowcolor{lightgray} 
\MI\MO1  &        &        &        & \genc  &        &        &        &        & \basicc \\
\MI\MO2  &        &        &        & \genc  &        &        &        &        & \basicc \\  \rowcolor{lightgray} 
\MI\MO3  &        &        &        & \genc  &        &        &        &        & \basicc \\ 
\MI\MO4  &        &        &        & \genc  &        &        &        &        & \basicc \\%\midrule

\rowcolor{lightgray} 
\ME\MA1  &        & \basicc&        &        &        &        &        &        & \\ %\rowcolor{lightgray} 
\ME\MA2  &        &\contrac&        &        &        & \basicc&        &        &  \\\rowcolor{lightgray} 
\ME\MA3  &        &\contrac&        &        &        &\contrac&        &        & \basicc \\% \rowcolor{lightgray} 
\ME\MA4  &        &\contrac&        &        &        & \basicc&        &        &\\  %\midrule 
 \rowcolor{lightgray} 
\ME\MI1  &        & \genc   &        &	 &	  &	   &	    &	     & \basicc \\
\ME\MI2  &        &	       &	&	 &	  &	   &	    & \basicc \\ \rowcolor{lightgray}  
\ME\MI3  &        & \genc   &        &	 &	  &	   &	    &\contrac& \basicc \\
\ME\MI4  &        &	       &	&	 &	  &	   &	    & \basicc \\%\midrule

\rowcolor{lightgray} 
\ME\ME1  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
\ME\ME2  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ \rowcolor{lightgray} 
\ME\ME3  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
\ME\ME4  &        &	       &	&	 &	  &	   &	    &        &\basicc \\%\midrule
 \rowcolor{lightgray} 
\ME\MO1  &        &	       &	&	 &	  &	   &	    &        &\basicc \\
\ME\MO2  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ \rowcolor{lightgray} 
\ME\MO3  &        &	       &	&	 &	  &	   &	    &        &\basicc \\
\ME\MO4  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ %\midrule

\rowcolor{lightgray} 
\MO\MA1  &        &            &	& \basicc&        &        &        &        & \\ %\rowcolor{lightgray} 
\MO\MA2  &        &	       &	&	 &	  &	   &	    &  \abdc &\basicc \\\rowcolor{lightgray} 
\MO\MA3  &        &	       &	&\contrac&	  &	   &	    &        &\basicc \\% \rowcolor{lightgray} 
\MO\MA4  &        &	       &	&	 &	  &	   &	    & \basicc& \\  %\midrule
 \rowcolor{lightgray} 
\MO\MI1  &        &	       &	&\genc   &	  &	   &	    &        &\basicc \\
\MO\MI2  &        &	       &	&	 &	  &	   &	    &\genc   &\basicc \\  \rowcolor{lightgray} 
\MO\MI3  &        &	       &	&\genc   &	  &	   &	    &        &\basicc \\ 
\MO\MI4  &        &	       &	&	 &	  &	   &	    &\genc   &\basicc\\ %\midrule

\rowcolor{lightgray} 
\MO\ME1  &        &	       &	&        &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
\MO\ME2  &        &	       &	&        &	  &	   &	    &        &\basicc \\\rowcolor{lightgray} 
\MO\ME3  &        &	       &	&        &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
\MO\ME4  &        &	       &	&        &	  &	   &	    &        &\basicc \\ %\midrule
\rowcolor{lightgray} 
\MO\MO1  &        &	       &	&        &	  &	   &	    &        &\basicc \\
\MO\MO2  &        &	       &	&        &	  &	   &	    &        &\basicc \\\rowcolor{lightgray}  
\MO\MO3  &        &	       &	&        &	  &	   &	    &        &\basicc \\
\MO\MO4  &        &	       &	&        &	  &	   &	    &        &\basicc \\ \bottomrule
  \end{tabular}

\caption{\label{tab:cogprinc}Predictions according to the cognitive principles as specified in Section~\ref{sect:representing}.\betweentinyandsmall}
 \end{table}

\thispagestyle{empty}
\clearpage

\section{On non-aggregated Data} \label{sect:nonag}

A wide amount of cognitive theories for syllogistic reasoning have previously only be evaluated with respect to the aggregated data of psychological experiments.
Usually, as assessment, typical statistical evaluation criteria with respect to the experimental data 
 have been used (cf. root-mean-square error, Bayesian information criterion~\cite{schwarz1978}, Akaike information criterion~\cite{akaike1974new}).
 These measures are helpful as a general guideline, however, their major drawback is that they only compute the quantitative overall performance.
 In this section, we carry out a qualitative assessment of the suggested cognitive principles with respect to the individual differences
  among the participants.
% As already mentioned in the introduction, an issue with these evaluations is that these theories cannot account for individual differences. 
% Therefore, in this section, we will investigate an approach that accounts for non-aggregated data of participants in syllogistic reasoning.

\subsection{mReasoner}

\cite{khemlani:2016} proposed an approach to model the individual reasoning patterns of the participants, and 
showed that different participants seem to derive different conclusions.
Their approach is built on top of the mReasoner~\cite{khemlani:2013}, an extension of the mental model theory~\cite{johnsonlaird:1983}.
Reasoning depends on the construction and manipulation of mental models, based on 3 fundamental principles: 
(i) represent possibilities (ii) principle of iconicity and (iii) principle of dual processes.
 The procedure in the mReasoner consists of two steps:
(1) The construction of a small set of models that represents the terms of the premises, and
(2) the call of a deliberative component to search for counterexamples to conclusions.
They specified four parameters for the construction of a mental model and evaluated their
predictions based on 2 datasets, week 1 and week 2, for 20 participants.
The parameters for the construction of the mental model are (i) the 
size of a mental model (ii) the probability of drawing the complete set 
(iii) the probability to search for counter examples 
and (iv) the probability to weaken the conclusion.
The clustering with the mReasoner was based on partitioning around medoids clustering algorithm to determine the optimal number of clusters.
The K-means clustering algorithm was used to divide participants into clusters and then
the mReaoner to simulated the clusters by choosing appropriate parameter settings.
\cite{khemlani:2016} identified three different reasoning clusters: The intuitive cluster, the intermediate cluster and the deliberative cluster.
The approach is novel in the sense that until now no other cognitive theory tried to rigorously account for individual differences in human syllogistic reasoning. However, their approach seems to have two drawbacks. First, some of their 
parameters were of probabilistic nature, i.e.\ there is no explaining factor behind them. Second, the authors did not distinguish between the participants' 
responses in week 1 and week 2, but instead merged the responses. This might give a distorted picture, as participants could
have learned from week 1 to week 2.

% The clusters described in~\cite{declare:2017} seem not to correspond exactly to the individual participants. 
% and observed an interesting effect: Participants seemed to improve their logical ability when carrying out the task for the second time.
% Furthermore, participants seemed not to reason according to one specific cluster during the whole task. This puts forward
% two possible assumptions: Either the clusters in~\cite{declare:2017} were not well defined or people changed their reasoning techniques within the task. 
% % According to these new findings, we formulate the hypothesis that the clusters do not represent different human reasoners, but different stages of reasoning. Depending on the persistence and the previous training of the individual participants, they might be in different stages.
% taking as starting point the principles in~\cite{cogsci:2017}.
% This dataset consists of the individual conclusions of 20 participants, who carried out the same 
% task two times, with an interval of one week.

\subsection{Evaluation}

\begin{table}[htp]
\betweentinyandsmall\centering
\newcolumntype{a}{>{\columncolor{white}}c@{\hspace{2mm}}}
\newcolumntype{b}{>{\columncolor{gray!40}}c@{\hspace{2mm}}}  
\captionsetup{font=footnotesize}
\begin{tabular}{cccccccccccc}\toprule
Week & Part. & NVC & CL & \basicc & \contrac & \abdc & \genc & Changes & Cluster & Cluster in~\cite{khemlani:2016}  \\ \midrule 
\rowcolor{lightgray} 
1&0&$\downarrow$&$\downarrow$&$\downarrow$&$\downarrow$&$\downarrow$&$\uparrow$& & &  \\
\rowcolor{lightgray}  
2&0&$\downarrow$&$\leftrightarrow$&$\leftrightarrow$&$\uparrow$ $\uparrow$&$\downarrow$&$\downarrow$ &
\multirow{-2}{*}{$\uparrow$}  & \multirow{-2}{*}{INTUITIVE}& 
\multirow{-2}{*}{INTUITIVE} \\

1&1&$\downarrow$&$\downarrow$&$\downarrow$&$\uparrow$&$\uparrow$& $\uparrow$& &&\\
2&1&$\uparrow$&$\uparrow$&$\uparrow$&$\leftrightarrow$&$\uparrow$&$\downarrow$& \multirow{-2}{*}{$\leftrightarrow$}&
\multirow{-2}{*}{INTUITIVE} &\multirow{-2}{*}{INTER}  \\

\rowcolor{lightgray} 
1&2&$\downarrow$&$\downarrow\downarrow$&$\downarrow\downarrow$&$\downarrow$&$\uparrow$&$\uparrow\uparrow$& && \\ \rowcolor{lightgray} 
2&2&$\downarrow$&$\downarrow$&$\leftrightarrow$&$\downarrow$&$\uparrow$&$\uparrow\uparrow$ & 
\multirow{-2}{*}{$\uparrow$} & \multirow{-2}{*}{INTUITIVE}&  \multirow{-2}{*}{INTUITIVE} \\

1&3&$\downarrow$&$\downarrow\downarrow$&$\downarrow\downarrow$&$\leftrightarrow$&$\uparrow$&$\leftrightarrow$ &&& \\
2&3&$\downarrow$&$\downarrow$&$\downarrow$&$\downarrow\downarrow$&$\uparrow$&$\leftrightarrow$ &
\multirow{-2}{*}{$\uparrow$} &\multirow{-2}{*}{INTUITIVE}& \multirow{-2}{*}{INTUITIVE}\\

\rowcolor{lightgray} 
1&4&$\uparrow$&$\uparrow$&$\uparrow$&$\uparrow\uparrow$&$\downarrow$&$\downarrow\downarrow$ &&& \\ 
\rowcolor{lightgray} 
2&4&$\uparrow$&$\uparrow$&$\uparrow$&$\uparrow\uparrow$&$\downarrow$&$\downarrow\downarrow$ &\multirow{-2}{*}{$\downarrow$}&
\multirow{-2}{*}{DEL}& \multirow{-2}{*}{DEL} \\

1&5&$\uparrow$&$\uparrow$&$\uparrow$&$\uparrow$&$\downarrow$&$\downarrow$ &&& \\
2&5&$\uparrow$&$\uparrow$&$\uparrow\uparrow$&$\uparrow\uparrow$&$\downarrow$&$\downarrow\downarrow$ &\multirow{-2}{*}{$\downarrow$}&
\multirow{-2}{*}{DEL}& \multirow{-2}{*}{DEL}\\

\rowcolor{lightgray} 
1&6&$\leftrightarrow$&$\leftrightarrow$&$\leftrightarrow$&$\downarrow$&$\leftrightarrow$&$\uparrow\uparrow$ &&&\\  \rowcolor{lightgray} 
2&6&$\leftrightarrow$&$\leftrightarrow$&$\downarrow$&$\uparrow$&$\uparrow\uparrow$&$\uparrow$ &\multirow{-2}{*}{$\uparrow$}&\multirow{-2}{*}{INTER}&\multirow{-2}{*}{INTER}\\

1&7&$\uparrow$&$\uparrow$&$\uparrow$&$\uparrow$&$\downarrow$&$\downarrow$ &&& \\
2&7&$\uparrow$&$\uparrow$&$\uparrow$&$\leftrightarrow$&$\downarrow$&$\downarrow\downarrow$ &\multirow{-2}{*}{$\downarrow$}&\multirow{-2}{*}{DEL} &\multirow{-2}{*}{DEL}\\

\rowcolor{lightgray} 
1&8&$\uparrow$&$\uparrow$&$\uparrow$&$\uparrow$&$\uparrow$&$\downarrow$ &&& \\
\rowcolor{lightgray} 
2&8&$\uparrow$&$\uparrow$&$\uparrow$&$\uparrow\uparrow$&$\uparrow$&$\downarrow\downarrow$ &\multirow{-2}{*}{$\downarrow$}&\multirow{-2}{*}{DEL}&\multirow{-2}{*}{DEL}\\

1&9&$\leftrightarrow$&$\leftrightarrow$&$\leftrightarrow$&$\leftrightarrow$&$\uparrow$&$\uparrow$ &&& \\
2&9&$\leftrightarrow$&$\uparrow$&$\uparrow$&$\leftrightarrow$&$\downarrow$&$\downarrow\downarrow$ &\multirow{-2}{*}{$\uparrow$} &\multirow{-2}{*}{INTER}& \multirow{-2}{*}{INTER}\\

\rowcolor{lightgray} 
1&10&$\uparrow$&$\uparrow$&$\uparrow$&$\uparrow$&$\downarrow$&$\downarrow\downarrow$ &&&\\
\rowcolor{lightgray} 
2&10&$\uparrow$&$\uparrow$&$\uparrow\uparrow$&$\uparrow$&$\downarrow$&$\downarrow\downarrow$ & \multirow{-2}{*}{$\downarrow$}&\multirow{-2}{*}{DEL}&\multirow{-2}{*}{DEL}\\

1&11&$\downarrow$&$\leftrightarrow$&$\uparrow$&$\uparrow$&$\uparrow$&$\uparrow$ &&&\\
2&11&$\downarrow$&$\downarrow$&$\downarrow$&$\downarrow$&$\uparrow$&$\uparrow$ &\multirow{-2}{*}{$\downarrow$}&\multirow{-2}{*}{INTER}&\multirow{-2}{*}{INTUITIVE}\\

\rowcolor{lightgray} 
1&12&$\leftrightarrow$&$\leftrightarrow$&$\leftrightarrow$&$\leftrightarrow$&$\downarrow$&$\leftrightarrow$ &&&\\
\rowcolor{lightgray} 
2&12&$\leftrightarrow$&$\leftrightarrow$&$\leftrightarrow$&$\downarrow$&$\uparrow$&$\uparrow$ &\multirow{-2}{*}{$\leftrightarrow$}&\multirow{-2}{*}{INTER}&\multirow{-2}{*}{INTER} \\

1&13&$\leftrightarrow$&$\downarrow$&$\leftrightarrow$&$\uparrow$&$\downarrow$&$\uparrow$ &&&\\
2&13&$\uparrow$&$\downarrow$&$\leftrightarrow$&$\downarrow$&$\uparrow$&$\downarrow$& \multirow{-2}{*}{$\uparrow$}&\multirow{-2}{*}{INTER}&\multirow{-2}{*}{INTER} \\

\rowcolor{lightgray} 
1&14&$\downarrow$&$\downarrow$&$\uparrow$&$\downarrow$&$\uparrow\uparrow$&$\uparrow\uparrow$ &&&\\
\rowcolor{lightgray} 
2&14&$\uparrow$&$\leftrightarrow$&$\uparrow\uparrow$&$\downarrow$&$\uparrow$&$\uparrow$ & \multirow{-2}{*}{$\downarrow$}&\multirow{-2}{*}{INTER}&\multirow{-2}{*}{INTER} \\

1&15&$\downarrow$&$\downarrow$&$\downarrow$&$\downarrow\downarrow$&$\downarrow$&$\uparrow$ &&&\\
2&15&$\downarrow$&$\downarrow$&$\downarrow$&$\downarrow$&$\uparrow\uparrow$&$\uparrow\uparrow$ &\multirow{-2}{*}{$\uparrow\uparrow$}&\multirow{-2}{*}{INTUITIVE}&\multirow{-2}{*}{INTUITIVE}\\

\rowcolor{lightgray} 
1&16&$\uparrow$&$\downarrow$&$\leftrightarrow$&$\downarrow$&$\uparrow$&$\downarrow$ &&& \\
\rowcolor{lightgray} 
2&16&$\downarrow$&$\downarrow$&$\leftrightarrow$&$\uparrow$&$\uparrow$&$\downarrow$&\multirow{-2}{*}{$\uparrow$}&\multirow{-2}{*}{INTER}&\multirow{-2}{*}{INTER}\\

1&17&$\downarrow$&$\downarrow$&$\downarrow$&$\downarrow$&$\uparrow\uparrow$&$\uparrow\uparrow$ &&&\\
2&17&$\leftrightarrow$&$\downarrow$&$\downarrow$&$\uparrow\uparrow$&$\uparrow$&$\uparrow\uparrow$ &\multirow{-2}{*}{$\uparrow$} &
\multirow{-2}{*}{INTUITIVE}& \multirow{-2}{*}{INTUITIVE}\\

\rowcolor{lightgray} 
1&18&$\uparrow$&$\uparrow$&$\uparrow$&$\uparrow$&$\downarrow$&$\downarrow\downarrow$ &&&\\
\rowcolor{lightgray} 
2&18&$\uparrow$&$\uparrow$&$\uparrow\uparrow$&$\downarrow$&$\downarrow$&$\downarrow\downarrow$ &\multirow{-2}{*}{$\downarrow$}& \multirow{-2}{*}{DEL} & \multirow{-2}{*}{DEL}\\

1&19&$\leftrightarrow$&$\downarrow$&$\uparrow$&$\downarrow$&$\downarrow$&$\downarrow$ &&&\\
2&19&$\leftrightarrow$&$\leftrightarrow$&$\downarrow$&$\downarrow$&$\downarrow$&$\downarrow$ &\multirow{-2}{*}{$\uparrow$}& \multirow{-2}{*}{INTUITIVE}& \multirow{-2}{*}{INTER}\\ \bottomrule
\end{tabular}
\caption{Overview of each participants' performance with respect to the specified criteria in week 1 and week 2.
$\uparrow$, $\downarrow$, and $\leftrightarrow$ denote above, below, and average, respectively.
\label{tab:individual}}
\end{table}

The non-aggregated data provided by~\cite{khemlani:2016} included a dataset of 20 participants,
for which the task was carried twice, with a break of one week. On average, the basic principles explained the results on about 10\% better than CL.
 Generally, the participants who were strong in CL/ contraposition,drew fewer conclusions according to abduction/ generalization, and vice versa.
 The changes within participant's answers from one to the other week was high, with 52/64 on average. 
 From week 1 to week 2, an increase with respect to CL (19/20, 6\%), basic (18/20, 6\%), contraposition (15/20, 5\%)
 and a decrease on generalization (15/20, 8\%) and no change on abduction (10/20, 0\%) answers could be observed.

Table~\ref{tab:individual} provides an overview of the performance of of each participant (column 2), for week 1 and week 2 (column 1)
with respect to CL (column 4) and the cognitive principles (column 5, 6, 7, and 8).
NVC (column 3) denotes how frequently participants responded \textit{no valid conclusion} and Changes (column 9) denotes how
frequently participants changed their answers from week 1 to week 2. 
The symbols $\uparrow$, $\downarrow$, and $\leftrightarrow$ denote above, below, and average, respectively.

Similar to~\cite{khemlani:2016}, we identified three different clusters: 
The last two columns show the classification of the participants as deliberative, intuitive or intermediate. The participants in the deliberative cluster here corresponds exactly to~\cite{khemlani:2016}. 
Our classification only differs to~\cite{khemlani:2016} for the intermediate and intuitive cluster with respect to three participants (1, 11 and 19).
A natural explanation for this difference, could be that~\cite{khemlani:2016}
did not take into account the participants' changes between week 1 and week 2.

Two of the three clusters can be specified through a set of cognitive principles, and for one, no distinguishable pattern could be identified:
% The summary is shown in Table~\ref{tab:clusters}.
\begin{description}
 \item[Deliberative cluster] (6/20) is characterized by the participants whose answers corresponded to the predictions made by basic/ contraposition. Almost none of their conclusions corresponded to predictions made by abduction/ generalization. Theses participants did change their answers from week 1 to week 2 
 lower than average and concluded higher than average NVC. Generally, their answers improved from week 1 to week 2 with respect to CL.
 \item[Intuitive cluster] (7/20)
is characterized by the participants whose answers corresponded to the predictions given by abduction/ generalization. Almost none of their conclusions corresponded to the predictions made by contraposition (except of 2 participants in week 2). They showed a high change within their own answers from week 1 to week 2 
 and concluded lower than average NVC. Generally, their answers improved from week 1 to week 2 with respect to CL, but they never performed \textit{as well as} the participants in 
 the deliberative cluster.
 \item[Intermediate cluster] (7/20) is characterized by the participants whose answers did not corresponded to any particular principle. These participants had a high change within their own answers from week 1 to week 2
 and no particular pattern with respect to NVC conclusions. Generally they were not improving their answers from week 1 to week 2 with respect to CL, but normally compensated for 
 right answers in some cases (that corresponded to contraposition) with wrong answers in the other cases (that correspond to abduction/ generalization).
\end{description}

 \section{Conclusions and Future Directions}

We have investigated the cognitive adequacy of WCS clustering with respect to the individual answer patterns reported by~\cite{khemlani:2016}.
\cite{khemlani:2016} is the only approach that we are aware of, that has evaluated their predictions on human syllogistic reasoning
with respect to the non-aggregated data.
However different than here, they did not distinguish between the participants' responses in week 1 and week 2.

Our results are two-fold: First, we were able to show that the reasoning clusters as specified in~\cite{declare:2017}, could not be validated
with respect to these individual reasoning patterns. However, and this leads us to the second more interesting result,
we could show that participants' answers seemed to be explainable through a set of cognitive principles. In particular,
some principles seemed to exclude each other by the individual reasoners throughout the experiment (contraposition and abduction), whereas others often appeared together (basic/ contraposition and abduction/ generalization). 
The individual participants' data allowed us to observe other reasoning patterns (high/ low frequency of NVC conclusions and high/ low frequency of changing answers between week 1 and week 2 answers). 
Interestingly, there was one group of participants for which we could not identify any pattern (intermediate). 
They seemed to compensate for their wrong/ right answers, and might have changed their strategies within the task.
Another explanation might be that the set of cognitive principles presented in Section~\ref{sect:representing} is not complete yet.

% Static clusters for each participant might not be appropriate, rather, a sequential application of cognitive principles seems to be more likely
For the future, we need to investigate whether the sequential application of the cognitive principles 
can give us more insights on the participants' reasoning pattern. In particular, our hypothesis is that these principles might correspond to some of the reasoning steps within the insight model~\cite{johnsonlaird:1970} for the selection task~\cite{wason:68}. Similarly, a meta-analysis of the individual response patterns for the selection task has been suggested in~\cite{ragni:2017}.
This in turn might give us a new starting point to investigate whether the cognitive principles  specified here can characterize participants' conclusions in other reasoning tasks.
% episodes.
% In order to validate the idea of a sequential reasoning procedure, cognitive scientists will have to set-up new 
% experiments, where possibly, participants are explicitly asked (not) to assume certain cognitive principles.
% Another aspect that needs to be examined, is

\section{Acknowledgments}

Thanks to the reviewer for pointing us to the relationship of the $\cxt$ operator
with {\L}ukasiewicz' extensional necessity operator.

\bibliographystyle{splncs}
\bibliography{bibv4}

\end{document}

\begin{table}[htp]\tiny\centering
\newcolumntype{a}{>{\columncolor{white}}c@{\hspace{2mm}}}
\newcolumntype{b}{>{\columncolor{gray!40}}c@{\hspace{2mm}}}  
\captionsetup{font=footnotesize}
 \begin{tabular}{llllllllll}
 \toprule
	 & \MA ac & \ME ac & \MI ac & \MO ac & \MA ca & \ME ca & \MI ca & \MO ca & \NVC \\ \midrule

\rowcolor{lightgray} 
\MA\MA1  & \basicc&        &        &        &        &        &        &        &      \\ \rowcolor{lightgray} 
\MA\MA2  &        &        &        &        & \basicc&        &        &        & \\ \rowcolor{lightgray} 
\MA\MA3  &        &        &        &        &        &        &        &        & \basicc \\ \rowcolor{lightgray} 
\MA\MA4  & \basicc&        &        &        &\basicc &        &        &        & \\ %\midrule

\MA\MI1  &        &        & \abdc  &        &        &        & \abdc  &        & \basicc \\
\MA\MI2  &        &        & \basicc&        &        &        & \basicc \\
\MA\MI3  &        &        & \abdc  &        &        &        & \abdc  &        & \basicc \\
\MA\MI4  &        &        & \basicc&        &        &        & \basicc&        & \\%\midrule

\rowcolor{lightgray} 
\MA\ME1  &        & \basicc&        &        &        &        &        &        & \\ \rowcolor{lightgray} 
\MA\ME2  &        &\contrac&        &        &        & \basicc&        &        & \\ \rowcolor{lightgray} 
\MA\ME3  &        &\contrac&        &        &        &\contrac&        &        & \basicc \\ \rowcolor{lightgray} 
\MA\ME4  &        & \basicc&        &        &        &\contrac&        &        &  \\%\midrule

\MA\MO1  &        &        &        &\abdc   &        &        &        &        & \basicc \\
\MA\MO2  &        &        &        &        &        &        &        &\basicc \\
\MA\MO3  &        &        &        &        &        &        &\contrac&\basicc \\
\MA\MO4  &        &        &        &\basicc \\%\midrule

\rowcolor{lightgray} 
\MI\MA1  &        &        &\basicc &        &        &        &\basicc &        & \\\rowcolor{lightgray} 
\MI\MA2  &        &        &\abdc   &        &\genc   &        & \abdc  &        & \basicc \\ \rowcolor{lightgray} 
\MI\MA3  &        &        & \abdc  &        &        &        & \abdc  &        & \basicc \\ \rowcolor{lightgray} 
\MI\MA4  &        &        &\basicc &        &        &        &\basicc &        &  \\  %\midrule

\MI\MI1  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\
\MI\MI2  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\
\MI\MI3  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\
\MI\MI4  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\ %\midrule

\rowcolor{lightgray} 
\MI\ME1  &        &        &        &\basicc &        &        &        &        & \\\rowcolor{lightgray} 
\MI\ME2  &        &        &        &\contrac&        & \genc  &        &        & \basicc \\\rowcolor{lightgray} 
\MI\ME3  &        &        &        &\contrac&        & \genc  &        &        & \basicc  \\\rowcolor{lightgray} 
\MI\ME4  &        &        &        & \basicc&        &        &        &        & \\ %\midrule

\MI\MO1  &        &        &        & \genc  &        &        &        &        & \basicc \\
\MI\MO2  &        &        &        & \genc  &        &        &        &        & \basicc \\ 
\MI\MO3  &        &        &        & \genc  &        &        &        &        & \basicc \\ 
\MI\MO4  &        &        &        & \genc  &        &        &        &        & \basicc \\%\midrule

\rowcolor{lightgray} 
\ME\MA1  &        & \basicc&        &        &        &        &        &        & \\\rowcolor{lightgray} 
\ME\MA2  &        &\contrac&        &        &        & \basicc&        &        &  \\\rowcolor{lightgray} 
\ME\MA3  &        &\contrac&        &        &        &\contrac&        &        & \basicc \\\rowcolor{lightgray} 
\ME\MA4  &        &\contrac&        &        &        & \basicc&        &        &\\  %\midrule

\ME\MI1  &        & \genc   &        &	 &	  &	   &	    &	     & \basicc \\
\ME\MI2  &        &	       &	&	 &	  &	   &	    & \basicc \\
\ME\MI3  &        & \genc   &        &	 &	  &	   &	    &\contrac& \basicc \\
\ME\MI4  &        &	       &	&	 &	  &	   &	    & \basicc \\%\midrule

\rowcolor{lightgray} 
\ME\ME1  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ \rowcolor{lightgray} 
\ME\ME2  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ \rowcolor{lightgray} 
\ME\ME3  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ \rowcolor{lightgray} 
\ME\ME4  &        &	       &	&	 &	  &	   &	    &        &\basicc \\%\midrule

\ME\MO1  &        &	       &	&	 &	  &	   &	    &        &\basicc \\
\ME\MO2  &        &	       &	&	 &	  &	   &	    &        &\basicc \\
\ME\MO3  &        &	       &	&	 &	  &	   &	    &        &\basicc \\
\ME\MO4  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ %\midrule

\rowcolor{lightgray} 
\MO\MA1  &        &            &	& \basicc&        &        &        &        & \\\rowcolor{lightgray} 
\MO\MA2  &        &	       &	&	 &	  &	   &	    &  \abdc &\basicc \\\rowcolor{lightgray} 
\MO\MA3  &        &	       &	&\contrac&	  &	   &	    &        &\basicc \\\rowcolor{lightgray} 
\MO\MA4  &        &	       &	&	 &	  &	   &	    & \basicc& \\  %\midrule

\MO\MI1  &        &	       &	&\genc   &	  &	   &	    &        &\basicc \\
\MO\MI2  &        &	       &	&	 &	  &	   &	    &\genc   &\basicc \\
\MO\MI3  &        &	       &	&\genc   &	  &	   &	    &        &\basicc \\
\MO\MI4  &        &	       &	&	 &	  &	   &	    &\genc   &\basicc\\ %\midrule

\rowcolor{lightgray} 
\MO\ME1  &        &	       &	&        &	  &	   &	    &        &\basicc \\\rowcolor{lightgray} 
\MO\ME2  &        &	       &	&        &	  &	   &	    &        &\basicc \\\rowcolor{lightgray} 
\MO\ME3  &        &	       &	&        &	  &	   &	    &        &\basicc \\\rowcolor{lightgray} 
\MO\ME4  &        &	       &	&        &	  &	   &	    &        &\basicc \\ %\midrule

\MO\MO1  &        &	       &	&        &	  &	   &	    &        &\basicc \\
\MO\MO2  &        &	       &	&        &	  &	   &	    &        &\basicc \\
\MO\MO3  &        &	       &	&        &	  &	   &	    &        &\basicc \\
\MO\MO4  &        &	       &	&        &	  &	   &	    &        &\basicc \\ \bottomrule
  \end{tabular}

\caption{Predictions according to the cognitive principles as specified in Section~\ref{sect:representing}.\betweentinyandsmall}
 \end{table}

% \commentem{Dieser Table ist falsch!!!}
% \begin{table}
% \begin{center}
% \begin{adjustbox}{totalheight={20.5cm}} 
% 
% \begin{tabular}{c@{\hspace{3cm}}c}
% 
%   \begin{tabular}{r|c|c|c|c|c|c|c|c|c|} 
%  Premises & Aac & Iac & Eac & Oac & Aca & Ica & Eca & Oca & NVC \\ 
%  \cline{2-10} 
%  AA1 &\cellcolor{black} & & & & & & & & \\ 
%  \cline{2-10} 
%  AA2 &&&&&\cellcolor{black} & &&&\\ 
%  \cline{2-10} 
%  AA3 &&\cellcolor{gray!50} & &&\cellcolor{gray!50} & &&&\cellcolor{gray!100}  \\ 
%  \cline{2-10} 
%  AA4 &\cellcolor{gray!50} & \cellcolor{gray!50} & &&&&&&\cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  AE1 &&&\cellcolor{black} & &&&&&\\ 
%  \cline{2-10} 
%  AE2 &&&&&&&\cellcolor{gray!100} & &\cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  AE3 &&&\cellcolor{gray!50} & &&&\cellcolor{gray!100} & &\\ 
%  \cline{2-10} 
%  AE4 &&&\cellcolor{gray!50} & \cellcolor{gray!50} & &&&&\cellcolor{gray!50} \\ 
%  \cline{2-10} 
%  AI1 &&\cellcolor{black} & &&&&&&\cellcolor{gray!50} \\ 
%  \cline{2-10} 
%  AI2 &&&&&&\cellcolor{black} & &&\\ 
%  \cline{2-10} 
%  AI3 &&&&&&\cellcolor{gray!50} & &&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  AI4 &&\cellcolor{gray!100} & &&&\cellcolor{gray!100} & &&\\ 
%  \cline{2-10} 
%  AO1 &&&&\cellcolor{black} & &&&&\cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  AO2 &&&&&&&&\cellcolor{gray!100} & \cellcolor{gray!100} \\ 
%  \cline{2-10} 
%  AO3 &&&&&&&&\cellcolor{gray!100} & \cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  AO4 &&&&\cellcolor{black} & &&&&\\ 
%  \cline{2-10} 
%  EA1 &&&\cellcolor{gray!100} & &&&\cellcolor{gray!50} & &\cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  EA2 &&&&&&&\cellcolor{black} & &\\ 
%  \cline{2-10} 
%  EA3 &&&\cellcolor{gray!100} & &&&\cellcolor{gray!100} & &\\ 
%  \cline{2-10} 
%  EA4 &&&\cellcolor{gray!50} & &&&\cellcolor{gray!50} & \cellcolor{gray!50} & \cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  EE1 &&&\cellcolor{gray!50} & &&&&&\cellcolor{gray!100}  \\ 
%  \cline{2-10} 
%  EE2 &&&&&&&\cellcolor{gray!50} & &\cellcolor{black}  \\ 
%  \cline{2-10} 
%  EE3 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  EE4 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  EI1 &&&\cellcolor{gray!50} & &&&&\cellcolor{gray!50} & \cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  EI2 &&&&&&&&\cellcolor{black} & \cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  EI3 &&&&&&&&\cellcolor{gray!100} & \cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  EI4 &&&&&&&&\cellcolor{gray!100} & \cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  EO1 &&&&\cellcolor{gray!50} & &&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  EO2 &&&&&&&&\cellcolor{gray!50} & \cellcolor{black} \\ 
%  \cline{2-10} 
%  EO3 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  EO4 &&&&\cellcolor{gray!50} & &&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  IA1 &&\cellcolor{black} & &&&&&&\\ 
%  \cline{2-10} 
%  IA2 &&&&&&\cellcolor{black} & &&\cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  IA3 &&\cellcolor{gray!50} & &&&\cellcolor{gray!50} & &&\cellcolor{gray!100}  \\ 
%  \cline{2-10} 
%  IA4 &&\cellcolor{gray!100} & &&&\cellcolor{gray!100} & &&\\ 
%  \cline{2-10} 
%  IE1 &&&&\cellcolor{black} & &&&&\\ 
%  \cline{2-10} 
%  IE2 &&&&\cellcolor{gray!50} & &&\cellcolor{gray!50} & &\cellcolor{gray!100}  \\ 
%  \cline{2-10} 
%  IE3 &&&\cellcolor{gray!50} & \cellcolor{gray!50} & &&\cellcolor{gray!50} & &\cellcolor{gray!50} \\ 
%  \cline{2-10} 
%  IE4 &&&&\cellcolor{gray!100} & &&&&\cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  II1 &&\cellcolor{gray!50} & &&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  II2 &&&&&&\cellcolor{gray!50} & &&\cellcolor{black} \\ 
%  \cline{2-10} 
%  II3 &&\cellcolor{gray!50} & &&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  II4 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  IO1 &&&&\cellcolor{gray!50} & &&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  IO2 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  IO3 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  IO4 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  OA1 &&&&\cellcolor{gray!50} & &&&&\cellcolor{gray!100}  \\ 
%  \cline{2-10} 
%  OA2 &&&&&&&&\cellcolor{black} & \cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  OA3 &&&&\cellcolor{gray!50} & &&&\cellcolor{gray!50} & \cellcolor{gray!100}  \\ 
%  \cline{2-10} 
%  OA4 &&&&&&&&\cellcolor{black} & \cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  OE1 &&&&\cellcolor{gray!50} & &&&&\cellcolor{black} \\ 
%  \cline{2-10} 
%  OE2 &&&&&&&&\cellcolor{gray!50} & \cellcolor{black} \\ 
%  \cline{2-10} 
%  OE3 &&&&&&&&&\cellcolor{black} \\ 
%  \cline{2-10} 
%  OE4 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  OI1 &&&&\cellcolor{gray!50} & &&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  OI2 &&&&&&&&\cellcolor{gray!50} & \cellcolor{black}  \\ 
%  \cline{2-10} 
%  OI3 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  OI4 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  OO1 &&&&&&&&&\cellcolor{black} \\ 
%  \cline{2-10} 
%  OO2 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  OO3 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  OO4 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%   \end{tabular}
%   & 
%  
%   \begin{tabular}{r|c|c|c|c|c|c|c|c|c|} 
%  Premises & Aac & Iac & Eac & Oac & Aca & Ica & Eca & Oca & NVC \\ 
%  \cline{2-10} 
%  AA1 &\cellcolor{gray!100} & &&&&&&&\cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  AA2 &&&&&\cellcolor{gray!100} & &&&\cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  AA3 &&&&&&&&&\cellcolor{black} \\ 
%  \cline{2-10} 
%  AA4 &\cellcolor{gray!50} & &&&\cellcolor{gray!50} & &&&\cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  AE1 &&&\cellcolor{gray!50} & &&&\cellcolor{gray!50} & &\cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  AE2 &&&\cellcolor{gray!50} & &&&\cellcolor{gray!50} & &\cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  AE3 &&&\cellcolor{gray!50} & &&&\cellcolor{gray!50} & &\cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  AE4 &&&\cellcolor{gray!50} & &&&\cellcolor{gray!50} & &\cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  AI1 &&\cellcolor{gray!100} & &&&&&&\cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  AI2 &&&&&&\cellcolor{gray!100} & &&\cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  AI3 &&&&&&&&&\cellcolor{black} \\ 
%  \cline{2-10} 
%  AI4 &&\cellcolor{gray!100} & &&&&&&\cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  AO1 &&&&\cellcolor{gray!100} & &&&&\cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  AO2 &&&&&&&&\cellcolor{black} & \cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  AO3 &&&&&&&&\cellcolor{gray!50} & \cellcolor{black}  \\ 
%  \cline{2-10} 
%  AO4 &&&&\cellcolor{black} & &&&&\cellcolor{gray!50} \\ 
%  \cline{2-10} 
%  EA1 &&&\cellcolor{gray!50} & &&&\cellcolor{gray!50} & &\cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  EA2 &&&\cellcolor{gray!50} & &&&\cellcolor{gray!50} & &\cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  EA3 &&&\cellcolor{gray!50} & &&&\cellcolor{gray!50} & &\cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  EA4 &&&\cellcolor{gray!50} & &&&\cellcolor{gray!50} & &\cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  EE1 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  EE2 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  EE3 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  EE4 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  EI1 &&&&\cellcolor{black} & &&&&\cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  EI2 &&&&\cellcolor{gray!50} & &&&&\cellcolor{gray!100}  \\ 
%  \cline{2-10} 
%  EI3 &&&&\cellcolor{gray!50} & &&&&\cellcolor{gray!100}  \\ 
%  \cline{2-10} 
%  EI4 &&&&\cellcolor{black} & &&&&\cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  EO1 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  EO2 &&&&&&&&&\cellcolor{black} \\ 
%  \cline{2-10} 
%  EO3 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  EO4 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  IA1 &&\cellcolor{gray!100} & &&&&&&\cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  IA2 &&&&&&\cellcolor{gray!100} & &&\cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  IA3 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  IA4 &&&&&&\cellcolor{gray!100} & &&\cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  IE1 &&&&&&&&&\cellcolor{black} \\ 
%  \cline{2-10} 
%  IE2 &&&&&&&&\cellcolor{black} & \cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  IE3 &&&&&&&&\cellcolor{gray!50} & \cellcolor{gray!100}  \\ 
%  \cline{2-10} 
%  IE4 &&&&&&&&\cellcolor{black} & \cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  II1 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  II2 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  II3 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  II4 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  IO1 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  IO2 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  IO3 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  IO4 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  OA1 &&&&\cellcolor{black} & &&&&\cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  OA2 &&&&&&&&\cellcolor{gray!100} & \cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  OA3 &&&&\cellcolor{gray!50} & &&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  OA4 &&&&&&&&\cellcolor{black} & \cellcolor{gray!50}  \\ 
%  \cline{2-10} 
%  OE1 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  OE2 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  OE3 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  OE4 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  OI1 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  OI2 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  OI3 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  OI4 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  OO1 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  OO2 &&&&&&&&&\cellcolor{black} \\ 
%  \cline{2-10} 
%  OO3 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%  OO4 &&&&&&&&&\cellcolor{black}  \\ 
%  \cline{2-10} 
%   \end{tabular} 
% \end{tabular}
% \end{adjustbox} 
% \end{center}
% \caption{The data reported in~\cite{khemlani:2012} (left) 
% and the predictions made by WCS Clustering (right):
% White means $< 16\%$, {\color{gray!50}light gray means $< 40\%$}, {\color{gray!100}gray means $< 60\%$}
% and {black means $>60\%$}.
%  \label{tab:eval:aggregated}}
% \end{table}
% Table~\ref{tab:eval:aggregated} gives an overview of the participants conclusions reported in~\cite{khemlani:2012}
% and the predictions under the WCS clustering.

### 2019-11-06 paper/paper-v3.tex
small changes table Oca
changed a bit after first reading

@@
 
-Originally, one of the objectives of using (not necessarily classical) logic in the field of AI and logic programming was to formalize human reasoning and commonsense reasoning~\cite{McCarthy59,McCarthy98}.
-However, the proposed approaches did not have the expected successes in this~(cf.~\cite{mcdermott:87}),
-as proposed approaches were mostly theoretical and did not apply to real case studies.
-On the other hand, cognitive scientists have extensively investigated human thinking and acquired a great expertise.
-However, this expertise was not consulted, i.e. there was no communication among the relevant fields.
+Even though the main logical formalism of Classical Logic (CL) has successfully been applied to our scientific reasoning,
+various experimental studies have shown that human reasoning systematically deviates 
+from conclusions valid under CL (cf.~\cite{wason:68,byrne:89}).
+Originally, one of the objectives of using (non-classical) logic in the field of AI and logic programming was to formalize human reasoning and commonsense reasoning~\cite{McCarthy59,McCarthy98}.
+% However, the proposed approaches did not have the expected successes~(cf.~\cite{mcdermott:87}),
+% as they were mostly theoretical and did not apply to real case studies.
+On the other hand, cognitive scientists have extensively investigated and acquired a great expertise human thinking.
+However, only recently an attempt was made to bridge the gap among these disciplines.
+% this expertise was not consulted, i.e. there was no communication among the relevant fields.
 
@@
 
-Yet, the cognitive adequacy of a theory can only be measured, if that theory is evaluable.
-In this context, that means, if the theory can uniformally account for episodes of human reasoning systematically, according to its underlying assumptions.
-Only then, the cognitve adequacy of that theory can be compared to others.
+Yet, the cognitive adequacy of a theory can only be measured, if that theory is evaluable, i.e.\ if the theory can uniformally account for episodes of human reasoning systematically, according to its underlying assumptions.
+% Only then, the cognitve adequacy of that theory can be compared to others.
 
 Formal methods can offer the required tools to implement such cognitive theories.
-This in turn, can allow a new way of communication among the fields, by avoiding ambiguous understanding, showing their formal correspondences and finally providing an appropriate evaluation method with respect to their cognitive adequacy.
+This in turn, can allow a new way of communication among the fields, by avoiding ambiguous definitions, showing their formal correspondences and finally providing appropriate evaluation methods with respect to their cognitive adequacy.
 In order to determine the required factors that are necessary for the evaluation of an adequate cognitive theory, we take~\cite{ragni:stolzenburg:2015} as a starting point:
-We aim for a theory in which (i) all processes are comprehensible, (ii) the suggested principles represent general cognitive processes, and (iii) the predictions of our theory are evaluated with respect to a wide range of human reasoning tasks.
-% Even though the main logical formalism of Classical Logic (CL) has successfully been applied to our scientific reasoning,
-% various experimental studies have shown that human reasoning systematically deviates 
-% from classical logical conclusions (cf.~\cite{wason:68,byrne:89}).
+We aim for a theory in which (i) all processes are comprehensible, (ii) the suggested underlying represent general cognitive processes, and (iii) the predictions of our theory are evaluated with respect to a wide range of human reasoning tasks.
 
@@
 \begin{array}{@{\hspace{0.0cm}}l@{\hspace{0.1cm}}G@{\hspace{0.1cm}}l@{\hspace{0.3cm}}c}\toprule
-	\mbox{mood} & \mbox{natural language} & \mbox{FOL} &
+	\mbox{Mood} & \mbox{natural language} & \mbox{First-order Logic} &
 \mbox{Short} \\\midrule
@@
 		\toprule
-		& \mbox{premise 1} & \mbox{premise 2} \\
+		 \mbox{Figure} & \mbox{Premise 1} & \mbox{Premise 2} \\
 		\midrule
-		\mbox{figure 1} &  a-b & b-c \\
-		\mbox{figure 2} & b-a &  c-b \\
-		\mbox{figure 3} & a-b & c-b \\
-		\mbox{figure 4} & b-a & b-c \\
+		\mbox{1} &  a-b & b-c \\
+		\mbox{2} & b-a &  c-b \\
+		\mbox{3} & a-b & c-b \\
+		\mbox{4} & b-a & b-c \\
 		\bottomrule
@@
 According to~\cite{khemlani:2012}, the majority of participants in experimental studies
-concluded \emph{No chemists are artists} (29\%) and \emph{No valid conclusion} (27\%). The classical logical valid conclusion \emph{Some artists are not chemists}, was concluded 
-by only 16\%, which is exactly the threshold, that the answer had not been chosen randomly.\footnote{
- Given that there are nine different possible conclusions, the chance that a conclusion has been chosen randomly is 1/9 = 11.1 \%. 
- A binomial test shows that if a conclusion is drawn in
-more than 16\% of the cases by the participants it is unlikely that is has been
-chosen by just random guesses~\cite{khemlani:2012}. }
+concluded \emph{No chemists are artists} (29\%) and \emph{No valid conclusion} (27\%). The valid conclusion \emph{Some artists are not chemists}
+under CL, was concluded by only 16\%, which is exactly the threshold, that the answer had not been chosen randomly~\cite{khemlani:2012}.
+% \footnote{
+% See
+%  Given that there are nine different possible conclusions, the chance that a conclusion has been chosen randomly is 1/9 = 11.1 \%. 
+%  A binomial test shows that if a conclusion is drawn in
+% more than 16\% of the cases by the participants it is unlikely that is has been
+% chosen by just random guesses~\cite{khemlani:2012}. }
 % Recently, a new cognitive theory has been proposed for syllogistic reasoning~\cite{cogsci:2017}.
@@
 % The majority's conclusions exclude each other, as the test was made such that
-The ones who answered \emph{No valid conclusion}
-were not the same ones who answered \emph{No chemists are artists} or \emph{Some artists are not chemists}, and vice versa.\footnote{Note that usually in these experiments, participants were supposed to give only one answer}
+Usually in these experiments, participants are supposed to give only one answer, thus all three (significant) answers
+were given by different participants. 
+% The ones who answered \emph{No valid conclusion}
+% were not the same ones who answered \emph{No chemists are artists} or \emph{Some artists are not chemists}, and vice versa.\footnote{}
 
@@
 
-As the above example and recent investigations on psychological experiments show (cf.~\cite{khemlani:2016, ragni:2017}), it seems that different humans represent and reason about conditional information differently.
+As the above example and recent investigations on psychological experiments show (cf.~\cite{khemlani:2016, ragni:2017}), it seems that different humans represent and reason about conditionals differently.
 This implies that \textit{the} human reasoner does not exist, instead there 
@@
 in the sense that we investigate the individual participants' conclusions with respect to the non-aggregated dataset provided by~\cite{khemlani:2016}.
-In Section~\ref{sect:wcsclustering} and~\ref{sect:representing}, we introduce the underlying computational logic approach and the cognitive principles.
+In Section~\ref{sect:wcsclustering} and~\ref{sect:representing}, we introduce the underlying computational logic approach and the formal representation of cognitive principles within this approach.
 Section~\ref{sect:aggregateddata} illustrates an example and briefly discusses the performance of the approach in~\cite{declare:2017} on non-aggregated data.
-Section~\ref{sect:nonag} provides a new characterization of the individual's reasoning patterns that 
-seems to be more adequate.
-
-\section{WCS Clustering} \label{sect:wcsclustering}
-
-The authors in~\cite{declare:2017} put forward the hypothesis that reasoning clusters can be expressed through cognitive principles and heuristic strategies.
-Cognitive principles are (not necessarily valid) assumptions made by humans while reasoning,
+Section~\ref{sect:nonag} provides a more adequate characterization of the individual's reasoning patterns.
+
+\section{Clustering by Principles under WCS} \label{sect:wcsclustering}
+
+According to~\cite{declare:2017}, reasoning clusters can be expressed through cognitive principles and heuristic strategies.
+Cognitive principles are assumptions (not necessarily valid under CL) made by humans while reasoning,
 whereas heuristic strategies describe the methods that are applied when
@@
 The approach presented in~\cite{declare:2017} extended~\cite{cogsci:2017} by specifying three reasoning clusters and two heuristic strategies and achieved a match of 92\%, which was the highest result compared to the presented approaches in~\cite{khemlani:2012}.
-The system developed in~\cite{declare:2017}, was extended and participated at the Syllogism Challenge 2017.\footnote{\url{https://www.cc.uni-freiburg.de/modelingchallenge/challenge-2017}} Figure~\ref{overview:system:syllchal} illustrates the components of the system:
-First, groups of reasoners were specified based on cognitive principles and heuristic strategies.
-After that, 
-a translation into logic program clauses was provided, such that for each reasoning clusters, the corresponding logic program
-was constructed. Next, the desired models according to the Weak Completion Semantics~\cite{hk:2009a,hk:2009b} were computed with a corresponding semantic operator~\cite{stenning:vanlambalgen:2008}.
-After that, the entailed relation between $a$ and $c$ were extracted. If no relation between $a$ and $c$ could be entailed, then \textit{No valid conclusion}
-was derived. The final answer distribution was specified according to the number of reasoning clusters 
+% The system developed in~\cite{declare:2017}, was extended and participated at the Syllogism Challenge 2017.\footnote{\url{https://www.cc.uni-freiburg.de/modelingchallenge/challenge-2017}} 
+Figure~\ref{overview:system:syllchal} illustrates the components of the system:
+First, groups of reasoners were specified based on a set of underlying cognitive principles and heuristic strategies.
+After that, this set was translated into a logic program (Section~\ref{sect:basic}-\ref{sect:advanced}), representing the group of reasoners.
+Next, the desired models were computed (Section~\ref{sect:wcs}-\ref{sect:lfp}).
+After that, the entailed relations between $a$ and $c$ were extracted (Section~\ref{sect:entail}). 
+ The final answer distribution was specified according to the number of reasoning clusters 
 that entailed the conclusion.
-Finally, an illustrative representation of these clusters with respect to their distribution was done by means of of multinomial process trees~\cite{Riefer1988}.
+Finally, an illustrative representation of these clusters with respect to their distribution was done by means of multinomial process trees~\cite{Riefer1988}.
 
@@
 
-As environment for the formalization, we suggest an extension of the
-Weak Completion Semantics (WCS). WCS has first been introduced in~\cite{hk:2009a,hk:2009b} 
-and has shown to model a wide range of human reasoning tasks~\cite{btg:2015:h}. 
-WCS is the non-monotonic computational logic approach to consider
-weakly completed logic programs, to compute their least models
-under the three-valued {\L}ukasiewicz logic~\cite{lukasiewicz:20}, and to reason with respect to these models or search for explanations by means of (skeptical) abduction~\cite{kakas:etal:93}.
+The suggested computational logic approach for the formal representation and reasoning of these clusters,
+is the Weak Completion Semantics (WCS)~\cite{hk:2009a,hk:2009b}
+which has shown to model a wide range of human reasoning tasks~\cite{btg:2015:h}. 
+% WCS is the non-monotonic computational logic approach to consider
+% weakly completed logic programs, to compute their least models
+% under the three-valued {\L}ukasiewicz logic~\cite{lukasiewicz:20}, and to reason with respect to these models or search for explanations by means of (skeptical) abduction~\cite{kakas:etal:93}.
 
@@
 We introduce an extension of WCS first presented in~\cite{dietz:hoelldobler:pereira:2017}. 
-\textit{Contextual logic programs} are (data) logic programs extended by the truth-functional operator~$\cxt$, called \textit{context}. \textit{Contextual (logic) program clauses}
+\textit{Contextual (logic) program clauses}
 are expressions of the forms $A  \leftarrow  L_1 \land \ldots \land L_m \land
@@
 \textit{negative assumptions})\footnote{Under WCS, the negative
-assumption will become $A \leftrightarrow \bot$ and, hence, $A$ has to be false.
-This can, however, be overwritten by other rules and facts (\textit{defeating} the assumption).} and $A \leftarrow \udf$ (called
+assumption will become $A \leftrightarrow \bot$, which, however, can be overwritten by other rules and facts (\textit{defeating} the assumption).} and $A \leftarrow \udf$ (called
 \textit{unknown assumptions}), where
-$A$ is an atom and the~$L_i$ with~$1 \leq i \leq m+p$ is a literal.
+$A$ is an atom and~$L_i$ with~$1 \leq i \leq m+p$, is a literal.
 $A$ is called \textit{head}
@@
 If the clause contains variables, then they are implicitly universally quantified within the scope of the entire clause. 
-For instance, the clause $p(X) \leftarrow q(X)$ represents the formula $(\forall X) (p(X) \leftarrow q(X))$. Therefore, a clause never contains free variables.
+% For instance, the clause $p(X) \leftarrow q(X)$ represents the formula $(\forall X) (p(X) \leftarrow q(X))$. Therefore, a clause never contains free variables.
 A \defname{ground clause} is a clause that does not contain variables.
 $\Ground \CalP$ denotes the set of all its ground clauses.
-$A$ is \textit{defined in $\Ground \CalP$}  iff $\Ground \CalP$ contains a rule, a fact 
-or an assumption with head $A$.
+$A$ is \textit{defined in $\Ground \CalP$}  iff $\Ground \CalP$ contains a clause with head $A$.
 The \textit{definition of $A$ in $\Ground \CalP$} is defined as $\Def(A,\CalP)
-= \{ A \leftarrow \Body \mid A \leftarrow \Body \mbox{ is a rule, a fact or an assumption
+= \{ A \leftarrow \Body \mid A \leftarrow \Body \mbox{ is a clause
 occurring in } \Ground \CalP\}$.
@@
 The set of all atoms that are undefined in~$\Ground \CalP$ is denoted by~$\ud(\CalP)$.
+Consider the following transformation for a given $\CalP$:
+1.\ For each atom~$A$ defined in $\Ground \CalP$, replace
+  all clauses of the form $A \leftarrow \Body_1, \ldots,$ $A \leftarrow
+  \Body_m$ occurring in $\Ground \CalP$ by $ A \leftarrow \Body_1 \vee
+  \ldots \vee \Body_m$. 2.\ Replace all occurrences of $\leftarrow$ by $\leftrightarrow$.
+The obtained ground set of equivalences is called the \textit{weak completion}
+of~$\CalP$ or~$\WC\CalP$.
+
 % $\neg A$ is \textit{assumed in $\Ground \CalP$} iff $\Ground \CalP$
@@
 \]
-\caption{The truth tables for the connectives under the three-valued {\L}ukasiewicz logic and for $\cxt(L)$. $L$ is a literal, $\top$, $\bot$, and $\udf$ denote \textit{true}, \textit{false},
+\caption{The truth tables for the connectives under the three-valued {\L}ukasiewicz logic and for $\cxt$. $L$ is a literal, $\top$, $\bot$, and $\udf$ denote \textit{true}, \textit{false},
 and \textit{unknown}, respectively.\label{tab:3vldluka}}
@@
 
-\subsection{Reasoning with Respect to the Least Fixed Point of $\Svlp$}
-
-For a given $\CalP$, consider the following transformation:
-1.\ For each ground atom~$A$ which occurs as head of a clause in $\Ground \CalP$, replace
-  all clauses of the form $A \leftarrow \Body_1, \ldots,$ $A \leftarrow
-  \Body_m$ occurring in $\Ground \CalP$ by $ A \leftarrow \Body_1 \vee
-  \ldots \vee \Body_m$. 2.\ Replace all occurrences of $\leftarrow$ by $\leftrightarrow$.
-The obtained ground set of equivalences is called the \textit{weak completion}
-of~$\CalP$ or~$\WC\CalP$.
-Consider the following semantic operator, which is due to Stenning and van Lambalgen~\cite{stenning:vanlambalgen:2008}: Let $I = \langle I^\top,I^\bot
+\subsection{Reasoning with Respect to the Least Fixed Point of $\Svlp$}\label{sect:lfp}
+
+Consider the following semantic operator~\cite{stenning:vanlambalgen:2008}: Let $I = \langle I^\top,I^\bot
 \rangle$ be an interpretation. $\Phi_\CalP (I) =
@@
 \]
-\cite{hk:2009a} showed that the
-weak completion of non-contextual programs\footnote{Non-contextual programs refer programs that do not contain rules, in which $\cxt$ appears in the body of any rule, i.e.\ $p = 0$.} 
+The weak completion of non-contextual programs\footnote{Non-contextual programs refer programs that do not contain rules, in which $\cxt$ appears in the body of any rule, i.e.\ $p = 0$.} 
 always has a least model under  {\L}ukasiewicz logic,
-which can be computed by the least fixed point of~$\Svlp$. The correspondence to other well-known logic programming approaches, in particular 
-to the Well-founded semantics has been shown in~\cite{dietz:hoelldobler:wernhard:2014}.
+which can be computed by the least fixed point of~$\Svlp$\cite{hk:2009a}. The correspondence to other well-known logic programming approaches, in particular 
+to the well-founded semantics has been shown in~\cite{dietz:hoelldobler:wernhard:2014}.
 
@@
 the least fixed point of $\Svlp$ always exists. However,
-if these programs do not contain cycles, then there always exists a supported model, which can be 
-computed by iterating $\Svlp$: A least fixed point of $\Svlp$ always exists for contextual programs
-that do not contain cycles~\cite{dietz:hoelldobler:pereira:2017}.
+if these programs do not contain cycles, then there always exists a supported model, which 
+corresponds to the least fixed point of~$\Svlp$~\cite{dietz:hoelldobler:pereira:2017,dietz:hoelldobler:philipp:2017}.
 Let us denote the least fixed point of $\Svlp$ as $\CalM_\CalP$:
@@
 $\CalP \ModelsWCS F$ iff $\CalM_\CalP(F) = \top$.
-Note that for all programs that are considered in the following, a least fixed point of $\Svlp$ exists.
+For all programs considered in the sequel, $\CalM_\CalP$ exists.
 
@@
 
-\subsubsection{Conditional representation, license for inference and existential import}
+\subsubsection{Conditional Representation (\conditionals) and License for Inference (\licenses)}\label{sect:condli}
 Consider \textit{All y are z} ({\MA yz}), which in CL, is represented as $\forall X (y(X) \rightarrow z(X))$.
 Similarly, we assume that humans also understand this statement as a conditional sentence (\conditionals).
-However, it seems that humans do not understand it as \textit{strict}, but defeasible~\cite{stenning:vanlambalgen:2008},
-i.e.\ they implicitly assume a license for the inference (\licenses). For (\MA yz) this can be 
+However, it seems that humans understand them defeasibly, they assume a \textit{license for the inference}~\cite{stenning:vanlambalgen:2008},
+i.e.\ they might allow exceptions to the conditional (\licenses). For (\MA yz) this can be 
 explicitly expressed by 
@@
 % \]
+
+\subsubsection{Existential Import (\import)}
+
 Furthermore, humans have a pragmatic understanding of quantifiers in the language~\cite{grice1975}.
@@
 Consequently, \textit{All $z$} implies \textit{Some $z$ exist}.
-This is known as the \textit{Gricean implicature} or the \textit{existential import} (\import). 
-It also corresponds to the Aristotelian interpretation~\cite{parry:1991} and is assumed by several cognitive theories (cf.~\cite{johnsonlaird:1983,Rips1994}). 
- Likewise,~\cite{stenning:vanlambalgen:2008} have shown that humans
-require existential import for a conditional to be true.
-$\CalP_{Ayz}$, the logic program representing (\MA yz) is as follows:
+This is known as the \textit{Gricean implicature} or the \textit{existential import} (\import),
+which also corresponds to the Aristotelian interpretation~\cite{parry:1991} and is assumed by several cognitive theories (cf.~\cite{johnsonlaird:1983,Rips1994}). 
+%  Likewise,~\cite{stenning:vanlambalgen:2008} have shown that humans
+% require existential import for a conditional to be true.
+Together with the introduced cognitive principles in Section~\ref{sect:condli}, the logic program representing (\MA yz), $\CalP_{Ayz}$, is as follows:
 \begin{align} %{rll@{\hspace{\hfill}}r}
@@
 \end{align}
-Note that $z'(X)$ appears within the $\cxt$ operator in the body of the second clause, where $z'$ represents the negation of $z$:\footnote{$z'(X)$ can be introduced by the negation by transformation principle in the following.}
+Note that $z'(X)$ appears within the $\cxt$ operator in the body of the second clause, where $z'$ represents the negation of $z$:\footnote{$z'(X)$ will be introduced by the (\footnotesize{\textsf{transformation}}) principle in Section~\ref{sect:dneg}.}
 Nothing is abnormal, unless the negation of $z$ is true.
@@
 % \]
-We obtain \[\CalM_{\CalP_{Ayz}} = \langle \{y(o),z(o)\}, \{\Ab_{yz}(o)\} \rangle.\]
-
-\subsubsection{Negation by transformation and no double negation}
+We obtain \quad $\CalM_{\CalP_{Ayz}} = \langle \{y(o),z(o)\}, \{\Ab_{yz}(o)\} \rangle$.
+
+\subsubsection{Negation by Transformation (\transformation) and \\No Double Negation (\dnegation)} \label{sect:dneg}
 
@@
 This is a widely used technique in Logic Programming.
-Another principle that we introduce is the no double negation principle: Under the Weak Completion Semantics, a positive conclusion can be derived from double negation using two conditionals.
-However, it appears to be the case that humans do not reason in such a way~\cite{khemlani:2012}.
+Another principle that we introduce is the no double negation principle: Under the Weak Completion Semantics, a positive conclusion can be derived from double negation appearing in two conditionals.
+However, it seems that humans normally do not reason in such a way~\cite{khemlani:2012}.
 Hence, we block them with the help of abnormalities (\dnegation).
@@
 
-\subsubsection{Converse implication and unknown generalization}
+\subsubsection{Converse Implication (\converse) and Unknown Generalization (\unknownGen)}
 Consider the statement \textit{Some y are z.} ({\MI yz}), which in CL, is represented as $\exists X (y(X) \wedge z(X))$
@@
 In order to distinguish between \textit{some y are z} and \textit{all y are z}, we introduce the following principle:  If we know that \textit{some y are z}, then there must not only be an object~$o_1$, which belongs to $y$ and $z$ but there must be another object $o_2$, which belongs to
-$y$ and for which it is unknown whether it belongs to~$z$.
-$\CalP_{Iyz}$, the logic program representing (\MI yz), is as follows:\footnote{For simplicity, here we omit the clauses introduced by ({\small\converse}). They can be found in~\cite{declare:2017}.}
+$y$ and which does not belong to~$z$.
+$\CalP_{Iyz}$, the logic program representing (\MI yz), is as follows:\footnote{For simplicity, here we omit the clauses introduced by ({\footnotesize\textsf{converse}}).
+They can be found in~\cite{declare:2017}.}
 \begin{align}
@@
 Consider the statement \textit{Some y are not z} ({\MO yz}), which in CL is represented as $\exists X (y(X) \wedge \neg z(X)$.
-As in the previous cases, (\conditionals), (\licenses) and (\unknownGen) apply.
-Furthermore, as the conclusion is again a negation of a claim, we need to assume (\transformation) and
-(\dnegation). $\CalP_{Oyz}$, the logic program representing (\MO yz) is as follows:
+As in the previous cases, (\conditionals), (\licenses) and (\unknownGen) are assumed.
+As the conclusion is again a negation of a claim, (\transformation) and
+(\dnegation) apply as well. $\CalP_{Oyz}$, the logic program representing (\MO yz) is as follows:
 \begin{align}
@@
 The principles introduced here, are assumed not to be applied by all humans, 
-but characterize the reasoning clusters, that will be specified in the following section.
-
-\subsubsection{Contraposition}
+but characterize the reasoning clusters, specified in Section~\ref{sect:aggregateddata}.
+
+\subsubsection{Contraposition (\contraposition)}
 
@@
 There is
-evidence that some of the participants make use of this
-equivalence when solving syllogistic reasoning tasks~\cite{khemlani:2012}. We believe that when they
-encounter a premise with the mood \MA\ (e.g., \textit{all~a are~b}),
-then might reason with the
-contrapositive conditional as well (\contraposition).
+evidence that some of the participants might directly or indirectly 
+apply the contrapositive (\contraposition) within a reasoning task (cf.~\cite{rips:1994,mentallogic:1994}).
 Consider again $\CalP_{Ayz}$, the logic program representing (\MA yz)
@@
 
-\subsubsection{Generalization}  %\label{sect:naf}
-
-If all of the principles introduced so far are applied to an existential
-premise, the only object about which an inference can be made is the one
-resulting from (\import). This is because the
-abnormality introduced by (\licenses) and according to (\unknownGen)
- has to be false only for the object introduced by (\import), but unknown
-for other objects.
-There is, however, evidence that some humans still draw conclusions in such
-circumstances~\cite{khemlani:2012}, i.e.\ they generalize (universally quantify) the existential 
+\subsubsection{Generalization (\negFailure)}  %\label{sect:naf}
+
+If all principles introduced so far are applied to a pair of existentially quantified premises
+(\MI or \MO), then the only object about which an inference can be made is the one
+resulting from (\import) with respect to the premise it introduced (i.e.\ only inferences about the properties 
+of the object about $a$ and $b$ or
+$b$ and $c$). 
+% This is because the
+% abnormality (\licenses) and according to (\unknownGen)
+%  has to be false only for the object introduced by (\import), but unknown
+% for other objects.
+There is, however, evidence that some humans still draw conclusions about $a$ and $c$ in such
+circumstances, i.e.\ they seem to generalize (universally quantify) the existential 
 premise with respect to other objects (\negFailure).
@@
 into a universal one. In order to maintain (\unknownGen), the second clause is added:
-$\Ab_{yz}(o_2) \leftarrow \udf$ guarantees that $\Ab_{yz}(o_2)$ stays unknown, even though $\cxt(z'(o_2))$ might false.
+$\Ab_{yz}(o_2) \leftarrow \udf$ guarantees that $\Ab_{yz}(o_2)$ stays unknown, even though $\cxt(z'(o_2))$ might be false.
 Recall, programs are considered under their weak completion, i.e.\ if $\cxt(z'(o_2))$ is false, then
@@
 
-\subsubsection{Search for Alternative Conclusions}
+\subsubsection{Search for Alternative Conclusions (\abduction)}
 
@@
 facts introduced either from an existential import or from unknown generalization. 
-We use only the first as source for observations, as they are used directly to infer new information. 
+We use the first as source for observations, as they are used directly to infer new information. 
 Consider the following pair of syllogistic premises: 
@@
 for a given observation. The second premise states that some bakers exist, i.e.\ the observation is $\CalO = \{b(o_2) \}$. $\CalO$ can be explained by the first premise, namely that
-these bakers are artists. Under WCS, we formalize this reasoning procedure by means of skeptical abduction for logic programming~\cite{kakas:etal:93}.
-Due to lack of space we will not introduce the formal abductive approach for human syllogistic reasoning in this paper.
-The interested reader is referred to~\cite{cogsci:2017}.
+this baker is an artist. Under WCS, this reasoning procedure was formalized by means of skeptical abduction.
+The interested reader is referred to~\cite{cogsci:2017} for 
+formal abductive approach for human syllogistic reasoning in this paper.
 
@@
 When the contraposition principles applies, then additionally 
-the following condition needs to hold: There exists an object $o_2$ such that $\CalP \ModelsWCS
+the following condition needs to hold as well: There exists an object $o_2$ such that $\CalP \ModelsWCS
 z(o_2)$ and for all objects $o_2$ we find that if $\CalP \ModelsWCS z(o_2)$ then
@@
 $\CalP \not\ModelsWCS \neg z(o_2)$.
-\item[\NVC] When no previous conclusion can be derived, no valid conclusion holds.
+\item[\NVC (no valid conclusion)] iff previous conclusion can be derived.
 \end{description}
 For \MI yz (some) the conversion is explicitly specified, motivated by it (\converse) principle.
-By requiring for \MI\ and \MO, the existence of an object with the first property,
-for which the second property has to be either either false or unknown, \MA\ and \MI, and, \ME\ and \MO, exclude each other, respectively.
-
-\section{Preliminary Evaluation} \label{sect:aggregateddata}
-
-Let us again consider \mbox{\MI\ME 2} from the introduction: 
+By requiring for \MI yz and \MO yz, the existence of an object with the first property,
+for which the second property has to be either either false or unknown, \MA yz and \MI yz, and, \ME yz and \MO yz, exclude each other, respectively.
+
+\section{Evaluation} \label{sect:aggregateddata}
+
+\subsection{On aggregated Data}
+
+As illustration, let us consider \mbox{\MI\ME 2} from the introduction: 
 \begin{align}
@@
 concluded \emph{No chemists are artists} (29\%), \emph{No valid conclusion} (27\%)
-and \emph{Some artists are not chemists} (16\%): Participants who concluded the first answer, did not conclude the other two answers, and vice versa.
-In order to account for these individual differences, three clusters on principles and two on heuristic strategies have been suggested
+and \emph{Some artists are not chemists} (16\%).
+In order to account for the individual differences among participants, three clusters on principles and two on heuristic strategies have been suggested
 in~\cite{declare:2017}. The two heuristic strategies are the atmosphere bias~\cite{woodworth:sells:1935} and the matching bias~\cite{Wetherick1995}.\footnote{
-The interested reader is referred to~\cite{declare:2017} for details on their application on the syllogistic reasoning task. 
+The interested reader is referred to~\cite{declare:2017} for details on their application in syllogistic reasoning. 
 }
-The characterization of the three clusters specified through the cognitive principles is shown Table~\ref{tab:cluster}, together with the conclusions for \MI\ME 2 according to each cluster. 
+The three reasoning clusters have been characterized by the following principles:\footnote{{\footnotesize\textsf{basic}} refers to the set of basic principles introduced in Subsection~\ref{sect:basic}.} (1) \basic, \converse\ for \MI and \abduction,
+    (2) \basic, \converse\ for \MI\ and \negFailure, and
+    (3) \basic, \converse\ for \MI, \ME, and \contraposition\ for~\MA.
+This approach achieved a match of 92\% with the aggregated data reported in~\cite{khemlani:2012}, and was the highest result compared to the presented approaches in~\cite{khemlani:2012}.
+
+Table~\ref{tab:cluster} summarizes the three clusters together with their conclusions for \MI\ME 2.
 \begin{table}
@@
 \begin{tabular}{@{\hspace{0cm}}l@{\hspace{0.9cm}}c@{\hspace{0.9cm}}c@{\hspace{0.9cm}}cGG}
-Principle & Basic & Contraposition & Generalization & Heuristic 1 & Heuristic 2 \medskip \\ \midrule  
+Principle & (1) & (2) & (3) & Heuristic 1 & Heuristic 2 \medskip \\ \midrule  
 \conditionals & $\checkmark$   & $\checkmark$ & $\checkmark$ & - & - \smallskip\\
@@
 \unknownGen & $\checkmark$  & $\checkmark$    & $\checkmark$ & - & -     \smallskip\\
-\abduction &  - & -             & -            & - & - \smallskip\\
 \converse\ for \MI & $\checkmark$  & $\checkmark$      & $\checkmark$ & - & -   \smallskip\\
-\converse\ for \ME & -  & $\checkmark$      & -  & - & -   \smallskip\\
-{\small\textsf{contraposition}} & - & $\checkmark$            & -            & - & - \smallskip\\\midrule
+\abduction &  $\checkmark$ & -             & -            & - & - \smallskip\\
+\negFailure &  - &  $\checkmark$  & -             & - & - \smallskip\\
+\converse\ for \ME & -  & -  & $\checkmark$      & - & -   \smallskip\\
+\contraposition\ for \MA & - & -  & $\checkmark$          & - & - \smallskip\\\midrule
 % matching   &  -            & -             & -            & $\checkmark$ & -   \smallskip\\
@@
 
-Applying the contraposition principle to \textit{No chemists are bakers} yields to \textit{No bakers are chemists}.
+Applying the contraposition principle to \textit{No chemists are bakers} yields \textit{No bakers are chemists}.
 $\Piecontra$ consists of the following clauses:
@@
 Applying the generalization principle to \textit{Some bakers are artists} and its converse,
-yields to \textit{All bakers are artists} and \textit{All artists are bakers}.
+yields to \textit{All bakers are artists} and \textit{All artists are bakers}, respectively.
 $\Piegen$ consists of the clauses in $\CalP_{IE2}$ and the following ones:
@@
 \]
-\textit{No chemists are artists} is entailed by $\CalM_{\Piegen}$ because 
-for all $o$ for which $c(o)$ is true,
-which only applies for $c(o_5)$, we find that $a(o_5)$ is false.
-
-WCS clustering~\cite{declare:2017} suggests three reasoning clusters characterized by the following principles:\footnote{\textsf{basic} refers to the set of basic principles introduced in Subsection~\ref{sect:basic}} (1) \basic, \abduction, and \converse\ for \MI,
-    (2) \basic, \converse\ for \MI\ and \negFailure, and
-    (3) \basic, \converse\ for \MI, \ME, and \contraposition\ for~\MA.\footnote{\cite{declare:2017} also included two heuristic clusters based on
-         atmosphere bias~\cite{Woodworth1935} and matching bias~\cite{Wetherick1995}.}
-This approach achieved a match of 92\% with the aggregated data reported in~\cite{khemlani:2012}, and was the highest result compared to the presented approaches in~\cite{khemlani:2012}.
-
-We were interested whether these clusters were also cognitively adequate in the sense that they 
-would also correspond to the reasoning patterns of individuals.
-Therefore, we investigated the non-aggregated dataset provided by~\cite{khemlani:2016} consisting of the responses of 20 participants for two weeks,
-i.e. the participants carried out the task twice, with a break of one week. The results with respect to WCS clustering is two-fold:
+\textit{No chemists are artists} is entailed by $\CalM_{\Piegen}$, because 
+for all objects which have the property $c$, (which only applies for $o_5$), they do not have the property.
+
+On aggregated data, the clusters performed fairly well with a match of 92\%. 
+We are interested whether these clusters also correspond to the reasoning patterns observed among individuals. 
+For this purpose, we investigated the non-aggregated dataset provided by~\cite{khemlani:2016} consisting of the responses of 20 participants for two weeks,
+i.e.\ the participants carried out the task twice, with a break of one week. 
+The results are two-fold:
 \begin{description}
@@
  the majority's conclusion for 63 out of 64 pair of syllogistic premises and in the second week
- all majority's conclusions were predicted by WCS clustering.
-\item[Non-aggregated evaluation] None of the clusters fitted any individual participants' response pattern, i.e.\ no participant did belong to a single cluster.
+ all majority's conclusions were predicted.
+\item[Non-aggregated evaluation] None of the clusters fitted any individual participants' response pattern, i.e.\ none of the participants belonged to a certain cluster.
 \end{description}
-Given this sobering outcome of the non-aggregated evaluation, 
+
+\clearpage
+\begin{table}[t]\betweentinyandsmall\centering
+\newcolumntype{a}{>{\columncolor{white}}c@{\hspace{2mm}}}
+\newcolumntype{b}{>{\columncolor{gray!40}}c@{\hspace{2mm}}}  
+\captionsetup{font=footnotesize}
+ \begin{tabular}{llllllllll}
+ \toprule
+	 & \MA ac & \ME ac & \MI ac & \MO ac & \MA ca & \ME ca & \MI ca & \MO ca & \NVC \\ \midrule
+
+\rowcolor{lightgray} 
+\MA\MA1  & \basicc&        &        &        &        &        &        &        &      \\ 
+\MA\MA2  &        &        &        &        & \basicc&        &        &        & \\ \rowcolor{lightgray} 
+\MA\MA3  &        &        &        &        &        &        &        &        & \basicc \\
+\MA\MA4  & \basicc&        &        &        &\basicc &        &        &        & \\  \rowcolor{lightgray}  %\midrule
+
+\MA\MI1  &        &        & \abdc  &        &        &        & \abdc  &        & \basicc \\
+\MA\MI2  &        &        & \basicc&        &        &        & \basicc \\  \rowcolor{lightgray} 
+\MA\MI3  &        &        & \abdc  &        &        &        & \abdc  &        & \basicc \\
+\MA\MI4  &        &        & \basicc&        &        &        & \basicc&        & \\%\midrule
+
+\rowcolor{lightgray} 
+\MA\ME1  &        & \basicc&        &        &        &        &        &        & \\ %\rowcolor{lightgray} 
+\MA\ME2  &        &\contrac&        &        &        & \basicc&        &        & \\ \rowcolor{lightgray} 
+\MA\ME3  &        &\contrac&        &        &        &\contrac&        &        & \basicc \\ %\rowcolor{lightgray} 
+\MA\ME4  &        & \basicc&        &        &        &\contrac&        &        &  \\  \rowcolor{lightgray}  %\midrule
+
+\MA\MO1  &        &        &        &\abdc   &        &        &        &        & \basicc \\
+\MA\MO2  &        &        &        &        &        &        &        &\basicc \\  \rowcolor{lightgray} 	
+\MA\MO3  &        &        &        &        &        &        &        &\contrac&\basicc \\
+\MA\MO4  &        &        &        &\basicc \\ %\midrule
+
+\rowcolor{lightgray} 
+\MI\MA1  &        &        &\basicc &        &        &        &\basicc &        & \\ % \rowcolor{lightgray} 
+\MI\MA2  &        &        &\abdc   &        &\genc   &        & \abdc  &        & \basicc \\ \rowcolor{lightgray} 
+\MI\MA3  &        &        & \abdc  &        &        &        & \abdc  &        & \basicc \\ % \rowcolor{lightgray} 
+\MI\MA4  &        &        &\basicc &        &        &        &\basicc &        &  \\  %\midrule
+ \rowcolor{lightgray} 
+\MI\MI1  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\
+\MI\MI2  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\ \rowcolor{lightgray} 
+\MI\MI3  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\
+\MI\MI4  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\ %\midrule
+
+\rowcolor{lightgray} 
+\MI\ME1  &        &        &        &\basicc &        &        &        &        & \\ %\rowcolor{lightgray} 
+\MI\ME2  &        &        &        &\contrac&        & \genc  &        &        & \basicc \\\rowcolor{lightgray} 
+\MI\ME3  &        &        &        &\contrac&        & \genc  &        &        & \basicc  \\% \rowcolor{lightgray} 
+\MI\ME4  &        &        &        & \basicc&        &        &        &        & \\ %\midrule 
+ \rowcolor{lightgray} 
+\MI\MO1  &        &        &        & \genc  &        &        &        &        & \basicc \\
+\MI\MO2  &        &        &        & \genc  &        &        &        &        & \basicc \\  \rowcolor{lightgray} 
+\MI\MO3  &        &        &        & \genc  &        &        &        &        & \basicc \\ 
+\MI\MO4  &        &        &        & \genc  &        &        &        &        & \basicc \\%\midrule
+
+\rowcolor{lightgray} 
+\ME\MA1  &        & \basicc&        &        &        &        &        &        & \\ %\rowcolor{lightgray} 
+\ME\MA2  &        &\contrac&        &        &        & \basicc&        &        &  \\\rowcolor{lightgray} 
+\ME\MA3  &        &\contrac&        &        &        &\contrac&        &        & \basicc \\% \rowcolor{lightgray} 
+\ME\MA4  &        &\contrac&        &        &        & \basicc&        &        &\\  %\midrule 
+ \rowcolor{lightgray} 
+\ME\MI1  &        & \genc   &        &	 &	  &	   &	    &	     & \basicc \\
+\ME\MI2  &        &	       &	&	 &	  &	   &	    & \basicc \\ \rowcolor{lightgray}  
+\ME\MI3  &        & \genc   &        &	 &	  &	   &	    &\contrac& \basicc \\
+\ME\MI4  &        &	       &	&	 &	  &	   &	    & \basicc \\%\midrule
+
+\rowcolor{lightgray} 
+\ME\ME1  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
+\ME\ME2  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ \rowcolor{lightgray} 
+\ME\ME3  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
+\ME\ME4  &        &	       &	&	 &	  &	   &	    &        &\basicc \\%\midrule
+ \rowcolor{lightgray} 
+\ME\MO1  &        &	       &	&	 &	  &	   &	    &        &\basicc \\
+\ME\MO2  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ \rowcolor{lightgray} 
+\ME\MO3  &        &	       &	&	 &	  &	   &	    &        &\basicc \\
+\ME\MO4  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ %\midrule
+
+\rowcolor{lightgray} 
+\MO\MA1  &        &            &	& \basicc&        &        &        &        & \\ %\rowcolor{lightgray} 
+\MO\MA2  &        &	       &	&	 &	  &	   &	    &  \abdc &\basicc \\\rowcolor{lightgray} 
+\MO\MA3  &        &	       &	&\contrac&	  &	   &	    &        &\basicc \\% \rowcolor{lightgray} 
+\MO\MA4  &        &	       &	&	 &	  &	   &	    & \basicc& \\  %\midrule
+ \rowcolor{lightgray} 
+\MO\MI1  &        &	       &	&\genc   &	  &	   &	    &        &\basicc \\
+\MO\MI2  &        &	       &	&	 &	  &	   &	    &\genc   &\basicc \\  \rowcolor{lightgray} 
+\MO\MI3  &        &	       &	&\genc   &	  &	   &	    &        &\basicc \\ 
+\MO\MI4  &        &	       &	&	 &	  &	   &	    &\genc   &\basicc\\ %\midrule
+
+\rowcolor{lightgray} 
+\MO\ME1  &        &	       &	&        &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
+\MO\ME2  &        &	       &	&        &	  &	   &	    &        &\basicc \\\rowcolor{lightgray} 
+\MO\ME3  &        &	       &	&        &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
+\MO\ME4  &        &	       &	&        &	  &	   &	    &        &\basicc \\ %\midrule
+\rowcolor{lightgray} 
+\MO\MO1  &        &	       &	&        &	  &	   &	    &        &\basicc \\
+\MO\MO2  &        &	       &	&        &	  &	   &	    &        &\basicc \\\rowcolor{lightgray}  
+\MO\MO3  &        &	       &	&        &	  &	   &	    &        &\basicc \\
+\MO\MO4  &        &	       &	&        &	  &	   &	    &        &\basicc \\ \bottomrule
+  \end{tabular}
+
+\caption{\label{tab:cogprinc}Predictions according to the cognitive principles as specified in Section~\ref{sect:representing}.\betweentinyandsmall}
+ \end{table}
+
+\thispagestyle{empty}
+\clearpage
+
+Given this sobering outcome with respect to the non-aggregated evaluation, 
 we investigated whether patterns could be found on the level of the
@@
 (iii) Do participants learn, i.e.\ do they improve their logical ability by doing the same reasoning task again.
-We address these questions in the following section.
-
-\clearpage
-\begin{table}[t]\betweentinyandsmall\centering
-\newcolumntype{a}{>{\columncolor{white}}c@{\hspace{2mm}}}
-\newcolumntype{b}{>{\columncolor{gray!40}}c@{\hspace{2mm}}}  
-\captionsetup{font=footnotesize}
- \begin{tabular}{llllllllll}
- \toprule
-	 & \MA ac & \ME ac & \MI ac & \MO ac & \MA ca & \ME ca & \MI ca & \MO ca & \NVC \\ \midrule
-
-\rowcolor{lightgray} 
-\MA\MA1  & \basicc&        &        &        &        &        &        &        &      \\ 
-\MA\MA2  &        &        &        &        & \basicc&        &        &        & \\ \rowcolor{lightgray} 
-\MA\MA3  &        &        &        &        &        &        &        &        & \basicc \\
-\MA\MA4  & \basicc&        &        &        &\basicc &        &        &        & \\  \rowcolor{lightgray}  %\midrule
-
-\MA\MI1  &        &        & \abdc  &        &        &        & \abdc  &        & \basicc \\
-\MA\MI2  &        &        & \basicc&        &        &        & \basicc \\  \rowcolor{lightgray} 
-\MA\MI3  &        &        & \abdc  &        &        &        & \abdc  &        & \basicc \\
-\MA\MI4  &        &        & \basicc&        &        &        & \basicc&        & \\%\midrule
-
-\rowcolor{lightgray} 
-\MA\ME1  &        & \basicc&        &        &        &        &        &        & \\ %\rowcolor{lightgray} 
-\MA\ME2  &        &\contrac&        &        &        & \basicc&        &        & \\ \rowcolor{lightgray} 
-\MA\ME3  &        &\contrac&        &        &        &\contrac&        &        & \basicc \\ %\rowcolor{lightgray} 
-\MA\ME4  &        & \basicc&        &        &        &\contrac&        &        &  \\  \rowcolor{lightgray}  %\midrule
-
-\MA\MO1  &        &        &        &\abdc   &        &        &        &        & \basicc \\
-\MA\MO2  &        &        &        &        &        &        &        &\basicc \\  \rowcolor{lightgray} 	
-\MA\MO3  &        &        &        &        &        &        &        &\contrac&\basicc \\
-\MA\MO4  &        &        &        &\basicc \\ %\midrule
-
-\rowcolor{lightgray} 
-\MI\MA1  &        &        &\basicc &        &        &        &\basicc &        & \\ % \rowcolor{lightgray} 
-\MI\MA2  &        &        &\abdc   &        &\genc   &        & \abdc  &        & \basicc \\ \rowcolor{lightgray} 
-\MI\MA3  &        &        & \abdc  &        &        &        & \abdc  &        & \basicc \\ % \rowcolor{lightgray} 
-\MI\MA4  &        &        &\basicc &        &        &        &\basicc &        &  \\  %\midrule
- \rowcolor{lightgray} 
-\MI\MI1  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\
-\MI\MI2  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\ \rowcolor{lightgray} 
-\MI\MI3  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\
-\MI\MI4  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\ %\midrule
-
-\rowcolor{lightgray} 
-\MI\ME1  &        &        &        &\basicc &        &        &        &        & \\ %\rowcolor{lightgray} 
-\MI\ME2  &        &        &        &\contrac&        & \genc  &        &        & \basicc \\\rowcolor{lightgray} 
-\MI\ME3  &        &        &        &\contrac&        & \genc  &        &        & \basicc  \\% \rowcolor{lightgray} 
-\MI\ME4  &        &        &        & \basicc&        &        &        &        & \\ %\midrule 
- \rowcolor{lightgray} 
-\MI\MO1  &        &        &        & \genc  &        &        &        &        & \basicc \\
-\MI\MO2  &        &        &        & \genc  &        &        &        &        & \basicc \\  \rowcolor{lightgray} 
-\MI\MO3  &        &        &        & \genc  &        &        &        &        & \basicc \\ 
-\MI\MO4  &        &        &        & \genc  &        &        &        &        & \basicc \\%\midrule
-
-\rowcolor{lightgray} 
-\ME\MA1  &        & \basicc&        &        &        &        &        &        & \\ %\rowcolor{lightgray} 
-\ME\MA2  &        &\contrac&        &        &        & \basicc&        &        &  \\\rowcolor{lightgray} 
-\ME\MA3  &        &\contrac&        &        &        &\contrac&        &        & \basicc \\% \rowcolor{lightgray} 
-\ME\MA4  &        &\contrac&        &        &        & \basicc&        &        &\\  %\midrule 
- \rowcolor{lightgray} 
-\ME\MI1  &        & \genc   &        &	 &	  &	   &	    &	     & \basicc \\
-\ME\MI2  &        &	       &	&	 &	  &	   &	    & \basicc \\ \rowcolor{lightgray}  
-\ME\MI3  &        & \genc   &        &	 &	  &	   &	    &\contrac& \basicc \\
-\ME\MI4  &        &	       &	&	 &	  &	   &	    & \basicc \\%\midrule
-
-\rowcolor{lightgray} 
-\ME\ME1  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
-\ME\ME2  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ \rowcolor{lightgray} 
-\ME\ME3  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
-\ME\ME4  &        &	       &	&	 &	  &	   &	    &        &\basicc \\%\midrule
- \rowcolor{lightgray} 
-\ME\MO1  &        &	       &	&	 &	  &	   &	    &        &\basicc \\
-\ME\MO2  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ \rowcolor{lightgray} 
-\ME\MO3  &        &	       &	&	 &	  &	   &	    &        &\basicc \\
-\ME\MO4  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ %\midrule
-
-\rowcolor{lightgray} 
-\MO\MA1  &        &            &	& \basicc&        &        &        &        & \\ %\rowcolor{lightgray} 
-\MO\MA2  &        &	       &	&	 &	  &	   &	    &  \abdc &\basicc \\\rowcolor{lightgray} 
-\MO\MA3  &        &	       &	&\contrac&	  &	   &	    &        &\basicc \\% \rowcolor{lightgray} 
-\MO\MA4  &        &	       &	&	 &	  &	   &	    & \basicc& \\  %\midrule
- \rowcolor{lightgray} 
-\MO\MI1  &        &	       &	&\genc   &	  &	   &	    &        &\basicc \\
-\MO\MI2  &        &	       &	&	 &	  &	   &	    &\genc   &\basicc \\  \rowcolor{lightgray} 
-\MO\MI3  &        &	       &	&\genc   &	  &	   &	    &        &\basicc \\ 
-\MO\MI4  &        &	       &	&	 &	  &	   &	    &\genc   &\basicc\\ %\midrule
-
-\rowcolor{lightgray} 
-\MO\ME1  &        &	       &	&        &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
-\MO\ME2  &        &	       &	&        &	  &	   &	    &        &\basicc \\\rowcolor{lightgray} 
-\MO\ME3  &        &	       &	&        &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
-\MO\ME4  &        &	       &	&        &	  &	   &	    &        &\basicc \\ %\midrule
-\rowcolor{lightgray} 
-\MO\MO1  &        &	       &	&        &	  &	   &	    &        &\basicc \\
-\MO\MO2  &        &	       &	&        &	  &	   &	    &        &\basicc \\\rowcolor{lightgray}  
-\MO\MO3  &        &	       &	&        &	  &	   &	    &        &\basicc \\
-\MO\MO4  &        &	       &	&        &	  &	   &	    &        &\basicc \\ \bottomrule
-  \end{tabular}
-
-\caption{\label{tab:cogprinc}Predictions according to the cognitive principles as specified in Section~\ref{sect:representing}.\betweentinyandsmall}
- \end{table}
-
-\thispagestyle{empty}
-\clearpage
-
-\section{On non-aggregated Data} \label{sect:nonag}
+We address these questions in the following.
+
+\subsection{On non-aggregated Data} \label{sect:nonag}
 
@@
 
-\subsection{mReasoner}
+\subsubsection{mReasoner}
 
@@
 
-\subsection{Evaluation}
+\subsubsection{Individual Reasoning Patterns}
 
@@
 \end{tabular}
-\caption{Overview of each participants' performance with respect to the specified criteria in week 1 and week 2.
+\caption{Overview of the individual participants' responses with respect to the specified criteria in week 1 and week 2.
 $\uparrow$, $\downarrow$, and $\leftrightarrow$ denote above, below, and average, respectively.
@@
 The non-aggregated data provided by~\cite{khemlani:2016} included a dataset of 20 participants,
-for which the task was carried twice, with a break of one week. On average, the basic principles explained the results on about 10\% better than CL.
- Generally, the participants who were strong in CL/ contraposition,drew fewer conclusions according to abduction/ generalization, and vice versa.
+for which the task was carried out twice, with a break of one week. On average, the basic principles explained the results about 10\% better than CL.
+ Generally, the participants who were strong in CL/ contraposition, drew fewer conclusions according to abduction/ generalization, and vice versa.
  The changes within participant's answers from one to the other week was high, with 52/64 on average. 
@@
 Our classification only differs to~\cite{khemlani:2016} for the intermediate and intuitive cluster with respect to three participants (1, 11 and 19).
-A natural explanation for this difference, could be that~\cite{khemlani:2016}
+A possible explanation for this difference, could be that~\cite{khemlani:2016}
 did not take into account the participants' changes between week 1 and week 2.
@@
 \begin{description}
- \item[Deliberative cluster] (6/20) is characterized by the participants whose answers corresponded to the predictions made by basic/ contraposition.
+ \item[Deliberative cluster] (6 out of 20 participants) is characterized by the participants whose answers corresponded to the predictions made by basic/ contraposition.
 Almost none of their conclusions corresponded to predictions made by abduction/ generalization.
@@
  lower than average and concluded higher than average NVC. Generally, their answers improved from week 1 to week 2 with respect to CL.
- \item[Intuitive cluster] (7/20)
+ \item[Intuitive cluster] (7 out of 20 participants)
 is characterized by the participants whose answers corresponded to the predictions given by abduction/ generalization.
@@
  the deliberative cluster.
- \item[Intermediate cluster] (7/20) is characterized by the participants whose answers did not corresponded to any particular principle.
+ \item[Intermediate cluster] (7 out of 20 participants) is characterized by the participants whose answers did not corresponded to any particular principle.
 These participants had a high change within their own answers from week 1 to week 2
- and no particular pattern with respect to NVC conclusions. Generally they were not improving their answers from week 1 to week 2 with respect to CL, but normally compensated for 
- right answers in some cases (that corresponded to contraposition) with wrong answers in the other cases (that correspond to abduction/ generalization).
+ and no particular pattern with respect to NVC conclusions.
+Generally they were not improving their answers from week 1 to week 2 with respect to CL, but normally compensated for answers valid in CL in some cases (that corresponded to contraposition) with answers not valid in CL in the other cases (that correspond to abduction/ generalization).
 \end{description}
@@
 
-We have investigated the cognitive adequacy of WCS clustering with respect to the individual answer patterns reported by~\cite{khemlani:2016}.
+We have investigated the cognitive adequacy of clustering by principles under WCS with respect to the individual answer patterns reported by~\cite{khemlani:2016}.
 \cite{khemlani:2016} is the only approach that we are aware of, that has evaluated their predictions on human syllogistic reasoning
@@
 Similarly, a meta-analysis of the individual response patterns for the selection task has been suggested in~\cite{ragni:2017}.
-This in turn might give us a new starting point to investigate whether the cognitive principles  specified here can characterize participants' conclusions in other reasoning tasks.
+This in turn might give us a new starting point to investigate whether the cognitive principles  specified here can characterize participants' conclusions in other reasoning episodes.
 % episodes.
@@
 % Another aspect that needs to be examined, is
-
-\section{Acknowledgments}
-
-Thanks to the reviewer for pointing us to the relationship of the $\cxt$ operator
-with {\L}ukasiewicz' extensional necessity operator.
 

### 2019-11-07 paper/paper-v3.tex
reference to choice rules, asp book and future work extended with rel luka extensional necessity operator

@@
 by introducing an auxiliary formula $p'(X)$ together with the clause $p(X) \leftarrow \neg p'(X)$ and the integrity constraint $\udf \leftarrow p(X) \wedge p'(X)$ (\transformation).
-This is a widely used technique in Logic Programming.
+This is a widely used technique in Logic Programming (e.g.
+choice rules in~\cite{GebserKaminskiKaufmannSchaub12}).
 Another principle that we introduce is the no double negation principle: Under the Weak Completion Semantics, a positive conclusion can be derived from double negation appearing in two conditionals.
@@
 The individual participants' data allowed us to observe other reasoning patterns (high/ low frequency of NVC conclusions and high/ low frequency of changing answers between week 1 and week 2 answers).
-Interestingly, there was one group of participants for which we could not identify any pattern (intermediate). 
-They seemed to compensate for their wrong/ right answers, and might have changed their strategies within the task.
-Another explanation might be that the set of cognitive principles presented in Section~\ref{sect:representing} is not complete yet.
+Interestingly, there was one group of participants for which we could not identify any pattern (intermediate)
+and might have changed their strategies within the task.
+Furthermore, it is quite likely that the set of cognitive principles presented in Section~\ref{sect:representing} is not complete.
 
@@
 This in turn might give us a new starting point to investigate whether the cognitive principles  specified here can characterize participants' conclusions in other reasoning episodes.
+
+Another aspect that would be intersting to investigate is in how far the $\cxt$ operator within the extended three-valued {\L}ukasiewicz logic,
+originally presented in~\cite{}, 
+ corresponds to the {\L}ukasiewicz extensional necessity operator~\cite{}.\footnote{Thanks to the reviewer for this observation.}
+
 % episodes.

### 2019-11-07 paper/paper-v3.tex
rewrote intro

@@
 from conclusions valid under CL (cf.~\cite{wason:68,byrne:89}).
-Originally, one of the objectives of using (non-classical) logic in the field of AI and logic programming was to formalize human reasoning and commonsense reasoning~\cite{McCarthy59,McCarthy98}.
-% However, the proposed approaches did not have the expected successes~(cf.~\cite{mcdermott:87}),
+Originally, one of the objectives in the field of AI and logic programming was to formalize human reasoning and commonsense reasoning~\cite{McCarthy59,McCarthy98}. 
+However, the proposed approaches have been strongly critized~(cf.~\cite{mcdermott:87}).
+On the other hand, cognitive scientists have extensively investigated and acquired a great expertise
+in human thinking. Recently, several attempts have been made to bridge the gap among these disciplines. 
 % as they were mostly theoretical and did not apply to real case studies.
-On the other hand, cognitive scientists have extensively investigated and acquired a great expertise human thinking.
-However, only recently an attempt was made to bridge the gap among these disciplines.
+% However, only recently an attempt was made 
 % this expertise was not consulted, i.e. there was no communication among the relevant fields.
@@
 
-Another aspect that would be intersting to investigate is in how far the $\cxt$ operator within the extended three-valued {\L}ukasiewicz logic,
-originally presented in~\cite{}, 
- corresponds to the {\L}ukasiewicz extensional necessity operator~\cite{}.\footnote{Thanks to the reviewer for this observation.}
+Another aspect that would be intersting to investigate is in how far the semantics of the $\cxt$ operator within the extended three-valued {\L}ukasiewicz logic in Section~\ref{sub:3valuedL}, 
+ corresponds to the {\L}ukasiewicz extensional necessity operator~\cite{}.\footnote{Thanks to the
+ anonymous reviewer for this observation.}
 

### 2019-11-08 paper/paper-old.tex
added paper version as it was submitted

--- 2019-11-06 paper/paper-v3.tex
@@
 
-Even though the main logical formalism of Classical Logic (CL) has successfully been applied to our scientific reasoning,
-various experimental studies have shown that human reasoning systematically deviates 
-from conclusions valid under CL (cf.~\cite{wason:68,byrne:89}).
-Originally, one of the objectives of using (non-classical) logic in the field of AI and logic programming was to formalize human reasoning and commonsense reasoning~\cite{McCarthy59,McCarthy98}.
-% However, the proposed approaches did not have the expected successes~(cf.~\cite{mcdermott:87}),
-% as they were mostly theoretical and did not apply to real case studies.
-On the other hand, cognitive scientists have extensively investigated and acquired a great expertise human thinking.
-However, only recently an attempt was made to bridge the gap among these disciplines.
-% this expertise was not consulted, i.e. there was no communication among the relevant fields.
-
-Cognitive theories that try to explain human reasoning need to be evaluated by their cognitive adequacy~\cite{strube:1992} and their performance should be compared to other theories.
+Originally, one of the objectives of using logic in the field of AI and logic programming was to formalize human reasoning and commonsense reasoning~\cite{McCarthy59,McCarthy98}.
+However, it did not have the expected successes~(cf.~\cite{mcdermott:87}).
+Although several alternatives have been proposed, such as non-monotonic or many-valued logics, the emphasis had shifted away from their application to real case studies of human reasoning.
+On the other hand, cognitive scientists have extensively investigated human thinking and acquired a great expertise.
+However, this expertise was not consulted, i.e. there was no communication among the relevant fields.
+
+Following~\cite{strube:1992}, we argue that a theory that aims at modeling episodes of human reasoning needs to be evaluable according to its cognitive adequacy and should be comparable and compared to other cognitive theories.
 Conceptual cognitive adequacy describes to what extent a system corresponds to human conceptual knowledge, whereas inferential cognitive adequacy indicates whether the reasoning process of a system is structured similarly to the way humans reason~\cite{knauff:rauh:renz:1997}.
-
-Yet, the cognitive adequacy of a theory can only be measured, if that theory is evaluable, i.e.\ if the theory can uniformally account for episodes of human reasoning systematically, according to its underlying assumptions.
-% Only then, the cognitve adequacy of that theory can be compared to others.
-
-Formal methods can offer the required tools to implement such cognitive theories.
-This in turn, can allow a new way of communication among the fields, by avoiding ambiguous definitions, showing their formal correspondences and finally providing appropriate evaluation methods with respect to their cognitive adequacy.
+Formal methods can offer the required tools to implement elaborate cognitive theories.
+This in turn, can allow a new way of communication among the fields, by avoiding ambiguous understanding of the theories, showing their formal correspondences and finally providing an appropriate evaluation method with respect to their cognitive adequacy.
 In order to determine the required factors that are necessary for the evaluation of an adequate cognitive theory, we take~\cite{ragni:stolzenburg:2015} as a starting point:
-We aim for a theory in which (i) all processes are comprehensible, (ii) the suggested underlying represent general cognitive processes, and (iii) the predictions of our theory are evaluated with respect to a wide range of human reasoning tasks.
-
-The results of the meta-study on syllogistic reasoning in~\cite{khemlani:2012} show that the answers given by humans does not only systematically differ from the valid conclusions under classical logic, but from any other proposed cognitive theory (such as the theories proposed in~\cite{BJ89,johnsonlaird:1983,Rips1994,Polk1995,chater:oaksford:1999}).
+We aim for a theory in which (i) all processes are comprehensible, (ii) the suggested principles represent general cognitive processes, and (iii) the predictions of our theory are evaluated with respect to a wide range of human reasoning tasks.
+% Even though the main logical formalism of Classical Logic (CL) has successfully been applied to our scientific reasoning,
+% various experimental studies have shown that human reasoning systematically deviates 
+% from classical logical conclusions (cf.~\cite{wason:68,byrne:89}).
+
+The results of the meta-study on syllogistic reasoning in~\cite{khemlani:2012} show that humans do not only systematically deviate from the assertions in CL, but from any other proposed cognitive theory (such as the theories proposed in~\cite{BJ89,johnsonlaird:1983,Rips1994,Polk1995,chater:oaksford:1999}).
 % The approach suggested in~\cite{stolzenburg:luederitz:2017}, is based on venn diagrams, where moods are understood as conditional probabilities, and implemented 
@@
 \begin{array}{@{\hspace{0.0cm}}l@{\hspace{0.1cm}}G@{\hspace{0.1cm}}l@{\hspace{0.3cm}}c}\toprule
-	\mbox{Mood} & \mbox{natural language} & \mbox{First-order Logic} &
+	\mbox{mood} & \mbox{natural language} & \mbox{FOL} &
 \mbox{Short} \\\midrule
@@
 		\toprule
-		 \mbox{Figure} & \mbox{Premise 1} & \mbox{Premise 2} \\
+		& \mbox{premise 1} & \mbox{premise 2} \\
 		\midrule
-		\mbox{1} &  a-b & b-c \\
-		\mbox{2} & b-a &  c-b \\
-		\mbox{3} & a-b & c-b \\
-		\mbox{4} & b-a & b-c \\
+		\mbox{figure 1} &  a-b & b-c \\
+		\mbox{figure 2} & b-a &  c-b \\
+		\mbox{figure 3} & a-b & c-b \\
+		\mbox{figure 4} & b-a & b-c \\
 		\bottomrule
@@
 According to~\cite{khemlani:2012}, the majority of participants in experimental studies
-concluded \emph{No chemists are artists} (29\%) and \emph{No valid conclusion} (27\%). The valid conclusion \emph{Some artists are not chemists}
-under CL, was concluded by only 16\%, which is exactly the threshold, that the answer had not been chosen randomly~\cite{khemlani:2012}.
-% \footnote{
-% See
-%  Given that there are nine different possible conclusions, the chance that a conclusion has been chosen randomly is 1/9 = 11.1 \%. 
-%  A binomial test shows that if a conclusion is drawn in
-% more than 16\% of the cases by the participants it is unlikely that is has been
-% chosen by just random guesses~\cite{khemlani:2012}. }
+concluded \emph{No chemists are artists} (29\%) and \emph{No valid conclusion} (27\%). The classical logical valid conclusion \emph{Some artists are not chemists}, was concluded 
+by only 16\%, which is exactly the threshold, that the answer had not been chosen randomly.\footnote{
+ Given that there are nine different possible conclusions, the chance that a conclusion has been chosen randomly is 1/9 = 11.1 \%. 
+ A binomial test shows that if a conclusion is drawn in
+more than 16\% of the cases by the participants it is unlikely that is has been
+chosen by just random guesses~\cite{khemlani:2012}. }
 % Recently, a new cognitive theory has been proposed for syllogistic reasoning~\cite{cogsci:2017}.
@@
 % The majority's conclusions exclude each other, as the test was made such that 
-Usually in these experiments, participants are supposed to give only one answer, thus all three (significant) answers
-were given by different participants. 
-% The ones who answered \emph{No valid conclusion}
-% were not the same ones who answered \emph{No chemists are artists} or \emph{Some artists are not chemists}, and vice versa.\footnote{}
+The ones who answered \emph{No valid conclusion}
+were not the same ones who answered \emph{No chemists are artists} or \emph{Some artists are not chemists}, and vice versa.
 
@@
 
-As the above example and recent investigations on psychological experiments show (cf.~\cite{khemlani:2016, ragni:2017}), it seems that different humans represent and reason about conditionals differently.
+As the above example and recent investigations on psychological experiments show (cf.~\cite{khemlani:2016, ragni:2017}), it seems that different humans represent and reason about conditional information differently.
 This implies that \textit{the} human reasoner does not exist, instead there 
@@
 in the sense that we investigate the individual participants' conclusions with respect to the non-aggregated dataset provided by~\cite{khemlani:2016}.
-In Section~\ref{sect:wcsclustering} and~\ref{sect:representing}, we introduce the underlying computational logic approach and the formal representation of cognitive principles within this approach.
+In Section~\ref{sect:wcsclustering} and~\ref{sect:representing}, we introduce the underlying computational logic approach and the cognitive principles.
 Section~\ref{sect:aggregateddata} illustrates an example and briefly discusses the performance of the approach in~\cite{declare:2017} on non-aggregated data.
-Section~\ref{sect:nonag} provides a more adequate characterization of the individual's reasoning patterns.
-
-\section{Clustering by Principles under WCS} \label{sect:wcsclustering}
-
-According to~\cite{declare:2017}, reasoning clusters can be expressed through cognitive principles and heuristic strategies.
-Cognitive principles are assumptions (not necessarily valid under CL) made by humans while reasoning,
+Section~\ref{sect:nonag} provides a new characterization of the individual's reasoning patterns that 
+seems to be more adequate.
+
+\section{WCS Clustering} \label{sect:wcsclustering}
+
+\cite{declare:2017} puts forward the hypothesis that reasoning clusters can be expressed through cognitive principles and heuristic strategies.
+Cognitive principles are (not necessarily valid) assumptions made by humans while reasoning,
 whereas heuristic strategies describe the methods that are applied when
 conclusions are not derived by reasoning but instead based on structural patterns or {biases}.
-The approach suggested in~\cite{declare:2017} is based on a new computational logic approach for human syllogistic reasoning~\cite{cogsci:2017}, which seems to perform better than other state-of-the-art cognitive theories when evaluated with aggregated data provided in~\cite{khemlani:2012}.
-The approach presented in~\cite{declare:2017} extended~\cite{cogsci:2017} by specifying three reasoning clusters and two heuristic strategies and achieved a match of 92\%, which was the highest result compared to the presented approaches in~\cite{khemlani:2012}.
-% The system developed in~\cite{declare:2017}, was extended and participated at the Syllogism Challenge 2017.\footnote{\url{https://www.cc.uni-freiburg.de/modelingchallenge/challenge-2017}} 
-Figure~\ref{overview:system:syllchal} illustrates the components of the system:
-First, groups of reasoners were specified based on a set of underlying cognitive principles and heuristic strategies.
-After that, this set was translated into a logic program (Section~\ref{sect:basic}-\ref{sect:advanced}), representing the group of reasoners.
-Next, the desired models were computed (Section~\ref{sect:wcs}-\ref{sect:lfp}).
-After that, the entailed relations between $a$ and $c$ were extracted (Section~\ref{sect:entail}). 
- The final answer distribution was specified according to the number of reasoning clusters 
+The approach suggested in~\cite{declare:2017} is based on a new computational logic approach for human syllogistic reasoning~\cite{cogsci:2017}, which seems to perform better than other state-of-the-art cognitive theories when evaluated with aggregated data from~\cite{khemlani:2012}.
+\cite{declare:2017} extended the approach in~\cite{cogsci:2017} by specifying three reasoning clusters and two heuristic strategies and achieved a match of 92\%, which was the highest result compared to the presented approaches in~\cite{khemlani:2012}.
+The system developed in~\cite{declare:2017}, was extended and participated at the Syllogism Challenge 2017.\footnote{\url{https://www.cc.uni-freiburg.de/modelingchallenge/challenge-2017}} Figure~\ref{overview:system:syllchal} illustrates the components of the system:
+First, groups of reasoners were specified based on cognitive principles and heuristic strategies.
+After that, 
+a translation into logic program clauses was provided, such that for each reasoning clusters, the corresponding logic program
+was constructed. Next, the desired models according to the Weak Completion Semantics~\cite{hk:2009a,hk:2009b} were computed with a corresponding semantic operator~\cite{stenning:vanlambalgen:2008}.
+After that, the entailed relation between $a$ and $c$ were extracted. If no relation between $a$ and $c$ could be entailed, then \textit{No valid conclusion}
+was derived. The final answer distribution was specified according to the number of reasoning clusters 
 that entailed the conclusion.
-Finally, an illustrative representation of these clusters with respect to their distribution was done by means of multinomial process trees~\cite{Riefer1988}.
+Finally, an illustrative representation of these clusters with respect to their distribution was done by means of of multinomial process trees~\cite{Riefer1988}.
 
@@
 
-The suggested computational logic approach for the formal representation and reasoning of these clusters,
-is the Weak Completion Semantics (WCS)~\cite{hk:2009a,hk:2009b}
-which has shown to model a wide range of human reasoning tasks~\cite{btg:2015:h}. 
-% WCS is the non-monotonic computational logic approach to consider
-% weakly completed logic programs, to compute their least models
-% under the three-valued {\L}ukasiewicz logic~\cite{lukasiewicz:20}, and to reason with respect to these models or search for explanations by means of (skeptical) abduction~\cite{kakas:etal:93}.
+As environment for the formalization, we suggest an extension of the
+Weak Completion Semantics (WCS). WCS has first been introduced in~\cite{hk:2009a,hk:2009b} 
+and has shown to model a wide range of human reasoning tasks~\cite{btg:2015:h}. 
+WCS is the non-monotonic computational logic approach to consider
+weakly completed logic programs, to compute their least models
+under the three-valued {\L}ukasiewicz logic~\cite{lukasiewicz:20}, and to reason with respect to these models or search for explanations by means of (skeptical) abduction~\cite{kakas:etal:93}.
 
@@
 We introduce an extension of WCS first presented in~\cite{dietz:hoelldobler:pereira:2017}. 
-\textit{Contextual (logic) program clauses}
+\textit{Contextual logic programs} are (data) logic programs extended by the truth-functional operator~$\cxt$, called \textit{context}. \textit{Contextual (logic) program clauses}
 are expressions of the forms $A  \leftarrow  L_1 \land \ldots \land L_m \land
@@
 \textit{negative assumptions})\footnote{Under WCS, the negative
-assumption will become $A \leftrightarrow \bot$, which, however, can be overwritten by other rules and facts (\textit{defeating} the assumption).} and $A \leftarrow \udf$ (called
+assumption will become $A \leftrightarrow \bot$ and, hence, $A$ has to be false.
+This can, however, be overwritten by other rules and facts (\textit{defeating} the assumption).} and $A \leftarrow \udf$ (called
 \textit{unknown assumptions}), where
-$A$ is an atom and~$L_i$ with~$1 \leq i \leq m+p$, is a literal.
+$A$ is an atom and the~$L_i$ with~$1 \leq i \leq m+p$ are
+literals.
 $A$ is called \textit{head}
@@
 If the clause contains variables, then they are implicitly universally quantified within the scope of the entire clause. 
-% For instance, the clause $p(X) \leftarrow q(X)$ represents the formula $(\forall X) (p(X) \leftarrow q(X))$. Therefore, a clause never contains free variables.
+For instance, the clause $p(X) \leftarrow q(X)$ represents the formula $(\forall X) (p(X) \leftarrow q(X))$. Therefore, a clause never contains free variables.
 A \defname{ground clause} is a clause that does not contain variables.
 $\Ground \CalP$ denotes the set of all its ground clauses.
-$A$ is \textit{defined in $\Ground \CalP$}  iff $\Ground \CalP$ contains a clause with head $A$.
+$A$ is \textit{defined in $\Ground \CalP$}  iff $\Ground \CalP$ contains a rule, a fact 
+or an assumption with head $A$.
 The \textit{definition of $A$ in $\Ground \CalP$} is defined as $\Def(A,\CalP)
-= \{ A \leftarrow \Body \mid A \leftarrow \Body \mbox{ is a clause
+= \{ A \leftarrow \Body \mid A \leftarrow \Body \mbox{ is a rule, a fact or an assumption
 occurring in } \Ground \CalP\}$.
@@
 The set of all atoms that are undefined in~$\Ground \CalP$ is denoted by~$\ud(\CalP)$.
-Consider the following transformation for a given $\CalP$:
-1.\ For each atom~$A$ defined in $\Ground \CalP$, replace
-  all clauses of the form $A \leftarrow \Body_1, \ldots,$ $A \leftarrow
-  \Body_m$ occurring in $\Ground \CalP$ by $ A \leftarrow \Body_1 \vee
-  \ldots \vee \Body_m$. 2.\ Replace all occurrences of $\leftarrow$ by $\leftrightarrow$.
-The obtained ground set of equivalences is called the \textit{weak completion}
-of~$\CalP$ or~$\WC\CalP$.
-
 % $\neg A$ is \textit{assumed in $\Ground \CalP$} iff $\Ground \CalP$
@@
 and \textit{unknown}, respectively.
-% As it seems,  $\cxt$ corresponds to {\L}ukasiewicz necessity operator~\cite{}, where the difference is, that $\cxt$ is restricted to be applicable to literals only.
 A \defname{three-valued interpretation}~$I$ is a mapping from the set of
@@
 \]
-\caption{The truth tables for the connectives under the three-valued {\L}ukasiewicz logic and for $\cxt$. $L$ is a literal, $\top$, $\bot$, and $\udf$ denote \textit{true}, \textit{false},
+\caption{The truth tables for the connectives under the three-valued {\L}ukasiewicz logic and for $\cxt(L)$. $L$ is a literal, $\top$, $\bot$, and $\udf$ denote \textit{true}, \textit{false},
 and \textit{unknown}, respectively.\label{tab:3vldluka}}
@@
 
-\subsection{Reasoning with Respect to the Least Fixed Point of $\Svlp$}\label{sect:lfp}
-
-Consider the following semantic operator~\cite{stenning:vanlambalgen:2008}: Let $I = \langle I^\top,I^\bot
+\subsection{Reasoning with Respect to the Least Fixed Point of $\Svlp$}
+
+For a given $\CalP$, consider the following transformation:
+1.\ For each ground atom~$A$ which occurs as head of a clause in $\Ground \CalP$, replace
+  all clauses of the form $A \leftarrow \Body_1, \ldots,$ $A \leftarrow
+  \Body_m$ occurring in $\Ground \CalP$ by $ A \leftarrow \Body_1 \vee
+  \ldots \vee \Body_m$. 2.\ Replace all occurrences of $\leftarrow$ by $\leftrightarrow$.
+The obtained ground set of equivalences is called the \textit{weak completion}
+of~$\CalP$ or~$\WC\CalP$.
+Consider the following semantic operator, which is due to Stenning and van Lambalgen~\cite{stenning:vanlambalgen:2008}: Let $I = \langle I^\top,I^\bot
 \rangle$ be an interpretation. $\Phi_\CalP (I) =
@@
 \]
-The weak completion of non-contextual programs\footnote{Non-contextual programs refer programs that do not contain rules, in which $\cxt$ appears in the body of any rule, i.e.\ $p = 0$.} 
+\cite{hk:2009a} showed that the
+weak completion of non-contextual programs\footnote{Non-contextual programs refer programs that do not contain rules, in which $\cxt$ appears in the body of any rule, i.e.\ $p = 0$.} 
 always has a least model under  {\L}ukasiewicz logic,
-which can be computed by the least fixed point of~$\Svlp$\cite{hk:2009a}. The correspondence to other well-known logic programming approaches, in particular 
-to the well-founded semantics has been shown in~\cite{dietz:hoelldobler:wernhard:2014}.
+which can be computed by the least fixed point of~$\Svlp$. The correspondence to other well-known logic programming approaches, in particular 
+to the Well-founded semantics has been shown in~\cite{dietz:hoelldobler:wernhard:2014}.
 
@@
 the least fixed point of $\Svlp$ always exists. However,
-if these programs do not contain cycles, then there always exists a supported model, which 
-corresponds to the least fixed point of~$\Svlp$~\cite{dietz:hoelldobler:pereira:2017,dietz:hoelldobler:philipp:2017}.
+if these programs do not contain cycles, then there always exists a supported model, which can be 
+computed by iterating $\Svlp$: A least fixed point of $\Svlp$ always exists for contextual programs
+that do not contain cycles~\cite{dietz:hoelldobler:pereira:2017}.
 Let us denote the least fixed point of $\Svlp$ as $\CalM_\CalP$:
@@
 $\CalP \ModelsWCS F$ iff $\CalM_\CalP(F) = \top$.
-For all programs considered in the sequel, $\CalM_\CalP$ exists.
+Note that for all programs that are considered in the following, a least fixed point of $\Svlp$ exists.
 
@@
 
-\subsubsection{Conditional Representation (\conditionals) and License for Inference (\licenses)}\label{sect:condli}
+\subsubsection{Conditional representation, license for inference and existential import}
 Consider \textit{All y are z} ({\MA yz}), which in CL, is represented as $\forall X (y(X) \rightarrow z(X))$.
 Similarly, we assume that humans also understand this statement as a conditional sentence (\conditionals).
-However, it seems that humans understand them defeasibly, they assume a \textit{license for the inference}~\cite{stenning:vanlambalgen:2008},
-i.e.\ they might allow exceptions to the conditional (\licenses). For (\MA yz) this can be 
+However, it seems that humans do not understand it as \textit{strict}, but defeasible~\cite{stenning:vanlambalgen:2008},
+i.e.\ they implicitly assume a license for the inference (\licenses). For (\MA yz) this can be 
 explicitly expressed by 
@@
 % \]
-
-\subsubsection{Existential Import (\import)}
-
 Furthermore, humans have a pragmatic understanding of quantifiers in the language~\cite{grice1975}.
@@
 Consequently, \textit{All $z$} implies \textit{Some $z$ exist}.
-This is known as the \textit{Gricean implicature} or the \textit{existential import} (\import),
-which also corresponds to the Aristotelian interpretation~\cite{parry:1991} and is assumed by several cognitive theories (cf.~\cite{johnsonlaird:1983,Rips1994}). 
-%  Likewise,~\cite{stenning:vanlambalgen:2008} have shown that humans
-% require existential import for a conditional to be true.
-Together with the introduced cognitive principles in Section~\ref{sect:condli}, the logic program representing (\MA yz), $\CalP_{Ayz}$, is as follows:
+This is known as the \textit{Gricean implicature} or the \textit{existential import} (\import). 
+It is assumed by several cognitive theories (cf.~\cite{johnsonlaird:1983,Rips1994}). 
+ Likewise,~\cite{stenning:vanlambalgen:2008} have shown that humans
+require existential import for a conditional to be true.
+$\CalP_{Ayz}$, the logic program representing (\MA yz) is as follows:
 \begin{align} %{rll@{\hspace{\hfill}}r}
@@
 \end{align}
-Note that $z'(X)$ appears within the $\cxt$ operator in the body of the second clause, where $z'$ represents the negation of $z$:\footnote{$z'(X)$ will be introduced by the (\footnotesize{\textsf{transformation}}) principle in Section~\ref{sect:dneg}.}
+Note that $z'(X)$ appears within the $\cxt$ operator in the body of the second clause, where $z'$ represents the negation of $z$:\footnote{$z'(X)$ can be introduced by the negation by transformation principle in the following.}
 Nothing is abnormal, unless the negation of $z$ is true.
@@
 % \]
-We obtain \quad $\CalM_{\CalP_{Ayz}} = \langle \{y(o),z(o)\}, \{\Ab_{yz}(o)\} \rangle$.
-
-\subsubsection{Negation by Transformation (\transformation) and \\No Double Negation (\dnegation)} \label{sect:dneg}
+We obtain \[\CalM_{\CalP_{Ayz}} = \langle \{y(o),z(o)\}, \{\Ab_{yz}(o)\} \rangle.\]
+
+\subsubsection{Negation by transformation and no double negation}
 
@@
 This is a widely used technique in Logic Programming.
-Another principle that we introduce is the no double negation principle: Under the Weak Completion Semantics, a positive conclusion can be derived from double negation appearing in two conditionals.
-However, it seems that humans normally do not reason in such a way~\cite{khemlani:2012}.
+Another principle that we introduce is the no double negation principle: Under the Weak Completion Semantics, a positive conclusion can be derived from double negation using two conditionals.
+However, it appears to be the case that humans do not reason in such a way~\cite{khemlani:2012}.
 Hence, we block them with the help of abnormalities (\dnegation).
@@
 
-\subsubsection{Converse Implication (\converse) and Unknown Generalization (\unknownGen)}
+\subsubsection{Converse implication and unknown generalization}
 Consider the statement \textit{Some y are z.} ({\MI yz}), which in CL, is represented as $\exists X (y(X) \wedge z(X))$
@@
 In order to distinguish between \textit{some y are z} and \textit{all y are z}, we introduce the following principle:  If we know that \textit{some y are z}, then there must not only be an object~$o_1$, which belongs to $y$ and $z$ but there must be another object $o_2$, which belongs to
-$y$ and which does not belong to~$z$.
-$\CalP_{Iyz}$, the logic program representing (\MI yz), is as follows:\footnote{For simplicity, here we omit the clauses introduced by ({\footnotesize\textsf{converse}}).
-They can be found in~\cite{declare:2017}.}
+$y$ and for which it is unknown whether it belongs to~$z$.
+$\CalP_{Iyz}$, the logic program representing (\MI yz), is as follows:\footnote{For simplicity, here we omit the clauses introduced by ({\small\converse}). They can be found in~\cite{declare:2017}.}
 \begin{align}
@@
 Consider the statement \textit{Some y are not z} ({\MO yz}), which in CL is represented as $\exists X (y(X) \wedge \neg z(X)$.
-As in the previous cases, (\conditionals), (\licenses) and (\unknownGen) are assumed.
-As the conclusion is again a negation of a claim, (\transformation) and
-(\dnegation) apply as well. $\CalP_{Oyz}$, the logic program representing (\MO yz) is as follows:
+As in the previous cases, (\conditionals), (\licenses) and (\unknownGen) apply.
+Furthermore, as the conclusion is again a negation of a claim, we need to assume (\transformation) and
+(\dnegation). $\CalP_{Oyz}$, the logic program representing (\MO yz) is as follows:
 \begin{align}
@@
 The principles introduced here, are assumed not to be applied by all humans, 
-but characterize the reasoning clusters, specified in Section~\ref{sect:aggregateddata}.
-
-\subsubsection{Contraposition (\contraposition)}
+but characterize the reasoning clusters, that will be specified in the following section.
+
+\subsubsection{Contraposition}
 
@@
 There is
-evidence that some of the participants might directly or indirectly 
-apply the contrapositive (\contraposition) within a reasoning task (cf.~\cite{rips:1994,mentallogic:1994}).
+evidence that some of the participants make use of this
+equivalence when solving syllogistic reasoning tasks~\cite{khemlani:2012}. We believe that when they
+encounter a premise with the mood \MA\ (e.g., \textit{all~a are~b}),
+then might reason with the
+contrapositive conditional as well (\contraposition).
 Consider again $\CalP_{Ayz}$, the logic program representing (\MA yz)
@@
 
-\subsubsection{Generalization (\negFailure)}  %\label{sect:naf}
-
-If all principles introduced so far are applied to a pair of existentially quantified premises
-(\MI or \MO), then the only object about which an inference can be made is the one
-resulting from (\import) with respect to the premise it introduced (i.e.\ only inferences about the properties 
-of the object about $a$ and $b$ or
-$b$ and $c$). 
-% This is because the
-% abnormality (\licenses) and according to (\unknownGen)
-%  has to be false only for the object introduced by (\import), but unknown
-% for other objects.
-There is, however, evidence that some humans still draw conclusions about $a$ and $c$ in such
-circumstances, i.e.\ they seem to generalize (universally quantify) the existential 
+\subsubsection{Generalization}  %\label{sect:naf}
+
+If all of the principles introduced so far are applied to an existential
+premise, the only object about which an inference can be made is the one
+resulting from (\import). This is because the
+abnormality introduced by (\licenses) and according to (\unknownGen)
+ has to be false only for the object introduced by (\import), but unknown
+for other objects.
+There is, however, evidence that some humans still draw conclusions in such
+circumstances~\cite{khemlani:2012}, i.e.\ they generalize (universally quantify) the existential 
 premise with respect to other objects (\negFailure).
@@
 into a universal one. In order to maintain (\unknownGen), the second clause is added:
-$\Ab_{yz}(o_2) \leftarrow \udf$ guarantees that $\Ab_{yz}(o_2)$ stays unknown, even though $\cxt(z'(o_2))$ might be false.
+$\Ab_{yz}(o_2) \leftarrow \udf$ guarantees that $\Ab_{yz}(o_2)$ stays unknown, even though $\cxt(z'(o_2))$ might false.
 Recall, programs are considered under their weak completion, i.e.\ if $\cxt(z'(o_2))$ is false, then
@@
 
-\subsubsection{Search for Alternative Conclusions (\abduction)}
+\subsubsection{Search for Alternative Conclusions}
 
@@
 facts introduced either from an existential import or from unknown generalization. 
-We use the first as source for observations, as they are used directly to infer new information. 
+We use only the first as source for observations, as they are used directly to infer new information. 
 Consider the following pair of syllogistic premises: 
@@
 for a given observation. The second premise states that some bakers exist, i.e.\ the observation is $\CalO = \{b(o_2) \}$. $\CalO$ can be explained by the first premise, namely that
-this baker is an artist. Under WCS, this reasoning procedure was formalized by means of skeptical abduction.
-The interested reader is referred to~\cite{cogsci:2017} for 
-formal abductive approach for human syllogistic reasoning in this paper.
+these bakers are artists. Under WCS, we formalize this reasoning procedure by means of skeptical abduction for logic programming~\cite{kakas:etal:93}.
+Due to lack of space we will not introduce the formal abductive approach for human syllogistic reasoning in this paper.
+The interested reader is referred to~\cite{cogsci:2017}.
 
@@
 When the contraposition principles applies, then additionally 
-the following condition needs to hold as well: There exists an object $o_2$ such that $\CalP \ModelsWCS
+the following condition needs to hold: There exists an object $o_2$ such that $\CalP \ModelsWCS
 z(o_2)$ and for all objects $o_2$ we find that if $\CalP \ModelsWCS z(o_2)$ then
@@
 $\CalP \not\ModelsWCS z(o_2)$ and
-% there exists an object $o_3$ such that $\CalP \ModelsWCS
-% z(o_3) \wedge y(o_3)$ and 
-there exists an
-object $o_3$ such that $\CalP \ModelsWCS z(o_3)$ and
-$\CalP \not\ModelsWCS y(o_3)$.
-\item[\MO yz (some are not)]
+there exists an object $o_3$ such that $\CalP \ModelsWCS
+z(o_3) \wedge y(o_3)$ and there exists an
+object $o_4$ such that $\CalP \ModelsWCS z(o_4)$ and
+$\CalP \not\ModelsWCS y(o_4)$.
+\item[\MO yz (Some Are Not)]
 $\CalP \models Oyz$ iff there exists an object $o_1$ such that $\CalP \ModelsWCS
@@
 $\CalP \not\ModelsWCS \neg z(o_2)$.
-\item[\NVC (no valid conclusion)] iff previous conclusion can be derived.
+\item[\NVC] When no previous conclusion can be derived, no valid conclusion holds.
 \end{description}
-For \MI yz (some) the conversion is explicitly specified, motivated by it (\converse) principle.
-By requiring for \MI yz and \MO yz, the existence of an object with the first property,
-for which the second property has to be either either false or unknown, \MA yz and \MI yz, and, \ME yz and \MO yz, exclude each other, respectively.
-
-\section{Evaluation} \label{sect:aggregateddata}
-
-\subsection{On aggregated Data}
-
-As illustration, let us consider \mbox{\MI\ME 2} from the introduction: 
+
+\section{Preliminary Evaluation} \label{sect:aggregateddata}
+
+Let us again consider \mbox{\MI\ME 2} from the introduction: 
 \begin{align}
@@
 concluded \emph{No chemists are artists} (29\%), \emph{No valid conclusion} (27\%)
-and \emph{Some artists are not chemists} (16\%).
-In order to account for the individual differences among participants, three clusters on principles and two on heuristic strategies have been suggested
+and \emph{Some artists are not chemists} (16\%): Participants who concluded the first answer, did not conclude the other two answers, and vice versa.
+In order to account for these individual differences, three clusters on principles and two on heuristic strategies have been suggested
 in~\cite{declare:2017}. The two heuristic strategies are the atmosphere bias~\cite{woodworth:sells:1935} and the matching bias~\cite{Wetherick1995}.\footnote{
-The interested reader is referred to~\cite{declare:2017} for details on their application in syllogistic reasoning. 
+The interested reader is referred to~\cite{declare:2017} for details on their application on the syllogistic reasoning task. 
 }
-The three reasoning clusters have been characterized by the following principles:\footnote{{\footnotesize\textsf{basic}} refers to the set of basic principles introduced in Subsection~\ref{sect:basic}.} (1) \basic, \converse\ for \MI and \abduction,
-    (2) \basic, \converse\ for \MI\ and \negFailure, and
-    (3) \basic, \converse\ for \MI, \ME, and \contraposition\ for~\MA.
-This approach achieved a match of 92\% with the aggregated data reported in~\cite{khemlani:2012}, and was the highest result compared to the presented approaches in~\cite{khemlani:2012}.
-
-Table~\ref{tab:cluster} summarizes the three clusters together with their conclusions for \MI\ME 2.
+The characterization of the three clusters specified through the cognitive principles is shown Table~\ref{tab:cluster}, together with the conclusions for \MI\ME 2 according to each cluster. 
 \begin{table}
@@
 \begin{tabular}{@{\hspace{0cm}}l@{\hspace{0.9cm}}c@{\hspace{0.9cm}}c@{\hspace{0.9cm}}cGG}
-Principle & (1) & (2) & (3) & Heuristic 1 & Heuristic 2 \medskip \\ \midrule  
+Principle & Basic & Contraposition & Generalization & Heuristic 1 & Heuristic 2 \medskip \\ \midrule  
 \conditionals & $\checkmark$   & $\checkmark$ & $\checkmark$ & - & - \smallskip\\
@@
 \unknownGen & $\checkmark$  & $\checkmark$    & $\checkmark$ & - & -     \smallskip\\
+\abduction &  - & -             & -            & - & - \smallskip\\
 \converse\ for \MI & $\checkmark$  & $\checkmark$      & $\checkmark$ & - & -   \smallskip\\
-\abduction &  $\checkmark$ & -             & -            & - & - \smallskip\\
-\negFailure &  - &  $\checkmark$  & -             & - & - \smallskip\\
-\converse\ for \ME & -  & -  & $\checkmark$      & - & -   \smallskip\\
-\contraposition\ for \MA & - & -  & $\checkmark$          & - & - \smallskip\\\midrule
+\converse\ for \ME & -  & $\checkmark$      & -  & - & -   \smallskip\\
+{\small\textsf{contraposition}} & - & $\checkmark$            & -            & - & - \smallskip\\\midrule
 % matching   &  -            & -             & -            & $\checkmark$ & -   \smallskip\\
@@
 
-Applying the contraposition principle to \textit{No chemists are bakers} yields \textit{No bakers are chemists}.
+Applying the contraposition principle to \textit{No chemists are bakers} yields to \textit{No bakers are chemists}.
 $\Piecontra$ consists of the following clauses:
@@
 Applying the generalization principle to \textit{Some bakers are artists} and its converse,
-yields to \textit{All bakers are artists} and \textit{All artists are bakers}, respectively.
+yields to \textit{All bakers are artists} and \textit{All artists are bakers}.
 $\Piegen$ consists of the clauses in $\CalP_{IE2}$ and the following ones:
@@
 \]
-\textit{No chemists are artists} is entailed by $\CalM_{\Piegen}$, because 
-for all objects which have the property $c$, (which only applies for $o_5$), they do not have the property.
-
-On aggregated data, the clusters performed fairly well with a match of 92\%. 
-We are interested whether these clusters also correspond to the reasoning patterns observed among individuals. 
-For this purpose, we investigated the non-aggregated dataset provided by~\cite{khemlani:2016} consisting of the responses of 20 participants for two weeks,
-i.e.\ the participants carried out the task twice, with a break of one week. 
-The results are two-fold:
+\textit{No chemists are artists} is entailed by $\CalM_{\Piegen}$ because 
+for all $o$ for which $c(o)$ is true,
+which only applies for $c(o_5)$, we find that $a(o_5)$ is false.
+
+WCS clustering~\cite{declare:2017} suggests three reasoning clusters characterized by the following principles:\footnote{\textsf{basic} refers to the set of basic principles introduced in Subsection~\ref{sect:basic}} (1) \basic, \abduction, and \converse\ for \MI,
+    (2) \basic, \converse\ for \MI\ and \negFailure, and
+    (3) \basic, \converse\ for \MI, \ME, and \contraposition\ for~\MA.\footnote{\cite{declare:2017} also included two heuristic clusters based on
+         atmosphere bias~\cite{Woodworth1935} and matching bias~\cite{Wetherick1995}.}
+This approach achieved a match of 92\% with the aggregated data reported in~\cite{khemlani:2012}, and was the highest result compared to the presented approaches in~\cite{khemlani:2012}.
+
+We were interested whether these clusters were also cognitively adequate in the sense that they 
+would also correspond to the reasoning patterns of individuals.
+Therefore, we investigated the non-aggregated dataset provided by~\cite{khemlani:2016} consisting of the responses of 20 participants for two weeks,
+i.e. the participants carried out the task twice, with a break of one week. The results with respect to WCS clustering is two-fold:
 \begin{description}
@@
  the majority's conclusion for 63 out of 64 pair of syllogistic premises and in the second week
- all majority's conclusions were predicted.
-\item[Non-aggregated evaluation] None of the clusters fitted any individual participants' response pattern, i.e.\ none of the participants belonged to a certain cluster.
+ all majority's conclusions were predicted by WCS clustering.
+\item[Non-aggregated evaluation] None of the clusters fitted any individual participants' response pattern, i.e.\ no participant did belong to a single cluster.
 \end{description}
-
-\clearpage
-\begin{table}[t]\betweentinyandsmall\centering
-\newcolumntype{a}{>{\columncolor{white}}c@{\hspace{2mm}}}
-\newcolumntype{b}{>{\columncolor{gray!40}}c@{\hspace{2mm}}}  
-\captionsetup{font=footnotesize}
- \begin{tabular}{llllllllll}
- \toprule
-	 & \MA ac & \ME ac & \MI ac & \MO ac & \MA ca & \ME ca & \MI ca & \MO ca & \NVC \\ \midrule
-
-\rowcolor{lightgray} 
-\MA\MA1  & \basicc&        &        &        &        &        &        &        &      \\ 
-\MA\MA2  &        &        &        &        & \basicc&        &        &        & \\ \rowcolor{lightgray} 
-\MA\MA3  &        &        &        &        &        &        &        &        & \basicc \\
-\MA\MA4  & \basicc&        &        &        &\basicc &        &        &        & \\  \rowcolor{lightgray}  %\midrule
-
-\MA\MI1  &        &        & \abdc  &        &        &        & \abdc  &        & \basicc \\
-\MA\MI2  &        &        & \basicc&        &        &        & \basicc \\  \rowcolor{lightgray} 
-\MA\MI3  &        &        & \abdc  &        &        &        & \abdc  &        & \basicc \\
-\MA\MI4  &        &        & \basicc&        &        &        & \basicc&        & \\%\midrule
-
-\rowcolor{lightgray} 
-\MA\ME1  &        & \basicc&        &        &        &        &        &        & \\ %\rowcolor{lightgray} 
-\MA\ME2  &        &\contrac&        &        &        & \basicc&        &        & \\ \rowcolor{lightgray} 
-\MA\ME3  &        &\contrac&        &        &        &\contrac&        &        & \basicc \\ %\rowcolor{lightgray} 
-\MA\ME4  &        & \basicc&        &        &        &\contrac&        &        &  \\  \rowcolor{lightgray}  %\midrule
-
-\MA\MO1  &        &        &        &\abdc   &        &        &        &        & \basicc \\
-\MA\MO2  &        &        &        &        &        &        &        &\basicc \\  \rowcolor{lightgray} 	
-\MA\MO3  &        &        &        &        &        &        &        &\contrac&\basicc \\
-\MA\MO4  &        &        &        &\basicc \\ %\midrule
-
-\rowcolor{lightgray} 
-\MI\MA1  &        &        &\basicc &        &        &        &\basicc &        & \\ % \rowcolor{lightgray} 
-\MI\MA2  &        &        &\abdc   &        &\genc   &        & \abdc  &        & \basicc \\ \rowcolor{lightgray} 
-\MI\MA3  &        &        & \abdc  &        &        &        & \abdc  &        & \basicc \\ % \rowcolor{lightgray} 
-\MI\MA4  &        &        &\basicc &        &        &        &\basicc &        &  \\  %\midrule
- \rowcolor{lightgray} 
-\MI\MI1  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\
-\MI\MI2  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\ \rowcolor{lightgray} 
-\MI\MI3  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\
-\MI\MI4  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\ %\midrule
-
-\rowcolor{lightgray} 
-\MI\ME1  &        &        &        &\basicc &        &        &        &        & \\ %\rowcolor{lightgray} 
-\MI\ME2  &        &        &        &\contrac&        & \genc  &        &        & \basicc \\\rowcolor{lightgray} 
-\MI\ME3  &        &        &        &\contrac&        & \genc  &        &        & \basicc  \\% \rowcolor{lightgray} 
-\MI\ME4  &        &        &        & \basicc&        &        &        &        & \\ %\midrule 
- \rowcolor{lightgray} 
-\MI\MO1  &        &        &        & \genc  &        &        &        &        & \basicc \\
-\MI\MO2  &        &        &        & \genc  &        &        &        &        & \basicc \\  \rowcolor{lightgray} 
-\MI\MO3  &        &        &        & \genc  &        &        &        &        & \basicc \\ 
-\MI\MO4  &        &        &        & \genc  &        &        &        &        & \basicc \\%\midrule
-
-\rowcolor{lightgray} 
-\ME\MA1  &        & \basicc&        &        &        &        &        &        & \\ %\rowcolor{lightgray} 
-\ME\MA2  &        &\contrac&        &        &        & \basicc&        &        &  \\\rowcolor{lightgray} 
-\ME\MA3  &        &\contrac&        &        &        &\contrac&        &        & \basicc \\% \rowcolor{lightgray} 
-\ME\MA4  &        &\contrac&        &        &        & \basicc&        &        &\\  %\midrule 
- \rowcolor{lightgray} 
-\ME\MI1  &        & \genc   &        &	 &	  &	   &	    &	     & \basicc \\
-\ME\MI2  &        &	       &	&	 &	  &	   &	    & \basicc \\ \rowcolor{lightgray}  
-\ME\MI3  &        & \genc   &        &	 &	  &	   &	    &\contrac& \basicc \\
-\ME\MI4  &        &	       &	&	 &	  &	   &	    & \basicc \\%\midrule
-
-\rowcolor{lightgray} 
-\ME\ME1  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
-\ME\ME2  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ \rowcolor{lightgray} 
-\ME\ME3  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
-\ME\ME4  &        &	       &	&	 &	  &	   &	    &        &\basicc \\%\midrule
- \rowcolor{lightgray} 
-\ME\MO1  &        &	       &	&	 &	  &	   &	    &        &\basicc \\
-\ME\MO2  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ \rowcolor{lightgray} 
-\ME\MO3  &        &	       &	&	 &	  &	   &	    &        &\basicc \\
-\ME\MO4  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ %\midrule
-
-\rowcolor{lightgray} 
-\MO\MA1  &        &            &	& \basicc&        &        &        &        & \\ %\rowcolor{lightgray} 
-\MO\MA2  &        &	       &	&	 &	  &	   &	    &  \abdc &\basicc \\\rowcolor{lightgray} 
-\MO\MA3  &        &	       &	&\contrac&	  &	   &	    &        &\basicc \\% \rowcolor{lightgray} 
-\MO\MA4  &        &	       &	&	 &	  &	   &	    & \basicc& \\  %\midrule
- \rowcolor{lightgray} 
-\MO\MI1  &        &	       &	&\genc   &	  &	   &	    &        &\basicc \\
-\MO\MI2  &        &	       &	&	 &	  &	   &	    &\genc   &\basicc \\  \rowcolor{lightgray} 
-\MO\MI3  &        &	       &	&\genc   &	  &	   &	    &        &\basicc \\ 
-\MO\MI4  &        &	       &	&	 &	  &	   &	    &\genc   &\basicc\\ %\midrule
-
-\rowcolor{lightgray} 
-\MO\ME1  &        &	       &	&        &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
-\MO\ME2  &        &	       &	&        &	  &	   &	    &        &\basicc \\\rowcolor{lightgray} 
-\MO\ME3  &        &	       &	&        &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
-\MO\ME4  &        &	       &	&        &	  &	   &	    &        &\basicc \\ %\midrule
-\rowcolor{lightgray} 
-\MO\MO1  &        &	       &	&        &	  &	   &	    &        &\basicc \\
-\MO\MO2  &        &	       &	&        &	  &	   &	    &        &\basicc \\\rowcolor{lightgray}  
-\MO\MO3  &        &	       &	&        &	  &	   &	    &        &\basicc \\
-\MO\MO4  &        &	       &	&        &	  &	   &	    &        &\basicc \\ \bottomrule
-  \end{tabular}
-
-\caption{\label{tab:cogprinc}Predictions according to the cognitive principles as specified in Section~\ref{sect:representing}.\betweentinyandsmall}
- \end{table}
-
-\thispagestyle{empty}
-\clearpage
-
-Given this sobering outcome with respect to the non-aggregated evaluation, 
-we investigated whether patterns could be found on the level of the
+Given this sobering outcome of (2), we investigated whether patterns could be found on the level of the
 cognitive principles: \basic, \contraposition, \abduction\ and \negFailure.
@@
 (iii) Do participants learn, i.e.\ do they improve their logical ability by doing the same reasoning task again.
-We address these questions in the following.
-
-\subsection{On non-aggregated Data} \label{sect:nonag}
+We address these questions in the following section.
+
+\clearpage
+\begin{table}[t]\betweentinyandsmall\centering
+\newcolumntype{a}{>{\columncolor{white}}c@{\hspace{2mm}}}
+\newcolumntype{b}{>{\columncolor{gray!40}}c@{\hspace{2mm}}}  
+\captionsetup{font=footnotesize}
+ \begin{tabular}{llllllllll}
+ \toprule
+	 & \MA ac & \ME ac & \MI ac & \MO ac & \MA ca & \ME ca & \MI ca & \MO ca & \NVC \\ \midrule
+
+\rowcolor{lightgray} 
+\MA\MA1  & \basicc&        &        &        &        &        &        &        &      \\ 
+\MA\MA2  &        &        &        &        & \basicc&        &        &        & \\ \rowcolor{lightgray} 
+\MA\MA3  &        &        &        &        &        &        &        &        & \basicc \\
+\MA\MA4  & \basicc&        &        &        &\basicc &        &        &        & \\  \rowcolor{lightgray}  %\midrule
+
+\MA\MI1  &        &        & \abdc  &        &        &        & \abdc  &        & \basicc \\
+\MA\MI2  &        &        & \basicc&        &        &        & \basicc \\  \rowcolor{lightgray} 
+\MA\MI3  &        &        & \abdc  &        &        &        & \abdc  &        & \basicc \\
+\MA\MI4  &        &        & \basicc&        &        &        & \basicc&        & \\%\midrule
+
+\rowcolor{lightgray} 
+\MA\ME1  &        & \basicc&        &        &        &        &        &        & \\ %\rowcolor{lightgray} 
+\MA\ME2  &        &\contrac&        &        &        & \basicc&        &        & \\ \rowcolor{lightgray} 
+\MA\ME3  &        &\contrac&        &        &        &\contrac&        &        & \basicc \\ %\rowcolor{lightgray} 
+\MA\ME4  &        & \basicc&        &        &        &\contrac&        &        &  \\  \rowcolor{lightgray}  %\midrule
+
+\MA\MO1  &        &        &        &\abdc   &        &        &        &        & \basicc \\
+\MA\MO2  &        &        &        &        &        &        &        &\basicc \\  \rowcolor{lightgray} 	
+\MA\MO3  &        &        &        &        &        &        &\contrac&\basicc & \\
+\MA\MO4  &        &        &        &\basicc \\ %\midrule
+
+\rowcolor{lightgray} 
+\MI\MA1  &        &        &\basicc &        &        &        &\basicc &        & \\ % \rowcolor{lightgray} 
+\MI\MA2  &        &        &\abdc   &        &\genc   &        & \abdc  &        & \basicc \\ \rowcolor{lightgray} 
+\MI\MA3  &        &        & \abdc  &        &        &        & \abdc  &        & \basicc \\ % \rowcolor{lightgray} 
+\MI\MA4  &        &        &\basicc &        &        &        &\basicc &        &  \\  %\midrule
+ \rowcolor{lightgray} 
+\MI\MI1  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\
+\MI\MI2  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\ \rowcolor{lightgray} 
+\MI\MI3  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\
+\MI\MI4  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\ %\midrule
+
+\rowcolor{lightgray} 
+\MI\ME1  &        &        &        &\basicc &        &        &        &        & \\ %\rowcolor{lightgray} 
+\MI\ME2  &        &        &        &\contrac&        & \genc  &        &        & \basicc \\\rowcolor{lightgray} 
+\MI\ME3  &        &        &        &\contrac&        & \genc  &        &        & \basicc  \\% \rowcolor{lightgray} 
+\MI\ME4  &        &        &        & \basicc&        &        &        &        & \\ %\midrule 
+ \rowcolor{lightgray} 
+\MI\MO1  &        &        &        & \genc  &        &        &        &        & \basicc \\
+\MI\MO2  &        &        &        & \genc  &        &        &        &        & \basicc \\  \rowcolor{lightgray} 
+\MI\MO3  &        &        &        & \genc  &        &        &        &        & \basicc \\ 
+\MI\MO4  &        &        &        & \genc  &        &        &        &        & \basicc \\%\midrule
+
+\rowcolor{lightgray} 
+\ME\MA1  &        & \basicc&        &        &        &        &        &        & \\ %\rowcolor{lightgray} 
+\ME\MA2  &        &\contrac&        &        &        & \basicc&        &        &  \\\rowcolor{lightgray} 
+\ME\MA3  &        &\contrac&        &        &        &\contrac&        &        & \basicc \\% \rowcolor{lightgray} 
+\ME\MA4  &        &\contrac&        &        &        & \basicc&        &        &\\  %\midrule 
+ \rowcolor{lightgray} 
+\ME\MI1  &        & \genc   &        &	 &	  &	   &	    &	     & \basicc \\
+\ME\MI2  &        &	       &	&	 &	  &	   &	    & \basicc \\ \rowcolor{lightgray}  
+\ME\MI3  &        & \genc   &        &	 &	  &	   &	    &\contrac& \basicc \\
+\ME\MI4  &        &	       &	&	 &	  &	   &	    & \basicc \\%\midrule
+
+\rowcolor{lightgray} 
+\ME\ME1  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
+\ME\ME2  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ \rowcolor{lightgray} 
+\ME\ME3  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
+\ME\ME4  &        &	       &	&	 &	  &	   &	    &        &\basicc \\%\midrule
+ \rowcolor{lightgray} 
+\ME\MO1  &        &	       &	&	 &	  &	   &	    &        &\basicc \\
+\ME\MO2  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ \rowcolor{lightgray} 
+\ME\MO3  &        &	       &	&	 &	  &	   &	    &        &\basicc \\
+\ME\MO4  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ %\midrule
+
+\rowcolor{lightgray} 
+\MO\MA1  &        &            &	& \basicc&        &        &        &        & \\ %\rowcolor{lightgray} 
+\MO\MA2  &        &	       &	&	 &	  &	   &	    &  \abdc &\basicc \\\rowcolor{lightgray} 
+\MO\MA3  &        &	       &	&\contrac&	  &	   &	    &        &\basicc \\% \rowcolor{lightgray} 
+\MO\MA4  &        &	       &	&	 &	  &	   &	    & \basicc& \\  %\midrule
+ \rowcolor{lightgray} 
+\MO\MI1  &        &	       &	&\genc   &	  &	   &	    &        &\basicc \\
+\MO\MI2  &        &	       &	&	 &	  &	   &	    &\genc   &\basicc \\  \rowcolor{lightgray} 
+\MO\MI3  &        &	       &	&\genc   &	  &	   &	    &        &\basicc \\ 
+\MO\MI4  &        &	       &	&	 &	  &	   &	    &\genc   &\basicc\\ %\midrule
+
+\rowcolor{lightgray} 
+\MO\ME1  &        &	       &	&        &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
+\MO\ME2  &        &	       &	&        &	  &	   &	    &        &\basicc \\\rowcolor{lightgray} 
+\MO\ME3  &        &	       &	&        &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
+\MO\ME4  &        &	       &	&        &	  &	   &	    &        &\basicc \\ %\midrule
+\rowcolor{lightgray} 
+\MO\MO1  &        &	       &	&        &	  &	   &	    &        &\basicc \\
+\MO\MO2  &        &	       &	&        &	  &	   &	    &        &\basicc \\\rowcolor{lightgray}  
+\MO\MO3  &        &	       &	&        &	  &	   &	    &        &\basicc \\
+\MO\MO4  &        &	       &	&        &	  &	   &	    &        &\basicc \\ \bottomrule
+  \end{tabular}
+
+\caption{\label{tab:cogprinc}Predictions according to the cognitive principles as specified in Section~\ref{sect:representing}.\betweentinyandsmall}
+ \end{table}
+
+\thispagestyle{empty}
+\clearpage
+
+\section{On non-aggregated Data} \label{sect:nonag}
 
@@
 
-\subsubsection{mReasoner}
+\subsection{mReasoner}
 
@@
 
-\subsubsection{Individual Reasoning Patterns}
+\subsection{Evaluation}
 
@@
 \end{tabular}
-\caption{Overview of the individual participants' responses with respect to the specified criteria in week 1 and week 2.
+\caption{Overview of each participants' performance with respect to the specified criteria in week 1 and week 2.
 $\uparrow$, $\downarrow$, and $\leftrightarrow$ denote above, below, and average, respectively.
@@
 The non-aggregated data provided by~\cite{khemlani:2016} included a dataset of 20 participants,
-for which the task was carried out twice, with a break of one week. On average, the basic principles explained the results about 10\% better than CL.
- Generally, the participants who were strong in CL/ contraposition, drew fewer conclusions according to abduction/ generalization, and vice versa.
+for which the task was carried twice, with a break of one week. On average, the basic principles explained the results on about 10\% better than CL.
+ Generally, the participants who were strong in CL/ contraposition,drew fewer conclusions according to abduction/ generalization, and vice versa.
  The changes within participant's answers from one to the other week was high, with 52/64 on average. 
@@
 Our classification only differs to~\cite{khemlani:2016} for the intermediate and intuitive cluster with respect to three participants (1, 11 and 19).
-A possible explanation for this difference, could be that~\cite{khemlani:2016}
+A natural explanation for this difference, could be that~\cite{khemlani:2016}
 did not take into account the participants' changes between week 1 and week 2.
@@
 \begin{description}
- \item[Deliberative cluster] (6 out of 20 participants) is characterized by the participants whose answers corresponded to the predictions made by basic/ contraposition.
+ \item[Deliberative cluster] (6/20) is characterized by the participants whose answers corresponded to the predictions made by basic/ contraposition.
 Almost none of their conclusions corresponded to predictions made by abduction/ generalization.
@@
  lower than average and concluded higher than average NVC. Generally, their answers improved from week 1 to week 2 with respect to CL.
- \item[Intuitive cluster] (7 out of 20 participants)
+ \item[Intuitive cluster] (7/20)
 is characterized by the participants whose answers corresponded to the predictions given by abduction/ generalization.
@@
  the deliberative cluster.
- \item[Intermediate cluster] (7 out of 20 participants) is characterized by the participants whose answers did not corresponded to any particular principle.
+ \item[Intermediate cluster] (7/20) is characterized by the participants whose answers did not corresponded to any particular principle.
 These participants had a high change within their own answers from week 1 to week 2
- and no particular pattern with respect to NVC conclusions.
-Generally they were not improving their answers from week 1 to week 2 with respect to CL, but normally compensated for answers valid in CL in some cases (that corresponded to contraposition) with answers not valid in CL in the other cases (that correspond to abduction/ generalization).
+ and no particular pattern with respect to NVC conclusions. Generally they were not improving their answers from week 1 to week 2 with respect to CL, but normally compensated for 
+ right answers in some cases (that corresponded to contraposition) with wrong answers in the other cases (that correspond to abduction/ generalization).
 \end{description}
@@
 
-We have investigated the cognitive adequacy of clustering by principles under WCS with respect to the individual answer patterns reported by~\cite{khemlani:2016}.
+We have investigated the cognitive adequacy of WCS clustering with respect to the individual answer patterns reported by~\cite{khemlani:2016}.
 \cite{khemlani:2016} is the only approach that we are aware of, that has evaluated their predictions on human syllogistic reasoning
@@
 Similarly, a meta-analysis of the individual response patterns for the selection task has been suggested in~\cite{ragni:2017}.
-This in turn might give us a new starting point to investigate whether the cognitive principles  specified here can characterize participants' conclusions in other reasoning episodes.
+This in turn might give us a new starting point to investigate whether the cognitive principles  specified here can characterize participants' conclusions in other reasoning tasks.
 % episodes.

### 2019-11-08 paper/paper-v3.tex
rewrote first part of intro

@@
 various experimental studies have shown that human reasoning systematically deviates 
-from conclusions valid under CL (cf.~\cite{wason:68,byrne:89}).
-Originally, one of the objectives in the field of AI and logic programming was to formalize human reasoning and commonsense reasoning~\cite{McCarthy59,McCarthy98}. 
-However, the proposed approaches have been strongly critized~(cf.~\cite{mcdermott:87}).
-On the other hand, cognitive scientists have extensively investigated and acquired a great expertise
-in human thinking. Recently, several attempts have been made to bridge the gap among these disciplines. 
-% as they were mostly theoretical and did not apply to real case studies.
+from conclusions that are valid under CL (cf.~\cite{wason:68,byrne:89}).
+% Originally, one of the objectives in the field of AI and logic programming was to formalize human reasoning and commonsense reasoning~\cite{McCarthy59,McCarthy98}. 
+% However, the proposed approaches have been strongly critized~(cf.~\cite{mcdermott:87}).
+% On the other hand, cognitive scientists have extensively investigated and acquired a great expertise
+% in human thinking. Recently, several attempts have been made to bridge the gap among these disciplines. 
+% % as they were mostly theoretical and did not apply to real case studies.
 % However, only recently an attempt was made 
 % this expertise was not consulted, i.e. there was no communication among the relevant fields. 
-
 Cognitive theories that try to explain human reasoning need to be evaluated by their cognitive adequacy~\cite{strube:1992} and their performance should be compared to other theories.
 Conceptual cognitive adequacy describes to what extent a system corresponds to human conceptual knowledge, whereas inferential cognitive adequacy indicates whether the reasoning process of a system is structured similarly to the way humans reason~\cite{knauff:rauh:renz:1997}.
-
-Yet, the cognitive adequacy of a theory can only be measured, if that theory is evaluable, i.e.\ if the theory can uniformally account for episodes of human reasoning systematically, according to its underlying assumptions.
-% Only then, the cognitve adequacy of that theory can be compared to others.
-
-Formal methods can offer the required tools to implement such cognitive theories.
-This in turn, can allow a new way of communication among the fields, by avoiding ambiguous definitions, showing their formal correspondences and finally providing appropriate evaluation methods with respect to their cognitive adequacy.
 In order to determine the required factors that are necessary for the evaluation of an adequate cognitive theory, we take~\cite{ragni:stolzenburg:2015} as a starting point:
@@
 
-The results of the meta-study on syllogistic reasoning in~\cite{khemlani:2012} show that the answers given by humans does not only systematically differ from the valid conclusions under classical logic, but from any other proposed cognitive theory (such as the theories proposed in~\cite{BJ89,johnsonlaird:1983,Rips1994,Polk1995,chater:oaksford:1999}).
+% Yet, the cognitive adequacy of a theory can only be measured, if that theory is evaluable, i.e.\ if the theory can uniformally account for episodes of human reasoning systematically, according to its underlying assumptions.
+% Only then, the cognitve adequacy of that theory can be compared to others.
+% 
+% Formal methods can offer the required tools to implement such cognitive theories.
+This in turn, can allow a new way of communication among the fields, by avoiding ambiguous definitions, showing their formal correspondences and finally providing appropriate evaluation methods with respect to their cognitive adequacy.
+% A major drawback of 
+The results of the meta-study on syllogistic reasoning in~\cite{khemlani:2012} show that the answers given by humans doe not only systematically differ from the valid conclusions under CL, but from any other proposed cognitive theory (cf.~\cite{BJ89,johnsonlaird:1983,Rips1994,Polk1995,chater:oaksford:1999}).
+% Possibly one of the reasons could be that none of the suggested cognitive theories so far has accounted for differences among individuals, i.e.\ they only considered the data summed up with respect to the individual answer possibilities.
+% As recent investigations on psychological experiments show (cf.~\cite{khemlani:2016, ragni:2017}), it seems that different humans represent and reason about conditional information differently.
+% This implies that \textit{the} human reasoner does not exist, instead there 
+% seem to exist various (human) reasoning clusters.
+
 % The approach suggested in~\cite{stolzenburg:luederitz:2017}, is based on venn diagrams, where moods are understood as conditional probabilities, and implemented 
@@
 	\[
-		\begin{array}{@{\hspace{0cm}}r@{\hspace{0.1cm}}c@{\hspace{0.1cm}}c@{\hspace{0.1cm}}c@{\hspace{0.1cm}}c}
+		\begin{array}{@{\hspace{0cm}}c@{\hspace{0.2cm}}c@{\hspace{0.1cm}}c@{\hspace{0.1cm}}c@{\hspace{0.1cm}}c}
 		\toprule
@@
 of~$\CalP$ or~$\WC\CalP$.
-
 % $\neg A$ is \textit{assumed in $\Ground \CalP$} iff $\Ground \CalP$
@@
 
-\subsubsection{Converse Implication (\converse) and Unknown Generalization (\unknownGen)}
+\subsubsection{Converse Implication (\converse) and \\ Unknown Generalization (\unknownGen)}
 Consider the statement \textit{Some y are z.} ({\MI yz}), which in CL, is represented as $\exists X (y(X) \wedge z(X))$
@@
 \end{description}
-For \MI yz (some) the conversion is explicitly specified, motivated by it (\converse) principle.
+For \MI yz (some) the conversion is explicitly specified, motivated by the (\converse) principle.
 By requiring for \MI yz and \MO yz, the existence of an object with the first property,
-for which the second property has to be either either false or unknown, \MA yz and \MI yz, and, \ME yz and \MO yz, exclude each other, respectively.
+for which the second property has to be either either false or unknown, \MA yz and \MI yz, (\ME yz and \MO yz, resp.), exclude each other.
 
@@
 b(o_2) & \leftarrow  \top. & \tag{\unknownGen}
-% \medskip \\
-\end{align}
-\begin{align}
+\medskip \\
 b(X) & \leftarrow a(X) \wedge \neg \Ab_{ab}(X). & \tag{\conditionals\ \&\ \licenses\ \&\ \converse}\\
@@
 
-\clearpage
-\begin{table}[t]\betweentinyandsmall\centering
-\newcolumntype{a}{>{\columncolor{white}}c@{\hspace{2mm}}}
-\newcolumntype{b}{>{\columncolor{gray!40}}c@{\hspace{2mm}}}  
-\captionsetup{font=footnotesize}
- \begin{tabular}{llllllllll}
- \toprule
-	 & \MA ac & \ME ac & \MI ac & \MO ac & \MA ca & \ME ca & \MI ca & \MO ca & \NVC \\ \midrule
-
-\rowcolor{lightgray} 
-\MA\MA1  & \basicc&        &        &        &        &        &        &        &      \\ 
-\MA\MA2  &        &        &        &        & \basicc&        &        &        & \\ \rowcolor{lightgray} 
-\MA\MA3  &        &        &        &        &        &        &        &        & \basicc \\
-\MA\MA4  & \basicc&        &        &        &\basicc &        &        &        & \\  \rowcolor{lightgray}  %\midrule
-
-\MA\MI1  &        &        & \abdc  &        &        &        & \abdc  &        & \basicc \\
-\MA\MI2  &        &        & \basicc&        &        &        & \basicc \\  \rowcolor{lightgray} 
-\MA\MI3  &        &        & \abdc  &        &        &        & \abdc  &        & \basicc \\
-\MA\MI4  &        &        & \basicc&        &        &        & \basicc&        & \\%\midrule
-
-\rowcolor{lightgray} 
-\MA\ME1  &        & \basicc&        &        &        &        &        &        & \\ %\rowcolor{lightgray} 
-\MA\ME2  &        &\contrac&        &        &        & \basicc&        &        & \\ \rowcolor{lightgray} 
-\MA\ME3  &        &\contrac&        &        &        &\contrac&        &        & \basicc \\ %\rowcolor{lightgray} 
-\MA\ME4  &        & \basicc&        &        &        &\contrac&        &        &  \\  \rowcolor{lightgray}  %\midrule
-
-\MA\MO1  &        &        &        &\abdc   &        &        &        &        & \basicc \\
-\MA\MO2  &        &        &        &        &        &        &        &\basicc \\  \rowcolor{lightgray} 	
-\MA\MO3  &        &        &        &        &        &        &        &\contrac&\basicc \\
-\MA\MO4  &        &        &        &\basicc \\ %\midrule
-
-\rowcolor{lightgray} 
-\MI\MA1  &        &        &\basicc &        &        &        &\basicc &        & \\ % \rowcolor{lightgray} 
-\MI\MA2  &        &        &\abdc   &        &\genc   &        & \abdc  &        & \basicc \\ \rowcolor{lightgray} 
-\MI\MA3  &        &        & \abdc  &        &        &        & \abdc  &        & \basicc \\ % \rowcolor{lightgray} 
-\MI\MA4  &        &        &\basicc &        &        &        &\basicc &        &  \\  %\midrule
- \rowcolor{lightgray} 
-\MI\MI1  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\
-\MI\MI2  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\ \rowcolor{lightgray} 
-\MI\MI3  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\
-\MI\MI4  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\ %\midrule
-
-\rowcolor{lightgray} 
-\MI\ME1  &        &        &        &\basicc &        &        &        &        & \\ %\rowcolor{lightgray} 
-\MI\ME2  &        &        &        &\contrac&        & \genc  &        &        & \basicc \\\rowcolor{lightgray} 
-\MI\ME3  &        &        &        &\contrac&        & \genc  &        &        & \basicc  \\% \rowcolor{lightgray} 
-\MI\ME4  &        &        &        & \basicc&        &        &        &        & \\ %\midrule 
- \rowcolor{lightgray} 
-\MI\MO1  &        &        &        & \genc  &        &        &        &        & \basicc \\
-\MI\MO2  &        &        &        & \genc  &        &        &        &        & \basicc \\  \rowcolor{lightgray} 
-\MI\MO3  &        &        &        & \genc  &        &        &        &        & \basicc \\ 
-\MI\MO4  &        &        &        & \genc  &        &        &        &        & \basicc \\%\midrule
-
-\rowcolor{lightgray} 
-\ME\MA1  &        & \basicc&        &        &        &        &        &        & \\ %\rowcolor{lightgray} 
-\ME\MA2  &        &\contrac&        &        &        & \basicc&        &        &  \\\rowcolor{lightgray} 
-\ME\MA3  &        &\contrac&        &        &        &\contrac&        &        & \basicc \\% \rowcolor{lightgray} 
-\ME\MA4  &        &\contrac&        &        &        & \basicc&        &        &\\  %\midrule 
- \rowcolor{lightgray} 
-\ME\MI1  &        & \genc   &        &	 &	  &	   &	    &	     & \basicc \\
-\ME\MI2  &        &	       &	&	 &	  &	   &	    & \basicc \\ \rowcolor{lightgray}  
-\ME\MI3  &        & \genc   &        &	 &	  &	   &	    &\contrac& \basicc \\
-\ME\MI4  &        &	       &	&	 &	  &	   &	    & \basicc \\%\midrule
-
-\rowcolor{lightgray} 
-\ME\ME1  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
-\ME\ME2  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ \rowcolor{lightgray} 
-\ME\ME3  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
-\ME\ME4  &        &	       &	&	 &	  &	   &	    &        &\basicc \\%\midrule
- \rowcolor{lightgray} 
-\ME\MO1  &        &	       &	&	 &	  &	   &	    &        &\basicc \\
-\ME\MO2  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ \rowcolor{lightgray} 
-\ME\MO3  &        &	       &	&	 &	  &	   &	    &        &\basicc \\
-\ME\MO4  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ %\midrule
-
-\rowcolor{lightgray} 
-\MO\MA1  &        &            &	& \basicc&        &        &        &        & \\ %\rowcolor{lightgray} 
-\MO\MA2  &        &	       &	&	 &	  &	   &	    &  \abdc &\basicc \\\rowcolor{lightgray} 
-\MO\MA3  &        &	       &	&\contrac&	  &	   &	    &        &\basicc \\% \rowcolor{lightgray} 
-\MO\MA4  &        &	       &	&	 &	  &	   &	    & \basicc& \\  %\midrule
- \rowcolor{lightgray} 
-\MO\MI1  &        &	       &	&\genc   &	  &	   &	    &        &\basicc \\
-\MO\MI2  &        &	       &	&	 &	  &	   &	    &\genc   &\basicc \\  \rowcolor{lightgray} 
-\MO\MI3  &        &	       &	&\genc   &	  &	   &	    &        &\basicc \\ 
-\MO\MI4  &        &	       &	&	 &	  &	   &	    &\genc   &\basicc\\ %\midrule
-
-\rowcolor{lightgray} 
-\MO\ME1  &        &	       &	&        &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
-\MO\ME2  &        &	       &	&        &	  &	   &	    &        &\basicc \\\rowcolor{lightgray} 
-\MO\ME3  &        &	       &	&        &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
-\MO\ME4  &        &	       &	&        &	  &	   &	    &        &\basicc \\ %\midrule
-\rowcolor{lightgray} 
-\MO\MO1  &        &	       &	&        &	  &	   &	    &        &\basicc \\
-\MO\MO2  &        &	       &	&        &	  &	   &	    &        &\basicc \\\rowcolor{lightgray}  
-\MO\MO3  &        &	       &	&        &	  &	   &	    &        &\basicc \\
-\MO\MO4  &        &	       &	&        &	  &	   &	    &        &\basicc \\ \bottomrule
-  \end{tabular}
-
-\caption{\label{tab:cogprinc}Predictions according to the cognitive principles as specified in Section~\ref{sect:representing}.\betweentinyandsmall}
- \end{table}
-
-\thispagestyle{empty}
-\clearpage
-
 Given this sobering outcome with respect to the non-aggregated evaluation, 
@@
 
+\clearpage
+\begin{table}[t]\betweentinyandsmall\centering
+\newcolumntype{a}{>{\columncolor{white}}c@{\hspace{2mm}}}
+\newcolumntype{b}{>{\columncolor{gray!40}}c@{\hspace{2mm}}}  
+\captionsetup{font=footnotesize}
+ \begin{tabular}{llllllllll}
+ \toprule
+	 & \MA ac & \ME ac & \MI ac & \MO ac & \MA ca & \ME ca & \MI ca & \MO ca & \NVC \\ \midrule
+
+\rowcolor{lightgray} 
+\MA\MA1  & \basicc&        &        &        &        &        &        &        &      \\ 
+\MA\MA2  &        &        &        &        & \basicc&        &        &        & \\ \rowcolor{lightgray} 
+\MA\MA3  &        &        &        &        &        &        &        &        & \basicc \\
+\MA\MA4  & \basicc&        &        &        &\basicc &        &        &        & \\  \rowcolor{lightgray}  %\midrule
+
+\MA\MI1  &        &        & \abdc  &        &        &        & \abdc  &        & \basicc \\
+\MA\MI2  &        &        & \basicc&        &        &        & \basicc \\  \rowcolor{lightgray} 
+\MA\MI3  &        &        & \abdc  &        &        &        & \abdc  &        & \basicc \\
+\MA\MI4  &        &        & \basicc&        &        &        & \basicc&        & \\%\midrule
+
+\rowcolor{lightgray} 
+\MA\ME1  &        & \basicc&        &        &        &        &        &        & \\ %\rowcolor{lightgray} 
+\MA\ME2  &        &\contrac&        &        &        & \basicc&        &        & \\ \rowcolor{lightgray} 
+\MA\ME3  &        &\contrac&        &        &        &\contrac&        &        & \basicc \\ %\rowcolor{lightgray} 
+\MA\ME4  &        & \basicc&        &        &        &\contrac&        &        &  \\  \rowcolor{lightgray}  %\midrule
+
+\MA\MO1  &        &        &        &\abdc   &        &        &        &        & \basicc \\
+\MA\MO2  &        &        &        &        &        &        &        &\basicc \\  \rowcolor{lightgray} 	
+\MA\MO3  &        &        &        &        &        &        &        &\contrac&\basicc \\
+\MA\MO4  &        &        &        &\basicc \\ %\midrule
+
+\rowcolor{lightgray} 
+\MI\MA1  &        &        &\basicc &        &        &        &\basicc &        & \\ % \rowcolor{lightgray} 
+\MI\MA2  &        &        &\abdc   &        &\genc   &        & \abdc  &        & \basicc \\ \rowcolor{lightgray} 
+\MI\MA3  &        &        & \abdc  &        &        &        & \abdc  &        & \basicc \\ % \rowcolor{lightgray} 
+\MI\MA4  &        &        &\basicc &        &        &        &\basicc &        &  \\  %\midrule
+ \rowcolor{lightgray} 
+\MI\MI1  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\
+\MI\MI2  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\ \rowcolor{lightgray} 
+\MI\MI3  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\
+\MI\MI4  &        &        &\genc   &        &        &        & \genc  &        & \basicc \\ %\midrule
+
+\rowcolor{lightgray} 
+\MI\ME1  &        &        &        &\basicc &        &        &        &        & \\ %\rowcolor{lightgray} 
+\MI\ME2  &        &        &        &\contrac&        & \genc  &        &        & \basicc \\\rowcolor{lightgray} 
+\MI\ME3  &        &        &        &\contrac&        & \genc  &        &        & \basicc  \\% \rowcolor{lightgray} 
+\MI\ME4  &        &        &        & \basicc&        &        &        &        & \\ %\midrule 
+ \rowcolor{lightgray} 
+\MI\MO1  &        &        &        & \genc  &        &        &        &        & \basicc \\
+\MI\MO2  &        &        &        & \genc  &        &        &        &        & \basicc \\  \rowcolor{lightgray} 
+\MI\MO3  &        &        &        & \genc  &        &        &        &        & \basicc \\ 
+\MI\MO4  &        &        &        & \genc  &        &        &        &        & \basicc \\%\midrule
+
+\rowcolor{lightgray} 
+\ME\MA1  &        & \basicc&        &        &        &        &        &        & \\ %\rowcolor{lightgray} 
+\ME\MA2  &        &\contrac&        &        &        & \basicc&        &        &  \\\rowcolor{lightgray} 
+\ME\MA3  &        &\contrac&        &        &        &\contrac&        &        & \basicc \\% \rowcolor{lightgray} 
+\ME\MA4  &        &\contrac&        &        &        & \basicc&        &        &\\  %\midrule 
+ \rowcolor{lightgray} 
+\ME\MI1  &        & \genc   &        &	 &	  &	   &	    &	     & \basicc \\
+\ME\MI2  &        &	       &	&	 &	  &	   &	    & \basicc \\ \rowcolor{lightgray}  
+\ME\MI3  &        & \genc   &        &	 &	  &	   &	    &\contrac& \basicc \\
+\ME\MI4  &        &	       &	&	 &	  &	   &	    & \basicc \\%\midrule
+
+\rowcolor{lightgray} 
+\ME\ME1  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
+\ME\ME2  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ \rowcolor{lightgray} 
+\ME\ME3  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
+\ME\ME4  &        &	       &	&	 &	  &	   &	    &        &\basicc \\%\midrule
+ \rowcolor{lightgray} 
+\ME\MO1  &        &	       &	&	 &	  &	   &	    &        &\basicc \\
+\ME\MO2  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ \rowcolor{lightgray} 
+\ME\MO3  &        &	       &	&	 &	  &	   &	    &        &\basicc \\
+\ME\MO4  &        &	       &	&	 &	  &	   &	    &        &\basicc \\ %\midrule
+
+\rowcolor{lightgray} 
+\MO\MA1  &        &            &	& \basicc&        &        &        &        & \\ %\rowcolor{lightgray} 
+\MO\MA2  &        &	       &	&	 &	  &	   &	    &  \abdc &\basicc \\\rowcolor{lightgray} 
+\MO\MA3  &        &	       &	&\contrac&	  &	   &	    &        &\basicc \\% \rowcolor{lightgray} 
+\MO\MA4  &        &	       &	&	 &	  &	   &	    & \basicc& \\  %\midrule
+ \rowcolor{lightgray} 
+\MO\MI1  &        &	       &	&\genc   &	  &	   &	    &        &\basicc \\
+\MO\MI2  &        &	       &	&	 &	  &	   &	    &\genc   &\basicc \\  \rowcolor{lightgray} 
+\MO\MI3  &        &	       &	&\genc   &	  &	   &	    &        &\basicc \\ 
+\MO\MI4  &        &	       &	&	 &	  &	   &	    &\genc   &\basicc\\ %\midrule
+
+\rowcolor{lightgray} 
+\MO\ME1  &        &	       &	&        &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
+\MO\ME2  &        &	       &	&        &	  &	   &	    &        &\basicc \\\rowcolor{lightgray} 
+\MO\ME3  &        &	       &	&        &	  &	   &	    &        &\basicc \\ %\rowcolor{lightgray} 
+\MO\ME4  &        &	       &	&        &	  &	   &	    &        &\basicc \\ %\midrule
+\rowcolor{lightgray} 
+\MO\MO1  &        &	       &	&        &	  &	   &	    &        &\basicc \\
+\MO\MO2  &        &	       &	&        &	  &	   &	    &        &\basicc \\\rowcolor{lightgray}  
+\MO\MO3  &        &	       &	&        &	  &	   &	    &        &\basicc \\
+\MO\MO4  &        &	       &	&        &	  &	   &	    &        &\basicc \\ \bottomrule
+  \end{tabular}
+
+\caption{\label{tab:cogprinc}Predictions according to the cognitive principles as specified in Section~\ref{sect:representing}.\betweentinyandsmall}
+ \end{table}
+
+\thispagestyle{empty}
+\clearpage
 \subsection{On non-aggregated Data} \label{sect:nonag}
@@
 
-Another aspect that would be intersting to investigate is in how far the semantics of the $\cxt$ operator within the extended three-valued {\L}ukasiewicz logic in Section~\ref{sub:3valuedL}, 
+Another interesting aspect is to investigate in how far the semantics of the $\cxt$ operator within the extended three-valued {\L}ukasiewicz logic in Section~\ref{sub:3valuedL}, 
  corresponds to the {\L}ukasiewicz extensional necessity operator~\cite{}.\footnote{Thanks to the

### 2019-11-10 paper/paper-v3.tex
Improvements/suggestions

@@
 
+\usepackage[backgroundcolor=orange!5,bordercolor=orange,linecolor=orange,textsize=footnotesize]{todonotes}
+
 \input{abbreviations}
@@
 % A major drawback of 
-The results of the meta-study on syllogistic reasoning in~\cite{khemlani:2012} show that the answers given by humans doe not only systematically differ from the valid conclusions under CL, but from any other proposed cognitive theory (cf.~\cite{BJ89,johnsonlaird:1983,Rips1994,Polk1995,chater:oaksford:1999}).
+The results of the meta-study on syllogistic reasoning in~\cite{khemlani:2012} show that the answers given by humans do not only systematically differ from the valid conclusions under CL, but from any other proposed cognitive theory (cf.~\cite{BJ89,johnsonlaird:1983,Rips1994,Polk1995,chater:oaksford:1999}).
 % Possibly one of the reasons could be that none of the suggested cognitive theories so far has accounted for differences among individuals, i.e.\ they only considered the data summed up with respect to the individual answer possibilities.
@@
 A syllogism consists of two
-premises and one conclusion. 
-The assertions are statements about properties over terms and can have one of the classical quantifiers in Table~\ref{table:moods}, which are called moods.
-The pair of syllogistic premises can have four different orders as shown in Table~\ref{t:2}.
-Based on these moods and orders, 64 different pairs of syllogistic premises can be constructed.
+premises and one conclusion,
+each of which is a statement about properties over terms which can have one of the classical quantifiess in Table~\ref{table:moods} (called \emph{moods}).
+The pair of syllogistic premises can have four different orders as shown in Table~\ref{t:2} (called \emph{figures}).
+Based on these moods and figures, 64 different pairs of syllogistic premises can be constructed.
 A conclusion then consists of an assertion about two properties over terms where each of these properties occurs in only one of the premises, together with a third middle property that does not appear in the conclusion but appears in both premises.
+\todo[inline]{R: Für mich ist der letzte Satz unverständlich.
+Vorschlag: ``\dots each of these properties occurs in exactly one of the premises, and there is a third property connecting both premises which does not occur in the conclusion.''}\noindent%
 \begin{table}[t]
@@
 \end{table}
-Consider the following pair of syllogistic premises: 
+For instance, consider the following pair of syllogistic premises: 
 \begin{align}
@@
 According to~\cite{khemlani:2012}, the majority of participants in experimental studies
-concluded \emph{No chemists are artists} (29\%) and \emph{No valid conclusion} (27\%). The valid conclusion \emph{Some artists are not chemists}
-under CL, was concluded by only 16\%, which is exactly the threshold, that the answer had not been chosen randomly~\cite{khemlani:2012}.
+concluded \emph{No chemists are artists} (29\;\%) and \emph{No valid conclusion} (27\;\%).
+The valid conclusion under CL, \emph{Some artists are not chemists},
+was concluded by only 16\,\%, which is exactly the threshold that the answer had not been chosen randomly~\cite{khemlani:2012}.
 % \footnote{
@@
 Section~\ref{sect:aggregateddata} illustrates an example and briefly discusses the performance of the approach in~\cite{declare:2017} on non-aggregated data.
-Section~\ref{sect:nonag} provides a more adequate characterization of the individual's reasoning patterns.
+Section~\ref{sect:nonag} provides a more adequate characterization of the individuals' reasoning patterns.
 
@@
 The approach suggested in~\cite{declare:2017} is based on a new computational logic approach for human syllogistic reasoning~\cite{cogsci:2017}, which seems to perform better than other state-of-the-art cognitive theories when evaluated with aggregated data provided in~\cite{khemlani:2012}.
-The approach presented in~\cite{declare:2017} extended~\cite{cogsci:2017} by specifying three reasoning clusters and two heuristic strategies and achieved a match of 92\%, which was the highest result compared to the presented approaches in~\cite{khemlani:2012}.
+The approach presented in~\cite{declare:2017} extended~\cite{cogsci:2017} by specifying three reasoning clusters and two heuristic strategies and achieved a match of 92\;\%, which was the highest result compared to the presented approaches in~\cite{khemlani:2012}.
 % The system developed in~\cite{declare:2017}, was extended and participated at the Syllogism Challenge 2017.\footnote{\url{https://www.cc.uni-freiburg.de/modelingchallenge/challenge-2017}} 
@@
 After that, this set was translated into a logic program (Section~\ref{sect:basic}-\ref{sect:advanced}), representing the group of reasoners.
-Next, the desired models were computed (Section~\ref{sect:wcs}-\ref{sect:lfp}).
+Next, the desired\todo{R: desired? Entailed, oder was willst du sagen?} models were computed (Section~\ref{sect:wcs}-\ref{sect:lfp}).
 After that, the entailed relations between $a$ and $c$ were extracted (Section~\ref{sect:entail}). 
@@
 \end{tikzpicture}
-\caption{The system that participated at the Syllogism Challenge 2017\label{overview:system:syllchal}}
+\caption{The system that participated at the Syllogism Challenge 2017\label{overview:system:syllchal}.}
 
@@
 is a finite set of contextual program clauses.
-As we restrict terms to be either constants or variables only, we consider so-called \defname{data logic programs}.
-If the clause contains variables, then they are implicitly universally quantified within the scope of the entire clause. 
+We restrict terms to be either constants or variables only, hence we consider so-called \defname{data logic programs}.
+If a clause contains variables, then they are implicitly universally quantified within the scope of the entire clause. 
 % For instance, the clause $p(X) \leftarrow q(X)$ represents the formula $(\forall X) (p(X) \leftarrow q(X))$. Therefore, a clause never contains free variables.
 A \defname{ground clause} is a clause that does not contain variables.
-$\Ground \CalP$ denotes the set of all its ground clauses.
-$A$ is \textit{defined in $\Ground \CalP$}  iff $\Ground \CalP$ contains a clause with head $A$.
+$\Ground \CalP$ denotes the set of all ground clauses of $\CalP$.
+An atom $A$ is \textit{defined in $\Ground \CalP$}  iff $\Ground \CalP$ contains a clause with head $A$.
 The \textit{definition of $A$ in $\Ground \CalP$} is defined as $\Def(A,\CalP)
@@
 
+\todo[inline]{R: Ist es eine Überlegung wert, \L ukasiewicz logic vor contextual programs einzuführen? Oben benutzen wir schon $\udf$, was erst hier eingeführt wird.}
 We consider the three-valued {\L}ukasiewicz logic~\cite{lukasiewicz:20}
@@
 \]
-\caption{The truth tables for the connectives under the three-valued {\L}ukasiewicz logic and for $\cxt$. $L$ is a literal, $\top$, $\bot$, and $\udf$ denote \textit{true}, \textit{false},
+\caption{The truth tables for the connectives under the three-valued {\L}ukasiewicz logic and for $\cxt$. $L$ is a literal and $\top$, $\bot$, and $\udf$ denote \textit{true}, \textit{false},
 and \textit{unknown}, respectively.\label{tab:3vldluka}}
@@
 the form $\udf \leftarrow \Body$, where $\Body$ is a conjunction of
-literals and $\udf$ denotes unknown. An interpretation maps
+literals\todo{R: ``and $\udf$ denotes unknown'' entfernt, sollte an dieser Stelle klar sein}. An interpretation maps
 an integrity constraint to $\top$ iff $\Body$ is either mapped to
@@
 \]
-The weak completion of non-contextual programs\footnote{Non-contextual programs refer programs that do not contain rules, in which $\cxt$ appears in the body of any rule, i.e.\ $p = 0$.} 
+The weak completion of non-contextual programs\footnote{Non-contextual programs refer programs that do not contain rules in which $\cxt$ appears in the body of any rule, i.e.\ $p = 0$.} 
 always has a least model under  {\L}ukasiewicz logic,
@@
 the least fixed point of $\Svlp$ always exists. However,
-if these programs do not contain cycles, then there always exists a supported model, which 
+if these programs do not contain cycles, then there always exists a supported model\todo{R: Begriff ``supported model'' unklar}, which 
 corresponds to the least fixed point of~$\Svlp$~\cite{dietz:hoelldobler:pereira:2017,dietz:hoelldobler:philipp:2017}.
@@
 First, we introduce the basic principles that are assumed for all premises when applicable.
-Thus, we assume them to be assumed by all (human) reasoners.
+Thus, we assume them to be assumed\todo{R: assumed $\to$ applied?} by all (human) reasoners.
 In Section~\ref{sect:advanced}, we present the advanced principles, which will 
@@
 Similarly, we assume that humans also understand this statement as a conditional sentence (\conditionals).
-However, it seems that humans understand them defeasibly, they assume a \textit{license for the inference}~\cite{stenning:vanlambalgen:2008},
+However, it seems that humans understand them defeasibly, \todo{R: i.e.?} they assume a \textit{license for the inference}~\cite{stenning:vanlambalgen:2008},
 i.e.\ they might allow exceptions to the conditional (\licenses). For (\MA yz) this can be 
 explicitly expressed by 
-\textit{all $y$ that are not abnormal, are $z$}.
+\textit{all $y$ that are not abnormal are $z$}.
 % \\
@@
 choice rules in~\cite{GebserKaminskiKaufmannSchaub12}).
-Another principle that we introduce is the no double negation principle: Under the Weak Completion Semantics, a positive conclusion can be derived from double negation appearing in two conditionals.
+Another principle that we introduce is the \emph{no double negation} principle: Under the Weak Completion Semantics, a positive conclusion can be derived from double negation appearing in two conditionals.
 However, it seems that humans normally do not reason in such a way~\cite{khemlani:2012}.
@@
 Accordingly, if we only observe that an object $o$ belongs to $y$ and $z$ then we do not want to conclude both, \textit{some y are z} and \textit{all y are z}.
-In order to distinguish between \textit{some y are z} and \textit{all y are z}, we introduce the following principle:  If we know that \textit{some y are z}, then there must not only be an object~$o_1$, which belongs to $y$ and $z$ but there must be another object $o_2$, which belongs to
+In order to distinguish between \textit{some y are z} and \textit{all y are z}, we introduce the following principle:  If we know that \textit{some y are z}, then there must not only be an object~$o_1$ which belongs to $y$ and $z$, but there must be another object $o_2$ which belongs to
 $y$ and which does not belong to~$z$.
@@
 
-The principles introduced here, are assumed not to be applied by all humans, 
-but characterize the reasoning clusters, specified in Section~\ref{sect:aggregateddata}.
+The principles introduced here are assumed not to be applied by all humans, 
+but characterize the reasoning clusters (c.f.\ Section~\ref{sect:aggregateddata}).
 
@@
 evidence that some of the participants might directly or indirectly 
-apply the contrapositive (\contraposition) within a reasoning task (cf.~\cite{rips:1994,mentallogic:1994}).
+apply the contrapositive (\contraposition) within a reasoning task (cf.~\cite{rips:1994,mentallogic:1994})\todo{R: Die Referenzen gibt es bei mir nicht}.
 Consider again $\CalP_{Ayz}$, the logic program representing (\MA yz)
@@
 $\Ab_{yz}(o_2) \leftarrow \udf$ guarantees that $\Ab_{yz}(o_2)$ stays unknown, even though $\cxt(z'(o_2))$ might be false.
-Recall, programs are considered under their weak completion, i.e.\ if $\cxt(z'(o_2))$ is false, then
+Recall that programs are considered under their weak completion, i.e.\ if $\cxt(z'(o_2))$ is false, then
 \[\Ab_{yz}(o_2) \leftrightarrow (\cxt(z'(o_2)) \vee \udf)
@@
 
-When participants are faced with a \NVC\ conclusion (\textit{No valid conclusion}), they might not want to accept this conclusion and proceed to check whether there exists unknown information that might be relevant (\abduction).
+When participants are faced with an \NVC\ conclusion (\textit{No valid conclusion}), they might not want to accept this conclusion and proceed to check whether there exists unknown information that might be relevant (\abduction).
 This information may be explanations about
@@
 \end{align}
-Classical logically \textit{No valid conclusion} follows. However, 62\% of the participants concluded \textit{Some artists are not chemists}.
+Under CL, \textit{No valid conclusion} follows. However, 62\;\% of the participants concluded \textit{Some artists are not chemists}.
 This conclusion can be modeled by abductive reasoning, where the system searches for explanations
@@
 $\CalP \not\ModelsWCS \neg z(o_2)$.
-\item[\NVC (no valid conclusion)] iff previous conclusion can be derived.
+\item[\NVC\ (no valid conclusion)] iff none of the previous conclusions can be derived.
 \end{description}
 For \MI yz (some) the conversion is explicitly specified, motivated by the (\converse) principle.
-By requiring for \MI yz and \MO yz, the existence of an object with the first property,
-for which the second property has to be either either false or unknown, \MA yz and \MI yz, (\ME yz and \MO yz, resp.), exclude each other.
+By requiring for \MI yz and \MO yz the existence of an object $o_1$ with the first property $y$,
+for which the second property $z$ has to be either either false or unknown,
+\todo[inline]{R: Das ist doch nicht richtig, oder? Bei \MI yz muss $z(o_2)$ entweder false oder unknown sein, wie beschrieben, aber bei \MO yz muss es entweder true oder unknown sein.}
+\MA yz and \MI yz, (\ME yz and \MO yz, resp.), exclude each other.
 
@@
 The majority of participants in experimental studies
-concluded \emph{No chemists are artists} (29\%), \emph{No valid conclusion} (27\%)
-and \emph{Some artists are not chemists} (16\%).
+concluded \emph{No chemists are artists} (29\;\%), \emph{No valid conclusion} (27\;\%)
+and \emph{Some artists are not chemists} (16\;\%).
 In order to account for the individual differences among participants, three clusters on principles and two on heuristic strategies have been suggested
@@
 }
-The three reasoning clusters have been characterized by the following principles:\footnote{{\footnotesize\textsf{basic}} refers to the set of basic principles introduced in Subsection~\ref{sect:basic}.} (1) \basic, \converse\ for \MI and \abduction,
-    (2) \basic, \converse\ for \MI\ and \negFailure, and
-    (3) \basic, \converse\ for \MI, \ME, and \contraposition\ for~\MA.
-This approach achieved a match of 92\% with the aggregated data reported in~\cite{khemlani:2012}, and was the highest result compared to the presented approaches in~\cite{khemlani:2012}.
+The three reasoning clusters have been characterized by the following principles:\footnote{{\footnotesize\textsf{basic}} refers to the set of basic principles introduced in Subsection~\ref{sect:basic}.} (1) \basic, \converse\ for \MI, and \abduction,
+    (2) \basic, \converse\ for \MI, and \negFailure, and
+    (3) \basic, \converse\ for \MI\ and \ME, and \contraposition\ for~\MA.
+This approach achieved a match of 92\;\% with the aggregated data reported in~\cite{khemlani:2012}, and was the highest result compared to the presented approaches in~\cite{khemlani:2012}.
 
@@
 
-On aggregated data, the clusters performed fairly well with a match of 92\%. 
+On aggregated data, the clusters performed fairly well with a match of 92\;\%. 
 We are interested whether these clusters also correspond to the reasoning patterns observed among individuals. 
@@
 cognitive principles: \basic, \contraposition, \abduction\ and \negFailure.
+\todo[inline]{R: Ich finde, hier wird nicht ganz klar, was das ``level of cognitive principles'' ist}
 Table~\ref{tab:cogprinc} provides an overview of the predicted conclusions according to these principles.\footnote{Note that any advanced principle also includes the logic program clauses specified for \textsf{basic}.}
@@
  by these principles? and
-(iii) Do participants learn, i.e.\ do they improve their logical ability by doing the same reasoning task again.
+(iii) Do participants learn, i.e.\ do they improve their logical ability by doing the same reasoning task again?
 We address these questions in the following.
@@
 Reasoning depends on the construction and manipulation of mental models, based on 3 fundamental principles: 
-(i) represent possibilities (ii) principle of iconicity and (iii) principle of dual processes.
- The procedure in the mReasoner consists of two steps:
+(i) represent possibilities, (ii) principle of iconicity, and (iii) principle of dual processes.
+ The procedure of mReasoner consists of two steps:
 (1) The construction of a small set of models that represents the terms of the premises, and
 (2) the call of a deliberative component to search for counterexamples to conclusions.
-They specified four parameters for the construction of a mental model and evaluated their
+They\todo{R: statt ``they'' \cite{khemlani:2016}} specified four parameters for the construction of a mental model and evaluated their
 predictions based on 2 datasets, week 1 and week 2, for 20 participants.
-The parameters for the construction of the mental model are (i) the 
-size of a mental model (ii) the probability of drawing the complete set 
-(iii) the probability to search for counter examples 
+The parameters for the construction of the mental models are (i) the 
+size of a mental model, (ii) the probability of drawing\todo{R: was heißt draw hier?} the complete set, 
+(iii) the probability to search for counter examples,
 and (iv) the probability to weaken the conclusion.
-The clustering with the mReasoner was based on partitioning around medoids clustering algorithm to determine the optimal number of clusters.
+The clustering with mReasoner was based on partitioning around\todo{R: around \emph{the} medoids?} medoids clustering algorithm to determine the optimal number of clusters.
 The K-means clustering algorithm was used to divide participants into clusters and then
@@
 The non-aggregated data provided by~\cite{khemlani:2016} included a dataset of 20 participants,
-for which the task was carried out twice, with a break of one week. On average, the basic principles explained the results about 10\% better than CL.
+for which the task was carried out twice, with a break of one week. On average, the basic principles explained the results about 10\;\% better than CL.
  Generally, the participants who were strong in CL/ contraposition, drew fewer conclusions according to abduction/ generalization, and vice versa.
  The changes within participant's answers from one to the other week was high, with 52/64 on average. 
- From week 1 to week 2, an increase with respect to CL (19/20, 6\%), basic (18/20, 6\%), contraposition (15/20, 5\%)
- and a decrease on generalization (15/20, 8\%) and no change on abduction (10/20, 0\%) answers could be observed.
+ From week 1 to week 2, an increase with respect to CL (19/20, 6\;\%), basic (18/20, 6\;\%), contraposition (15/20, 5\;\%)
+ and a decrease on generalization (15/20, 8\;\%) and no change on abduction (10/20, 0\;\%) answers could be observed.
 
@@
 We have investigated the cognitive adequacy of clustering by principles under WCS with respect to the individual answer patterns reported by~\cite{khemlani:2016}.
-\cite{khemlani:2016} is the only approach that we are aware of, that has evaluated their predictions on human syllogistic reasoning
+\cite{khemlani:2016} is the only approach we are aware of that has evaluated predictions on human syllogistic reasoning
 with respect to the non-aggregated data.
-However different than here, they did not distinguish between the participants' responses in week 1 and week 2.
+However in contrast to here, they did not distinguish between the participants' responses in week 1 and week 2.
 
@@
 
-Another interesting aspect is to investigate in how far the semantics of the $\cxt$ operator within the extended three-valued {\L}ukasiewicz logic in Section~\ref{sub:3valuedL}, 
- corresponds to the {\L}ukasiewicz extensional necessity operator~\cite{}.\footnote{Thanks to the
+Another interesting aspect is to investigate in how far the semantics of the $\cxt$ operator within the extended three-valued {\L}ukasiewicz logic in Section~\ref{sub:3valuedL} 
+ corresponds to the {\L}ukasiewicz extensional necessity operator~\cite{}\todo{R: Referenz fehlt}.\footnote{Thanks to the
  anonymous reviewer for this observation.}

### 2019-11-11 paper/paper-v3.tex
richards kommentare eingearbeitet

@@
 Based on these moods and figures, 64 different pairs of syllogistic premises can be constructed.
-A conclusion then consists of an assertion about two properties over terms where each of these properties occurs in only one of the premises, together with a third middle property that does not appear in the conclusion but appears in both premises.
-\todo[inline]{R: Für mich ist der letzte Satz unverständlich.
-Vorschlag: ``\dots each of these properties occurs in exactly one of the premises, and there is a third property connecting both premises which does not occur in the conclusion.''}\noindent%
+A conclusion then consists of an assertion about two properties over terms where each of these properties occurs in exactly one of the premises, and there is a third property connecting both premises which does not occur in the conclusion
 \begin{table}[t]
@@
 two premises, which conclusion on the relation between \emph{artists} and \emph{chemists} necessarily follows?
-According to~\cite{khemlani:2012}, the majority of participants in experimental studies
+According to the authors in~\cite{khemlani:2012}, the majority of participants in experimental studies
 concluded \emph{No chemists are artists} (29\;\%) and \emph{No valid conclusion} (27\;\%).
@@
 
-Here, we focus on a qualitative assessment of the clustering approach in~\cite{declare:2017},
-in the sense that we investigate the individual participants' conclusions with respect to the non-aggregated dataset provided by~\cite{khemlani:2016}.
+Here, we focus on a qualitative assessment of the clustering approach proposed in~\cite{declare:2017},
+in the sense that we investigate the individual participants' conclusions with respect to the non-aggregated dataset provided by the authors in~\cite{khemlani:2016}.
 In Section~\ref{sect:wcsclustering} and~\ref{sect:representing}, we introduce the underlying computational logic approach and the formal representation of cognitive principles within this approach.
-Section~\ref{sect:aggregateddata} illustrates an example and briefly discusses the performance of the approach in~\cite{declare:2017} on non-aggregated data.
+Section~\ref{sect:aggregateddata} illustrates an example and briefly discusses the performance of the approach presented in~\cite{declare:2017} on non-aggregated data.
 Section~\ref{sect:nonag} provides a more adequate characterization of the individuals' reasoning patterns.
@@
 
-According to~\cite{declare:2017}, reasoning clusters can be expressed through cognitive principles and heuristic strategies.
+Clustering by Principles under WCS first presented in~\cite{declare:2017},
+assumes that reasoning clusters can be expressed through cognitive principles and heuristic strategies.
 Cognitive principles are assumptions (not necessarily valid under CL) made by humans while reasoning,
@@
 conclusions are not derived by reasoning but instead based on structural patterns or {biases}.
-The approach suggested in~\cite{declare:2017} is based on a new computational logic approach for human syllogistic reasoning~\cite{cogsci:2017}, which seems to perform better than other state-of-the-art cognitive theories when evaluated with aggregated data provided in~\cite{khemlani:2012}.
-The approach presented in~\cite{declare:2017} extended~\cite{cogsci:2017} by specifying three reasoning clusters and two heuristic strategies and achieved a match of 92\;\%, which was the highest result compared to the presented approaches in~\cite{khemlani:2012}.
+This approach is an extension of the computational logic approach for human syllogistic reasoning presented in~\cite{cogsci:2017}: Three reasoning clusters and two heuristic strategies were specified, and achieved a match of 92\;\%, which was the highest result compared to the presented approaches in~\cite{khemlani:2012}.
 % The system developed in~\cite{declare:2017}, was extended and participated at the Syllogism Challenge 2017.\footnote{\url{https://www.cc.uni-freiburg.de/modelingchallenge/challenge-2017}} 
@@
 After that, this set was translated into a logic program (Section~\ref{sect:basic}-\ref{sect:advanced}), representing the group of reasoners.
-Next, the desired\todo{R: desired? Entailed, oder was willst du sagen?} models were computed (Section~\ref{sect:wcs}-\ref{sect:lfp}).
+Next, the respective models were computed (Section~\ref{sect:wcs}-\ref{sect:lfp}).
 After that, the entailed relations between $a$ and $c$ were extracted (Section~\ref{sect:entail}). 
@@
 
-\todo[inline]{R: Ist es eine Überlegung wert, \L ukasiewicz logic vor contextual programs einzuführen? Oben benutzen wir schon $\udf$, was erst hier eingeführt wird.}
 We consider the three-valued {\L}ukasiewicz logic~\cite{lukasiewicz:20}
@@
 the form $\udf \leftarrow \Body$, where $\Body$ is a conjunction of
-literals\todo{R: ``and $\udf$ denotes unknown'' entfernt, sollte an dieser Stelle klar sein}. An interpretation maps
+literals. An interpretation maps
 an integrity constraint to $\top$ iff $\Body$ is either mapped to
@@
 the least fixed point of $\Svlp$ always exists. However,
-if these programs do not contain cycles, then there always exists a supported model\todo{R: Begriff ``supported model'' unklar}, which 
-corresponds to the least fixed point of~$\Svlp$~\cite{dietz:hoelldobler:pereira:2017,dietz:hoelldobler:philipp:2017}.
-Let us denote the least fixed point of $\Svlp$ as $\CalM_\CalP$:
-As the fixed point of $\Svlp$ is unique and only computes the supported minimal model of $\WComp\CalP$, we define
+for programs without cycles, there always exists a unique supported model,
+which corresponds to the least fixed point of~$\Svlp$~\cite{dietz:hoelldobler:pereira:2017,dietz:hoelldobler:philipp:2017}. This model is \textit{supported} in the sense that 
+it will always be computed by $\Svlp$, independent with which interpretation, we start to iterate $\Svlp$ with. Let us denote this model as $\CalM_\CalP$:
+% As the fixed point of $\Svlp$ is unique and only computes the supported minimal model of $\WComp\CalP$, 
+% we define 
 $\CalP \ModelsWCS F$ iff $\CalM_\CalP(F) = \top$.
@@
 First, we introduce the basic principles that are assumed for all premises when applicable.
-Thus, we assume them to be assumed\todo{R: assumed $\to$ applied?} by all (human) reasoners.
 In Section~\ref{sect:advanced}, we present the advanced principles, which will 
@@
 Similarly, we assume that humans also understand this statement as a conditional sentence (\conditionals).
-However, it seems that humans understand them defeasibly, \todo{R: i.e.?} they assume a \textit{license for the inference}~\cite{stenning:vanlambalgen:2008},
-i.e.\ they might allow exceptions to the conditional (\licenses). For (\MA yz) this can be 
+However, it seems that humans understand them defeasibly, i.e., they assume a \textit{license for the inference}~\cite{stenning:vanlambalgen:2008} and might allow exceptions to the conditional (\licenses).
+For (\MA yz) this can be
 explicitly expressed by 
@@
 evidence that some of the participants might directly or indirectly 
-apply the contrapositive (\contraposition) within a reasoning task (cf.~\cite{rips:1994,mentallogic:1994})\todo{R: Die Referenzen gibt es bei mir nicht}.
+apply the contrapositive (\contraposition) within a reasoning task (cf.~\cite{rips:1994,mentallogic:1994}).
 Consider again $\CalP_{Ayz}$, the logic program representing (\MA yz)
@@
 For \MI yz (some) the conversion is explicitly specified, motivated by the (\converse) principle.
-By requiring for \MI yz and \MO yz the existence of an object $o_1$ with the first property $y$,
-for which the second property $z$ has to be either either false or unknown,
-\todo[inline]{R: Das ist doch nicht richtig, oder? Bei \MI yz muss $z(o_2)$ entweder false oder unknown sein, wie beschrieben, aber bei \MO yz muss es entweder true oder unknown sein.}
+By requiring for \MI yz (\MO yz, resp) the existence of an object with the first property $y$,
+for which the second property $z$ has to be either either false or unknown (true or unknown, resp.),
 \MA yz and \MI yz, (\ME yz and \MO yz, resp.), exclude each other.
@@
 Given this sobering outcome with respect to the non-aggregated evaluation, 
-we investigated whether patterns could be found on the level of the
-cognitive principles: \basic, \contraposition, \abduction\ and \negFailure.
-\todo[inline]{R: Ich finde, hier wird nicht ganz klar, was das ``level of cognitive principles'' ist}
+we ignored the clusters, and investigated whether parallels between the applied cognitive principles could be observed: \basic, \contraposition, \abduction\ and \negFailure.
 Table~\ref{tab:cogprinc} provides an overview of the predicted conclusions according to these principles.\footnote{Note that any advanced principle also includes the logic program clauses specified for \textsf{basic}.}
@@
 (2) the call of a deliberative component to search for counterexamples to conclusions.
-They\todo{R: statt ``they'' \cite{khemlani:2016}} specified four parameters for the construction of a mental model and evaluated their
+They specified four parameters for the construction of a mental model and evaluated their
 predictions based on 2 datasets, week 1 and week 2, for 20 participants.
 The parameters for the construction of the mental models are (i) the 
-size of a mental model, (ii) the probability of drawing\todo{R: was heißt draw hier?} the complete set, 
+size of a mental model, (ii) the probability of drawing a conclusions only from the most common mental models, or from the complete set,
 (iii) the probability to search for counter examples,
 and (iv) the probability to weaken the conclusion.
-The clustering with mReasoner was based on partitioning around\todo{R: around \emph{the} medoids?} medoids clustering algorithm to determine the optimal number of clusters.
+The clustering with mReasoner was based on a particular clustering algorithm to determine the optimal number of clusters for their dataset.
 The K-means clustering algorithm was used to divide participants into clusters and then

### 2019-11-11 paper/paper-v3.tex
Noch kleinere Korrekturen/Anmerkungen

@@
 Based on these moods and figures, 64 different pairs of syllogistic premises can be constructed.
-A conclusion then consists of an assertion about two properties over terms where each of these properties occurs in exactly one of the premises, and there is a third property connecting both premises which does not occur in the conclusion
+A conclusion then consists of an assertion about two properties over terms where each of these properties occurs in exactly one of the premises, and there is a third property connecting both premises which does not occur in the conclusion.
 \begin{table}[t]
@@
 
-Clustering by Principles under WCS first presented in~\cite{declare:2017},
-assumes that reasoning clusters can be expressed through cognitive principles and heuristic strategies.
+Clustering by principles under WCS was first presented in~\cite{declare:2017};
+the authors conjectured that reasoning clusters can be expressed through cognitive principles and heuristic strategies.
 Cognitive principles are assumptions (not necessarily valid under CL) made by humans while reasoning,
@@
 always has a least model under  {\L}ukasiewicz logic,
-which can be computed by the least fixed point of~$\Svlp$\cite{hk:2009a}. The correspondence to other well-known logic programming approaches, in particular 
+which can be computed by the least fixed point of~$\Svlp$~\cite{hk:2009a}. The correspondence to other well-known logic programming approaches, in particular 
 to the well-founded semantics has been shown in~\cite{dietz:hoelldobler:wernhard:2014}.
@@
 which corresponds to the least fixed point of~$\Svlp$~\cite{dietz:hoelldobler:pereira:2017,dietz:hoelldobler:philipp:2017}. This model is \textit{supported} in the sense that 
-it will always be computed by $\Svlp$, independent with which interpretation, we start to iterate $\Svlp$ with. Let us denote this model as $\CalM_\CalP$:
+it will always be computed by $\Svlp$, independent from the interpretation we start to iterate $\Svlp$ with. Let us denote this model as $\CalM_\CalP$:
 % As the fixed point of $\Svlp$ is unique and only computes the supported minimal model of $\WComp\CalP$, 
@@
 and which is semantically equivalent to $\exists X(z(X) \wedge y(X))$ (\converse).
-As~\cite{khemlani:2012}
-shows, humans seem to
+As shown by~\cite{khemlani:2012}, humans seem to
 distinguish between \textit{some y are z} and \textit{all y are z} (\unknownGen).
@@
 For \MI yz (some) the conversion is explicitly specified, motivated by the (\converse) principle.
-By requiring for \MI yz (\MO yz, resp) the existence of an object with the first property $y$,
+By requiring for \MI yz (\MO yz, resp.) the existence of an object with the first property $y$,
 for which the second property $z$ has to be either either false or unknown (true or unknown, resp.),
@@
 
-\cite{khemlani:2016} proposed an approach to model the individual reasoning patterns of the participants, and 
+\cite{khemlani:2016}\todo{R: cite als Konstituent} proposed an approach to model the individual reasoning patterns of the participants, and 
 showed that different participants seem to derive different conclusions.
@@
 (2) the call of a deliberative component to search for counterexamples to conclusions.
-They specified four parameters for the construction of a mental model and evaluated their
+They\todo{R: oder doch ``Khemlani''? ;)} specified four parameters for the construction of a mental model and evaluated their
 predictions based on 2 datasets, week 1 and week 2, for 20 participants.
 The parameters for the construction of the mental models are (i) the 
-size of a mental model, (ii) the probability of drawing a conclusions only from the most common mental models, or from the complete set,
+size of a mental model, (ii) the probability of drawing conclusions only from the most common mental models, or from the complete set,
 (iii) the probability to search for counter examples,
@@
 the mReaoner to simulated the clusters by choosing appropriate parameter settings.
-\cite{khemlani:2016} identified three different reasoning clusters: The intuitive cluster, the intermediate cluster and the deliberative cluster.
+\cite{khemlani:2016}\todo{R: cite als Konstituent} identified three different reasoning clusters: The intuitive cluster, the intermediate cluster and the deliberative cluster.
 The approach is novel in the sense that until now no other cognitive theory tried to rigorously account for individual differences in human syllogistic reasoning.
@@
 We have investigated the cognitive adequacy of clustering by principles under WCS with respect to the individual answer patterns reported by~\cite{khemlani:2016}.
-\cite{khemlani:2016} is the only approach we are aware of that has evaluated predictions on human syllogistic reasoning
+\cite{khemlani:2016}\todo{R: cite als Konstituent} is the only approach we are aware of that has evaluated predictions on human syllogistic reasoning
 with respect to the non-aggregated data.

### 2019-11-12 paper/paper-v3.tex
added authors names for citations, where necessary

@@
 Conceptual cognitive adequacy describes to what extent a system corresponds to human conceptual knowledge, whereas inferential cognitive adequacy indicates whether the reasoning process of a system is structured similarly to the way humans reason~\cite{knauff:rauh:renz:1997}.
-In order to determine the required factors that are necessary for the evaluation of an adequate cognitive theory, we take~\cite{ragni:stolzenburg:2015} as a starting point:
+In order to determine the required factors that are necessary for the evaluation of an adequate cognitive theory, we take~the suggestion by Ragni and Stolzburg~\cite{ragni:stolzenburg:2015} as a starting point:
 We aim for a theory in which (i) all processes are comprehensible, (ii) the suggested underlying represent general cognitive processes, and (iii) the predictions of our theory are evaluated with respect to a wide range of human reasoning tasks.
@@
 % A major drawback of 
-The results of the meta-study on syllogistic reasoning in~\cite{khemlani:2012} show that the answers given by humans do not only systematically differ from the valid conclusions under CL, but from any other proposed cognitive theory (cf.~\cite{BJ89,johnsonlaird:1983,Rips1994,Polk1995,chater:oaksford:1999}).
+The results of the meta-study on syllogistic reasoning by Khemlani and Johnson-Laird~\cite{khemlani:2012} show that the answers given by humans do not only systematically differ from the valid conclusions under CL, but from any other proposed cognitive theory (cf.~\cite{BJ89,johnsonlaird:1983,Rips1994,Polk1995,chater:oaksford:1999}).
 % Possibly one of the reasons could be that none of the suggested cognitive theories so far has accounted for differences among individuals, i.e.\ they only considered the data summed up with respect to the individual answer possibilities.
@@
 two premises, which conclusion on the relation between \emph{artists} and \emph{chemists} necessarily follows?
-According to the authors in~\cite{khemlani:2012}, the majority of participants in experimental studies
+According to Khemlani and Johnson-Laird~\cite{khemlani:2012}, the majority of participants in experimental studies
 concluded \emph{No chemists are artists} (29\;\%) and \emph{No valid conclusion} (27\;\%).
@@
 Here, we focus on a qualitative assessment of the clustering approach proposed in~\cite{declare:2017},
-in the sense that we investigate the individual participants' conclusions with respect to the non-aggregated dataset provided by the authors in~\cite{khemlani:2016}.
+in the sense that we investigate the individual participants' conclusions with respect to the non-aggregated dataset provided by Johnson-Laird and Khemlani~\cite{khemlani:2016}.
 In Section~\ref{sect:wcsclustering} and~\ref{sect:representing}, we introduce the underlying computational logic approach and the formal representation of cognitive principles within this approach.
@@
 conclusions are not derived by reasoning but instead based on structural patterns or {biases}.
-This approach is an extension of the computational logic approach for human syllogistic reasoning presented in~\cite{cogsci:2017}: Three reasoning clusters and two heuristic strategies were specified, and achieved a match of 92\;\%, which was the highest result compared to the presented approaches in~\cite{khemlani:2012}.
+This approach is an extension of the computational logic approach for human syllogistic reasoning presented in~\cite{cogsci:2017}: Three reasoning clusters and two heuristic strategies were specified, and achieved a match of 92\;\%, which was the highest result compared to the presented approaches by Khemlani an Johnson-Laird~\cite{khemlani:2012}.
 % The system developed in~\cite{declare:2017}, was extended and participated at the Syllogism Challenge 2017.\footnote{\url{https://www.cc.uni-freiburg.de/modelingchallenge/challenge-2017}} 
@@
 
-Consider the following semantic operator~\cite{stenning:vanlambalgen:2008}: Let $I = \langle I^\top,I^\bot
+Consider the following semantic operator introduced by Stenning and van Lambalgen~\cite{stenning:vanlambalgen:2008}: Let $I = \langle I^\top,I^\bot
 \rangle$ be an interpretation. $\Phi_\CalP (I) =
@@
 Similarly, we assume that humans also understand this statement as a conditional sentence (\conditionals).
-However, it seems that humans understand them defeasibly, i.e., they assume a \textit{license for the inference}~\cite{stenning:vanlambalgen:2008} and might allow exceptions to the conditional (\licenses).
+However, it seems that humans understand them defeasibly.
+According to Stenning and van Lambalgen~\cite{stenning:vanlambalgen:2008}, they assume a \textit{license for the inference} and might allow exceptions to the conditional (\licenses).
 For (\MA yz) this can be
@@
 
-Furthermore, humans have a pragmatic understanding of quantifiers in the language~\cite{grice1975}.
+Furthermore, according to Grice~\cite{grice1975}, humans have a pragmatic understanding of quantifiers in the language.
 For instance, in natural language, we normally do not quantify over things that do not exist.
@@
 and which is semantically equivalent to $\exists X(z(X) \wedge y(X))$ (\converse).
-As shown by~\cite{khemlani:2012}, humans seem to
+As shown by Khemlani and Johnson-Laird~\cite{khemlani:2012}, humans seem to
 distinguish between \textit{some y are z} and \textit{all y are z} (\unknownGen).
@@
     (3) \basic, \converse\ for \MI\ and \ME, and \contraposition\ for~\MA.
-This approach achieved a match of 92\;\% with the aggregated data reported in~\cite{khemlani:2012}, and was the highest result compared to the presented approaches in~\cite{khemlani:2012}.
+This approach achieved a match of 92\;\% with the aggregated data reported in~\cite{khemlani:2012}, and was the highest result compared to the presented approaches.
 
@@
 We are interested whether these clusters also correspond to the reasoning patterns observed among individuals. 
-For this purpose, we investigated the non-aggregated dataset provided by~\cite{khemlani:2016} consisting of the responses of 20 participants for two weeks,
+For this purpose, we investigated the non-aggregated dataset provided by Johnson-Laird and Khemlani~\cite{khemlani:2016} consisting of the responses of 20 participants for two weeks,
 i.e.\ the participants carried out the task twice, with a break of one week. 
@@
 
-\cite{khemlani:2016}\todo{R: cite als Konstituent} proposed an approach to model the individual reasoning patterns of the participants, and 
+Johnson-Laird and Khemlani~\cite{khemlani:2016} proposed an approach to model the individual reasoning patterns of the participants, and 
 showed that different participants seem to derive different conclusions.
-Their approach is built on top of the mReasoner~\cite{khemlani:2013}, an extension of the mental model theory~\cite{johnsonlaird:1983}.
+Their approach is built on top of the mReasoner~\cite{khemlani:2013}, an extension of the mental model theory, first introduced by Johnson-Laird~\cite{johnsonlaird:1983}.
 Reasoning depends on the construction and manipulation of mental models, based on 3 fundamental principles: 
@@
 (2) the call of a deliberative component to search for counterexamples to conclusions.
-They\todo{R: oder doch ``Khemlani''? ;)} specified four parameters for the construction of a mental model and evaluated their
+Johnson-Laird and Khemlani specified four parameters for the construction of a mental model and evaluated their
 predictions based on 2 datasets, week 1 and week 2, for 20 participants.
@@
 the mReaoner to simulated the clusters by choosing appropriate parameter settings.
-\cite{khemlani:2016}\todo{R: cite als Konstituent} identified three different reasoning clusters: The intuitive cluster, the intermediate cluster and the deliberative cluster.
+Johnson-Laird and Khemlani~\cite{khemlani:2016} identified three different reasoning clusters: The intuitive cluster, the intermediate cluster and the deliberative cluster.
 The approach is novel in the sense that until now no other cognitive theory tried to rigorously account for individual differences in human syllogistic reasoning.
@@
 
-The non-aggregated data provided by~\cite{khemlani:2016} included a dataset of 20 participants,
-for which the task was carried out twice, with a break of one week. On average, the basic principles explained the results about 10\;\% better than CL.
+The non-aggregated data included a dataset of 20 participants,
+for which the task was carried out twice, with a break of one week~\cite{khemlani:2016}. On average, the basic principles explained the results about 10\;\% better than CL.
  Generally, the participants who were strong in CL/ contraposition, drew fewer conclusions according to abduction/ generalization, and vice versa.
@@
 
-Similar to~\cite{khemlani:2016}, we identified three different clusters: 
+We identified three different clusters: 
 The last two columns show the classification of the participants as deliberative, intuitive or intermediate. 
-The participants in the deliberative cluster here corresponds exactly to~\cite{khemlani:2016}.
-Our classification only differs to~\cite{khemlani:2016} for the intermediate and intuitive cluster with respect to three participants (1, 11 and 19).
-A possible explanation for this difference, could be that~\cite{khemlani:2016}
+% The participants in the deliberative cluster here corresponds exactly to~\cite{khemlani:2016}. 
+Our classification only differs wrt Johnson-Laird and Khemlani's 
+classification~\cite{khemlani:2016} for the intermediate and intuitive cluster with respect to three participants (1, 11 and 19).
+A possible explanation for this difference, could be that they
 did not take into account the participants' changes between week 1 and week 2.
@@
 
-We have investigated the cognitive adequacy of clustering by principles under WCS with respect to the individual answer patterns reported by~\cite{khemlani:2016}.
-\cite{khemlani:2016}\todo{R: cite als Konstituent} is the only approach we are aware of that has evaluated predictions on human syllogistic reasoning
+We have investigated the cognitive adequacy of clustering by principles under WCS with respect to the individual answer patterns reported by Johnson-Laird and Khemlanid~\cite{khemlani:2016}.
+Their approach is the only one we are aware of that has evaluated predictions on human syllogistic reasoning
 with respect to the non-aggregated data.
@@
 For the future, we need to investigate whether the sequential application of the cognitive principles 
-can give us more insights on the participants' reasoning pattern.
-In particular, our hypothesis is that these principles might correspond to some of the reasoning steps within the insight model~\cite{johnsonlaird:1970} for the selection task~\cite{wason:68}.
-Similarly, a meta-analysis of the individual response patterns for the selection task has been suggested in~\cite{ragni:2017}.
+can give us more insights on the participants' reasoning pattern. In particular, our hypothesis is that these principles might correspond to some of the reasoning steps within the insight model
+by Johnson-Laird and Wason~\cite{johnsonlaird:1970} for the selection task~\cite{wason:68}.
+Similarly, a meta-analysis of the individual response patterns for the selection task has been suggested by Ragni, Kola and Johnson-Laird~\cite{ragni:2017}.
 This in turn might give us a new starting point to investigate whether the cognitive principles  specified here can characterize participants' conclusions in other reasoning episodes.

### 2019-11-12 paper/paper-v3.tex
klammern dort wo nicht noetig weggelassen

@@
 According to Stenning and van Lambalgen~\cite{stenning:vanlambalgen:2008}, they assume a \textit{license for the inference} and might allow exceptions to the conditional (\licenses).
-For (\MA yz) this can be
+For \MA yz this can be
 explicitly expressed by 
@@
 % require existential import for a conditional to be true.
-Together with the introduced cognitive principles in Section~\ref{sect:condli}, the logic program representing (\MA yz), $\CalP_{Ayz}$, is as follows:
+Together with the introduced cognitive principles in Section~\ref{sect:condli}, the logic program representing \MA yz, $\CalP_{Ayz}$, is as follows:
 \begin{align} %{rll@{\hspace{\hfill}}r}
@@
 \end{align}
-Note that $z'(X)$ appears within the $\cxt$ operator in the body of the second clause, where $z'$ represents the negation of $z$:\footnote{$z'(X)$ will be introduced by the (\footnotesize{\textsf{transformation}}) principle in Section~\ref{sect:dneg}.}
+Note that $z'(X)$ appears within the $\cxt$ operator in the body of the second clause, where $z'$ represents the negation of $z$:\footnote{$z'(X)$ will be introduced by the \footnotesize{\textsf{transformation}} principle in Section~\ref{sect:dneg}.}
 Nothing is abnormal, unless the negation of $z$ is true.
@@
 Consider the statement \textit{No y is z.} ({\ME yz}):
-Similarly to the previous case, (\licenses) and (\import) apply here
+Similarly to the previous case, \licenses and \import apply here
 as well.
@@
 Hence, we block them with the help of abnormalities (\dnegation).
-$\CalP_{Eyz}$, the logic program representing (\ME yz) is as follows:
+$\CalP_{Eyz}$, the logic program representing \ME yz is as follows:
 % \footnote{For simplicity, we omitted to list the clauses that are introduced by the (\converse) principle in $\CalP_{Eyz}$: The same clauses just switching $y$ and $z$ are added and thus also the fact $z(o_2) \leftarrow \top$ for another object $o_2$.
@@
 $y$ and which does not belong to~$z$.
-$\CalP_{Iyz}$, the logic program representing (\MI yz), is as follows:\footnote{For simplicity, here we omit the clauses introduced by ({\footnotesize\textsf{converse}}).
+$\CalP_{Iyz}$, the logic program representing \MI yz, is as follows:\footnote{For simplicity, here we omit the clauses introduced by ({\footnotesize\textsf{converse}}).
 They can be found in~\cite{declare:2017}.}
@@
 Consider the statement \textit{Some y are not z} ({\MO yz}), which in CL is represented as $\exists X (y(X) \wedge \neg z(X)$.
-As in the previous cases, (\conditionals), (\licenses) and (\unknownGen) are assumed.
-As the conclusion is again a negation of a claim, (\transformation) and
-(\dnegation) apply as well. $\CalP_{Oyz}$, the logic program representing (\MO yz) is as follows:
+As in the previous cases, \conditionals, \licenses and \unknownGen are assumed.
+As the conclusion is again a negation of a claim, \transformation and
+\dnegation apply as well. $\CalP_{Oyz}$, the logic program representing \MO yz is as follows:
 \begin{align}
@@
 apply the contrapositive (\contraposition) within a reasoning task (cf.~\cite{rips:1994,mentallogic:1994}).
-Consider again $\CalP_{Ayz}$, the logic program representing (\MA yz)
+Consider again $\CalP_{Ayz}$, the logic program representing \MA yz
 from Section~\ref{sect:basic}. The clauses that represent the contraposition principles are as follows:
@@
 \end{align}
-The application of (\contraposition) for the \ME\ mood,
+The application of \contraposition for the \ME\ mood,
 \textit{No y is z}, yields \textit{No z is y}, and consists of the following clauses:
@@
 (\MI or \MO), then the only object about which an inference can be made is the one
-resulting from (\import) with respect to the premise it introduced (i.e.\ only inferences about the properties 
+resulting from \import with respect to the premise it introduced (i.e.\ only inferences about the properties 
 of the object about $a$ and $b$ or
@@
 Consider again $\CalP_{I yz}$ of Section~\ref{sect:basic}, where $\Ab_{yz}$ is only false for 
-$o_1$. In case (\negFailure) applies, then $\Ab_{yz}(o_1) \leftarrow \bot$
+$o_1$. In case \negFailure applies, then $\Ab_{yz}(o_1) \leftarrow \bot$
 in $\CalP_{I yz}$ is replaced by the following two clauses:
@@
 \end{align}
-Similar to the case of $\CalP_{Ayz}$ for the representation of the basic principles of (\MA yz),
+Similar to the case of $\CalP_{Ayz}$ for the representation of the basic principles of \MA yz,
 the $\cxt$ operator in the body of the first clause is false for all $o$
 for which $z'(o)$ is unknown. This clause transforms the representation for the existential quantification over \MI yz in $\CalP_{Iyz}$
-into a universal one. In order to maintain (\unknownGen), the second clause is added:
+into a universal one. In order to maintain \unknownGen, the second clause is added:
 $\Ab_{yz}(o_2) \leftarrow \udf$ guarantees that $\Ab_{yz}(o_2)$ stays unknown, even though $\cxt(z'(o_2))$ might be false.
@@
 
-When participants are faced with an \NVC\ conclusion (\textit{No valid conclusion}), they might not want to accept this conclusion and proceed to check whether there exists unknown information that might be relevant (\abduction).
+When participants are faced with \NVC\ (\textit{No valid conclusion}), they might not want to accept this conclusion and proceed to check whether there exists unknown information that might be relevant.
 This information may be explanations about
 facts introduced either from an existential import or from unknown generalization. 
-We use the first as source for observations, as they are used directly to infer new information. 
-Consider the following pair of syllogistic premises: 
+We use the first as source for observations, as they are used directly to infer new information  (\abduction). Consider the following pair of syllogistic premises: 
 \begin{align}
@@
 \end{description}
-For \MI yz (some) the conversion is explicitly specified, motivated by the (\converse) principle.
+For \MI yz (some) the conversion is explicitly specified, motivated by the \converse principle.
 By requiring for \MI yz (\MO yz, resp.) the existence of an object with the first property $y$,

### 2019-11-12 paper/paper-v3.tex
Fix typos

@@
 Conceptual cognitive adequacy describes to what extent a system corresponds to human conceptual knowledge, whereas inferential cognitive adequacy indicates whether the reasoning process of a system is structured similarly to the way humans reason~\cite{knauff:rauh:renz:1997}.
-In order to determine the required factors that are necessary for the evaluation of an adequate cognitive theory, we take~the suggestion by Ragni and Stolzburg~\cite{ragni:stolzenburg:2015} as a starting point:
+In order to determine the required factors that are necessary for the evaluation of an adequate cognitive theory, we take~the suggestion by Ragni and Stolzenburg~\cite{ragni:stolzenburg:2015} as a starting point:
 We aim for a theory in which (i) all processes are comprehensible, (ii) the suggested underlying represent general cognitive processes, and (iii) the predictions of our theory are evaluated with respect to a wide range of human reasoning tasks.
@@
 Consider the statement \textit{No y is z.} ({\ME yz}):
-Similarly to the previous case, \licenses and \import apply here
+Similarly to the previous case, \licenses\ and \import\ apply here
 as well.
@@
 Consider the statement \textit{Some y are not z} ({\MO yz}), which in CL is represented as $\exists X (y(X) \wedge \neg z(X)$.
-As in the previous cases, \conditionals, \licenses and \unknownGen are assumed.
-As the conclusion is again a negation of a claim, \transformation and
-\dnegation apply as well. $\CalP_{Oyz}$, the logic program representing \MO yz is as follows:
+As in the previous cases, \conditionals, \licenses, and \unknownGen are assumed.
+As the conclusion is again a negation of a claim, \transformation\ and
+\dnegation\ apply as well. $\CalP_{Oyz}$, the logic program representing \MO yz is as follows:
 \begin{align}
@@
 \end{align}
-The application of \contraposition for the \ME\ mood,
+The application of \contraposition\ for the \ME\ mood,
 \textit{No y is z}, yields \textit{No z is y}, and consists of the following clauses:
@@
 (\MI or \MO), then the only object about which an inference can be made is the one
-resulting from \import with respect to the premise it introduced (i.e.\ only inferences about the properties 
+resulting from \import\ with respect to the premise it introduced (i.e.\ only inferences about the properties 
 of the object about $a$ and $b$ or
@@
 Consider again $\CalP_{I yz}$ of Section~\ref{sect:basic}, where $\Ab_{yz}$ is only false for 
-$o_1$. In case \negFailure applies, then $\Ab_{yz}(o_1) \leftarrow \bot$
+$o_1$. In case \negFailure\ applies, then $\Ab_{yz}(o_1) \leftarrow \bot$
 in $\CalP_{I yz}$ is replaced by the following two clauses:
@@
 \end{description}
-For \MI yz (some) the conversion is explicitly specified, motivated by the \converse principle.
+For \MI yz (some) the conversion is explicitly specified, motivated by the \converse\ principle.
 By requiring for \MI yz (\MO yz, resp.) the existence of an object with the first property $y$,
@@
 % The participants in the deliberative cluster here corresponds exactly to~\cite{khemlani:2016}. 
-Our classification only differs wrt Johnson-Laird and Khemlani's 
+Our classification only differs wrt.\ Johnson-Laird and Khemlani's 
 classification~\cite{khemlani:2016} for the intermediate and intuitive cluster with respect to three participants (1, 11 and 19).
-A possible explanation for this difference, could be that they
+A possible explanation for this difference could be that they
 did not take into account the participants' changes between week 1 and week 2.
@@
 
-We have investigated the cognitive adequacy of clustering by principles under WCS with respect to the individual answer patterns reported by Johnson-Laird and Khemlanid~\cite{khemlani:2016}.
+We have investigated the cognitive adequacy of clustering by principles under WCS with respect to the individual answer patterns reported by Johnson-Laird and Khemlani~\cite{khemlani:2016}.
 Their approach is the only one we are aware of that has evaluated predictions on human syllogistic reasoning

### 2019-11-14 paper/paper-v3.tex
added Luka nec operator ref

@@
 Table~\ref{tab:3vldluka}.
+\footnote{Thanks to the
+ anonymous reviewer for the observation, that the $\cxt$ operator seems to be a special case of {\L}ukasiwicz's necessity operator~\cite[p. 20]{malinowski:1993}, restricted to literals.
+ }
+
 $I(F) = \top$ denotes that a formula $F$ is mapped to true under $I$.
@@
 
-Another interesting aspect is to investigate in how far the semantics of the $\cxt$ operator within the extended three-valued {\L}ukasiewicz logic in Section~\ref{sub:3valuedL} 
- corresponds to the {\L}ukasiewicz extensional necessity operator~\cite{}\todo{R: Referenz fehlt}.\footnote{Thanks to the
- anonymous reviewer for this observation.}
-
 % episodes.

### 2019-11-18 paper/paper-v3.tex
made the (hopfully very final changes

@@
 In order to determine the required factors that are necessary for the evaluation of an adequate cognitive theory, we take~the suggestion by Ragni and Stolzenburg~\cite{ragni:stolzenburg:2015} as a starting point:
-We aim for a theory in which (i) all processes are comprehensible, (ii) the suggested underlying represent general cognitive processes, and (iii) the predictions of our theory are evaluated with respect to a wide range of human reasoning tasks.
+We aim for a theory in which (i) all processes are comprehensible, (ii) the underlying assumptions represent general cognitive processes, and (iii) the predictions of our theory are evaluated with respect to a wide range of human reasoning tasks.
 
@@
 % A major drawback of 
-The results of the meta-study on syllogistic reasoning by Khemlani and Johnson-Laird~\cite{khemlani:2012} show that the answers given by humans do not only systematically differ from the valid conclusions under CL, but from any other proposed cognitive theory (cf.~\cite{BJ89,johnsonlaird:1983,Rips1994,Polk1995,chater:oaksford:1999}).
-% Possibly one of the reasons could be that none of the suggested cognitive theories so far has accounted for differences among individuals, i.e.\ they only considered the data summed up with respect to the individual answer possibilities.
+The results of the meta-study on syllogistic reasoning by Khemlani and Johnson-Laird~\cite{khemlani:2012} show that the responses given by humans do not only systematically differ from the valid conclusions under CL, but from any other proposed cognitive theory (cf.~\cite{BJ89,johnsonlaird:1983,Rips1994,Polk1995,chater:oaksford:1999}).
+% Possibly one of the reasons could be that none of the suggested cognitive theories so far has accounted for differences among individuals, i.e.\ they only considered the data summed up with respect to the individual response possibilities.
 % As recent investigations on psychological experiments show (cf.~\cite{khemlani:2016, ragni:2017}), it seems that different humans represent and reason about conditional information differently.
@@
 premises and one conclusion,
-each of which is a statement about properties over terms which can have one of the classical quantifiess in Table~\ref{table:moods} (called \emph{moods}).
+each of which is a statement about properties over terms which can have one of the classical quantifiers in Table~\ref{table:moods} (called \emph{moods}).
 The pair of syllogistic premises can have four different orders as shown in Table~\ref{t:2} (called \emph{figures}).
@@
 \begin{table}[t]
-    \begin{minipage}{0.6\textwidth}
+    \begin{minipage}{0.65\textwidth}
         \centering
 	\[	
-\begin{array}{@{\hspace{0.0cm}}l@{\hspace{0.1cm}}G@{\hspace{0.1cm}}l@{\hspace{0.3cm}}c}\toprule
+\begin{array}{@{\hspace{0.0cm}}l@{\hspace{0.4cm}}G@{\hspace{0.1cm}}l@{\hspace{0.3cm}}c}\toprule
 	\mbox{Mood} & \mbox{natural language} & \mbox{First-order Logic} &
@@
 	\]
-	\caption{The four moods and their
+	\caption{The classical moods and their
 formalization\label{table:moods}.}
 	\end{minipage}
-    \begin{minipage}{0.45\textwidth}
+    \begin{minipage}{0.34\textwidth}
         \centering
@@
 	\end{center}
-	\caption{The four figures.\label{t:2}}
+	\caption{The classical figures.\label{t:2}}
 	\end{minipage}
@@
 The valid conclusion under CL, \emph{Some artists are not chemists},
-was concluded by only 16\,\%, which is exactly the threshold that the answer had not been chosen randomly~\cite{khemlani:2012}.
+was concluded by only 16\,\%, which is exactly the threshold that the response had not been chosen randomly~\cite{khemlani:2012}.
 % \footnote{
@@
 % One reason is because often only {aggregated data} of the experiments are accessible.
-% i.e.\ the data summed up with respect to the individual answer possibilities.
+% i.e.\ the data summed up with respect to the individual response possibilities.
 % The experimental set-up was made such that participants would only choose one possible conclusion, which implies that 
-% these answers excluded each other. In particular, it can be assumed that participants were at least minimally consistent in the sense that
+% these responses excluded each other. In particular, it can be assumed that participants were at least minimally consistent in the sense that
 % i.e.\ 
 % The majority's conclusions exclude each other, as the test was made such that
-Usually in these experiments, participants are supposed to give only one answer, thus all three (significant) answers
+Usually in these experiments, participants are supposed to give only one response, thus all three responses above
 were given by different participants. 
-% The ones who answered \emph{No valid conclusion}
-% were not the same ones who answered \emph{No chemists are artists} or \emph{Some artists are not chemists}, and vice versa.\footnote{}
-
-A major drawback of the suggested cognitive theories so far is that only the participants'
+% The ones who responseed \emph{No valid conclusion}
+% were not the same ones who responseed \emph{No chemists are artists} or \emph{Some artists are not chemists}, and vice versa.\footnote{}
+
+A major drawback of most cognitive theories is that only the participants'
 aggregated data was considered: These theories did not account for differences
-among individuals, i.e.\ they only considered the data summed up with respect to the individual answer possibilities.
+among individuals, i.e.\ they only considered the data summed up with respect to the individual response possibilities.
 In contrast, non-aggregated data records the individual participants' response patterns, i.e.\ the complete set of responses that are given by each participant.
@@
 This implies that \textit{the} human reasoner does not exist, instead there 
-seem to exist various (human) reasoning clusters. The identification of these differences and the specification of these clusters is central 
+seem to exist various groups of reasoners, i.e.\ reasoning clusters. The identification of these differences and the specification of these clusters is central 
 for the development of an adequate cognitive theory.
@@
 % Interestingly, when comparing the changes of results of each individual participant
-% from the first to the second week, it seems that their answers in the second week fits better the classical logic derivations. 
+% from the first to the second week, it seems that their responses in the second week fits better the classical logic derivations. 
 % We argue that these reasoners can be characterized by a set of cognitive principles, motivated from studies made in Cognitive Science.
@@
 % Interestingly, when comparing the changes of results of each individual participant
-% from the first to the second week, it seems that their answers in the second week fits better the classical logic derivations. 
+% from the first to the second week, it seems that their responses in the second week fits better the classical logic derivations. 
 % We formulate the hypothesis that the individual clusters do not represent different human reasoners, but different stages of reasoning. Depending on the persistence and the
@@
 
-Here, we focus on a qualitative assessment of the clustering approach proposed in~\cite{declare:2017},
-in the sense that we investigate the individual participants' conclusions with respect to the non-aggregated dataset provided by Johnson-Laird and Khemlani~\cite{khemlani:2016}.
-In Section~\ref{sect:wcsclustering} and~\ref{sect:representing}, we introduce the underlying computational logic approach and the formal representation of cognitive principles within this approach.
+Here, we provide a qualitative assessment of the \textit{clustering by principles} approach proposed in~\cite{declare:2017},
+that performed well on aggregated data, and investigate whether these clusters are also adequate with respect to the non-aggregated dataset given by Johnson-Laird and Khemlani~\cite{khemlani:2016}.
+In Section~\ref{sect:wcsclustering} and~\ref{sect:representing}, we introduce the underlying computational logic approach, the cognitive principles and their formalization.
 Section~\ref{sect:aggregateddata} illustrates an example and briefly discusses the performance of the approach presented in~\cite{declare:2017} on non-aggregated data.
@@
 
-\section{Clustering by Principles under WCS} \label{sect:wcsclustering}
-
-Clustering by principles under WCS was first presented in~\cite{declare:2017};
-the authors conjectured that reasoning clusters can be expressed through cognitive principles and heuristic strategies.
+\section{Clustering by Principles} \label{sect:wcsclustering}
+
+Clustering by principles was first presented in~\cite{declare:2017}, which is an extension of
+the computational logic approach for human syllogistic reasoning presented in~\cite{cogsci:2017}.
+The authors in~\cite{declare:2017} conjectured that reasoning clusters can be expressed through cognitive principles and heuristic strategies.
 Cognitive principles are assumptions (not necessarily valid under CL) made by humans while reasoning,
-whereas heuristic strategies describe the methods that are applied when
-conclusions are not derived by reasoning but instead based on structural patterns or {biases}.
-This approach is an extension of the computational logic approach for human syllogistic reasoning presented in~\cite{cogsci:2017}: Three reasoning clusters and two heuristic strategies were specified, and achieved a match of 92\;\%, which was the highest result compared to the presented approaches by Khemlani an Johnson-Laird~\cite{khemlani:2012}.
+whereas heuristic strategies describe the methods that are applied when conclusions are not derived by reasoning but instead based on structural patterns or {biases}: Three reasoning clusters and two heuristic strategies were specified, and achieved a match of 92\;\%, which was the highest result compared to the presented approaches by Khemlani an Johnson-Laird~\cite{khemlani:2012}.
 % The system developed in~\cite{declare:2017}, was extended and participated at the Syllogism Challenge 2017.\footnote{\url{https://www.cc.uni-freiburg.de/modelingchallenge/challenge-2017}} 
 Figure~\ref{overview:system:syllchal} illustrates the components of the system:
-First, groups of reasoners were specified based on a set of underlying cognitive principles and heuristic strategies.
-After that, this set was translated into a logic program (Section~\ref{sect:basic}-\ref{sect:advanced}), representing the group of reasoners.
+First, groups of reasoners were specified through sets of underlying cognitive principles and heuristic strategies.
+After that, these sets were translated into logic programs (Section~\ref{sect:basic}-\ref{sect:advanced}), representing the reasoning clusters.
 Next, the respective models were computed (Section~\ref{sect:wcs}-\ref{sect:lfp}).
-After that, the entailed relations between $a$ and $c$ were extracted (Section~\ref{sect:entail}). 
- The final answer distribution was specified according to the number of reasoning clusters 
+After that, for each cluster, the entailed relations between $a$ and $c$ were extracted (Section~\ref{sect:entail}). 
+ The final response distribution was specified according to the number of reasoning clusters 
 that entailed the conclusion.
@@
 \node[anchor=west]  at (0,9.2) {\textbf{Cognitive Principles} presuppositions that humans make while reasoning};
-\node[anchor=west]  at (0,8.7) {\textbf{Heuristic Strategies} reasoning based on structure in patterns or biases};
+\node[anchor=west]  at (0,8.7) {\textbf{Heuristic Strategies} \hspace{0.05cm} reasoning based on structure in patterns or biases};
 
@@
 \end{tikzpicture}
-\caption{The system that participated at the Syllogism Challenge 2017\label{overview:system:syllchal}.}
+\caption{The methodology of \textit{clustering by principles}\label{overview:system:syllchal}.}
 
@@
 is the Weak Completion Semantics (WCS)~\cite{hk:2009a,hk:2009b}
-which has shown to model a wide range of human reasoning tasks~\cite{btg:2015:h}. 
+which has shown to model a wide range of human reasoning tasks~\cite{btg:2015:h}
+and will be introduced in the following.
 % WCS is the non-monotonic computational logic approach to consider
@@
 % contain an unknown assumption with head $A$, and $\Def(A,\CalP) = \emptyset$.
-We omit the word \textit{contextual} when we refer to programs, if not stated otherwise.
+In the sequel, we omit the word \textit{contextual} when we refer to programs, if not stated otherwise.
 
@@
 Table~\ref{tab:3vldluka}.
-\footnote{Thanks to the
- anonymous reviewer for the observation, that the $\cxt$ operator seems to be a special case of {\L}ukasiwicz's necessity operator~\cite[p. 20]{malinowski:1993}, restricted to literals.
- }
 
@@
 \begin{array}{c|c}
- L & \cxt(L)\\
+ L & \cxt(L)\footnotemark\\
 \midrule
@@
 \end{table}
+\footnotetext{Thanks to the
+ anonymous reviewer for the observation, that the $\cxt$ operator seems to be a special case of {\L}ukasiwicz's necessity operator~\cite[p. 20]{malinowski:1993}, restricted to literals.
+ }
 
@@
 \]
-The weak completion of non-contextual programs\footnote{Non-contextual programs refer programs that do not contain rules in which $\cxt$ appears in the body of any rule, i.e.\ $p = 0$.} 
+The weak completion of non-contextual programs\footnote{Non-contextual programs are programs that do not contain rules in which $\cxt$ appears in the body of any rule, i.e.\ $p = 0$.} 
 always has a least model under  {\L}ukasiewicz logic,
-which can be computed by the least fixed point of~$\Svlp$~\cite{hk:2009a}. The correspondence to other well-known logic programming approaches, in particular 
-to the well-founded semantics has been shown in~\cite{dietz:hoelldobler:wernhard:2014}.
-
-The weak completion of contextual
-programs might have more than one minimal model, and it is not guaranteed that 
-the least fixed point of $\Svlp$ always exists. However,
-for programs without cycles, there always exists a unique supported model,
+which corresponds to the least fixed point of~$\Svlp$~\cite{hk:2009a}. 
+The interested reader is referred to~\cite{dietz:hoelldobler:wernhard:2014} for the correspondence to other 
+logic programming approaches.
+
+Different than for non-contextual programs, the weak completion of contextual
+programs might have more than one minimal model. Furthermore, if $\CalP$ is a contextual program, the least fixed point of $\Svlp$ might not even exist. However,
+for contextual programs without cycles, there always exists a unique supported model,
 which corresponds to the least fixed point of~$\Svlp$~\cite{dietz:hoelldobler:pereira:2017,dietz:hoelldobler:philipp:2017}. This model is \textit{supported} in the sense that 
-it will always be computed by $\Svlp$, independent from the interpretation we start to iterate $\Svlp$ with. Let us denote this model as $\CalM_\CalP$:
-% As the fixed point of $\Svlp$ is unique and only computes the supported minimal model of $\WComp\CalP$, 
-% we define 
+it will always be computed by $\Svlp$, independent from the interpretation we start to iterate $\Svlp$ with. Let us denote this model as $\CalM_\CalP$.
+As the fixed point of $\Svlp$ is unique and only computes the supported minimal model of $\WComp\CalP$, 
+we define 
 $\CalP \ModelsWCS F$ iff $\CalM_\CalP(F) = \top$.
@@
 In the following, the programs are
-specified using the predicates $y$ and $z$, where~$yz$ can be replaced by $ab$, $ba$, $cb$, or $bc$ and depend on the figures shown in Table~\ref{t:2}.
+specified using the predicates $y$ and $z$, where~$yz$ can be replaced by $ab$, $ba$, $cb$, or $bc$ and depend on the figures in Table~\ref{t:2}.
 First, we introduce the basic principles that are assumed for all premises when applicable.
@@
 Consider \textit{All y are z} ({\MA yz}), which in CL, is represented as $\forall X (y(X) \rightarrow z(X))$.
-Similarly, we assume that humans also understand this statement as a conditional sentence (\conditionals).
-However, it seems that humans understand them defeasibly.
+Similarly, we assume that humans also understand this statement as a conditional sentence (\conditionals),
+but also defeasible.
 According to Stenning and van Lambalgen~\cite{stenning:vanlambalgen:2008}, they assume a \textit{license for the inference} and might allow exceptions to the conditional (\licenses).
@@
 \end{align}
-Note that $z'(X)$ appears within the $\cxt$ operator in the body of the second clause, where $z'$ represents the negation of $z$:\footnote{$z'(X)$ will be introduced by the \footnotesize{\textsf{transformation}} principle in Section~\ref{sect:dneg}.}
+Note that $z'(X)$ appears within the $\cxt$ operator in the body of the second clause, where $z'$ represents the negation of $z$ (which will be introduced by the \transformation\ principle in Section~\ref{sect:dneg}).
 Nothing is abnormal, unless the negation of $z$ is true.
@@
 
-\subsubsection{Negation by Transformation (\transformation) and \\No Double Negation (\dnegation)} \label{sect:dneg}
+\subsubsection{Negation by Transformation (\transformation) and \\No Inference through Double Negation (\dnegation)} \label{sect:dneg}
 
@@
 by introducing an auxiliary formula $p'(X)$ together with the clause $p(X) \leftarrow \neg p'(X)$ and the integrity constraint $\udf \leftarrow p(X) \wedge p'(X)$ (\transformation).
-This is a widely used technique in Logic Programming (e.g.
+This is a widely used technique in Logic Programming (cf.
 choice rules in~\cite{GebserKaminskiKaufmannSchaub12}).
-Another principle that we introduce is the \emph{no double negation} principle: Under the Weak Completion Semantics, a positive conclusion can be derived from double negation appearing in two conditionals.
+We also introduce the \emph{no inference through double negation} principle (\dnegation): Under the Weak Completion Semantics, a positive conclusion can be derived from two negations appearing in two conditionals.
 However, it seems that humans normally do not reason in such a way~\cite{khemlani:2012}.
-Hence, we block them with the help of abnormalities (\dnegation).
+Hence, we block them with the help of abnormalities.
 $\CalP_{Eyz}$, the logic program representing \ME yz is as follows:
@@
 and which is semantically equivalent to $\exists X(z(X) \wedge y(X))$ (\converse).
-As shown by Khemlani and Johnson-Laird~\cite{khemlani:2012}, humans seem to
-distinguish between \textit{some y are z} and \textit{all y are z} (\unknownGen).
+Another observation is that humans seem to
+distinguish between \textit{some y are z} and \textit{all y are z} (cf.~\cite{grice1975,khemlani:2012}).
 Accordingly, if we only observe that an object $o$ belongs to $y$ and $z$ then we do not want to conclude both, \textit{some y are z} and \textit{all y are z}.
-In order to distinguish between \textit{some y are z} and \textit{all y are z}, we introduce the following principle:  If we know that \textit{some y are z}, then there must not only be an object~$o_1$ which belongs to $y$ and $z$, but there must be another object $o_2$ which belongs to
+In order to distinguish between \textit{some y are z} and \textit{all y are z}, we introduce the \textit{unknown generalization} principle (\unknownGen):  If we know that \textit{some y are z}, then there must not only be an object~$o_1$ which belongs to $y$ and $z$, but there must be another object $o_2$ which belongs to
 $y$ and which does not belong to~$z$.
-$\CalP_{Iyz}$, the logic program representing \MI yz, is as follows:\footnote{For simplicity, here we omit the clauses introduced by ({\footnotesize\textsf{converse}}).
+$\CalP_{Iyz}$, the logic program representing \MI yz, is as follows:\footnote{We omit the clauses introduced by ({\footnotesize\textsf{converse}}), as they are symmetric.
 They can be found in~\cite{declare:2017}.}
@@
 Consider the statement \textit{Some y are not z} ({\MO yz}), which in CL is represented as $\exists X (y(X) \wedge \neg z(X)$.
-As in the previous cases, \conditionals, \licenses, and \unknownGen are assumed.
+As in the previous cases, \conditionals, \licenses, and \unknownGen\ are assumed.
 As the conclusion is again a negation of a claim, \transformation\ and
@@
 \end{align}
-We have to add the integrity constraint $U \leftarrow z(X) \wedge
+We have to add the integrity constraint $\udf \leftarrow z(X) \wedge
 z'(X)$
@@
 \subsection{Advanced Principles}\label{sect:advanced}
-
-The principles introduced here are assumed not to be applied by all humans, 
-but characterize the reasoning clusters (c.f.\ Section~\ref{sect:aggregateddata}).
+Different than the basic principles, we assume that the principles introduced now, are not assumed to be applied by all humans.
+% % 
+% but characterize the reasoning clusters which will be discussed in Section~\ref{sect:aggregateddata}.
 
@@
 
-In FOL, the conditional statement $\forall (X) (a(X) \rightarrow b(X))$, is logically
+In CL, the conditional statement $\forall (X) (a(X) \rightarrow b(X))$, is logically
 equivalent to its contrapositive, $\forall (X) (\neg b(X) \rightarrow \neg a(X))$.
@@
 If all principles introduced so far are applied to a pair of existentially quantified premises
-(\MI or \MO), then the only object about which an inference can be made is the one
+(\MI\ or \MO), then the only object about which an inference can be made is the one
 resulting from \import\ with respect to the premise it introduced (i.e.\ only inferences about the properties 
@@
 Consider again $\CalP_{I yz}$ of Section~\ref{sect:basic}, where $\Ab_{yz}$ is only false for 
-$o_1$. In case \negFailure\ applies, then $\Ab_{yz}(o_1) \leftarrow \bot$
-in $\CalP_{I yz}$ is replaced by the following two clauses:
+$o_1$. In case \negFailure\ applies, then
+% $\Ab_{yz}(o_1) \leftarrow \bot$ in
+$\CalP_{I yz}$ is extended by the following two clauses:
 \begin{align}
@@
 This information may be explanations about
-facts introduced either from an existential import or from unknown generalization. 
-We use the first as source for observations, as they are used directly to infer new information  (\abduction). Consider the following pair of syllogistic premises: 
+facts introduced from either \import\ or \unknownGen. 
+We use the first as source for observations, as they are used directly to infer new information. Consider the following pair of syllogistic premises: 
 \begin{align}
@@
 This conclusion can be modeled by abductive reasoning, where the system searches for explanations
-for a given observation. The second premise states that some bakers exist, i.e.\ the observation is $\CalO = \{b(o_2) \}$. $\CalO$ can be explained by the first premise, namely that
+for a given observation (\abduction). The second premise states that some bakers exist, i.e.\ the observation is $\CalO = \{b(o_2) \}$. $\CalO$ can be explained by the first premise, namely that
 this baker is an artist. Under WCS, this reasoning procedure was formalized by means of skeptical abduction.
-The interested reader is referred to~\cite{cogsci:2017} for 
-formal abductive approach for human syllogistic reasoning in this paper.
+The interested reader is referred to~\cite{cogsci:2017} for more details.
 
@@
 
-\section{Evaluation} \label{sect:aggregateddata}
-
-\subsection{On aggregated Data}
+\section{Evaluation}
+
+\subsection{On aggregated Data}\label{sect:aggregateddata}
 
@@
     (3) \basic, \converse\ for \MI\ and \ME, and \contraposition\ for~\MA.
-This approach achieved a match of 92\;\% with the aggregated data reported in~\cite{khemlani:2012}, and was the highest result compared to the presented approaches.
+This approach achieved a match of 92\;\% with the aggregated data reported in~\cite{khemlani:2012}, and was the highest result compared to the other approaches.
 
@@
 \begin{center}
-\begin{tabular}{@{\hspace{0cm}}l@{\hspace{0.9cm}}c@{\hspace{0.9cm}}c@{\hspace{0.9cm}}cGG}
-Principle & (1) & (2) & (3) & Heuristic 1 & Heuristic 2 \medskip \\ \midrule  
+\begin{tabular}{@{\hspace{0cm}}l@{\hspace{0.4cm}}c@{\hspace{0.4cm}}c@{\hspace{0.9cm}}cGG}
+Principle & Cluster (1) & Cluster (2) & Cluster (3) & Heuristic 1 & Heuristic 2 \medskip \\ \midrule  
 \conditionals & $\checkmark$   & $\checkmark$ & $\checkmark$ & - & - \smallskip\\
@@
 \end{table}
-The corresponding logic program $\Pie$ that represents the basic cluster, consists of the following clauses:
+The corresponding logic program $\Pie$ for cluster (1) consists of the following clauses:
 \begin{align}
@@
 for which according to the entailment of conclusions as defined in Section~\ref{sect:entail}, \textit{No valid conclusion}
-follows.
+follows.\footnote{For simplicity we ignore the application of {\footnotesize\textsf{searchAlt}}, which is described in more detail in~\cite{cogsci:2017}.}
 
@@
 \textit{No chemists are artists} is entailed by $\CalM_{\Piegen}$, because 
-for all objects which have the property $c$, (which only applies for $o_5$), they do not have the property.
+for all objects which have the property $c$, (which only applies for $o_5$), they do not have the property $a$.
 
@@
 \begin{description}
- \item[Aggregated evaluation] Most of the majority's answers can be explained by the clusters' predictions. In the first week 
+ \item[Aggregated evaluation] Most of the majority's responses can be explained by the clusters' predictions. In the first week 
  the majority's conclusion for 63 out of 64 pair of syllogistic premises and in the second week
- all majority's conclusions were predicted.
+ all majority's conclusions were predicted by the clustering by principles approach, i.e.\
+ by cluster (1), (2) or (3) as specified in Table~\ref{tab:cluster}.
 \item[Non-aggregated evaluation] None of the clusters fitted any individual participants' response pattern, i.e.\ none of the participants belonged to a certain cluster.
@@
 
-Given this sobering outcome with respect to the non-aggregated evaluation, 
-we ignored the clusters, and investigated whether parallels between the applied cognitive principles could be observed: \basic, \contraposition, \abduction\ and \negFailure.
+iven this sobering outcome with respect to the non-aggregated evaluation, 
+we decided to ignore the clusters, and investigated whether parallels between the following cognitive principles could be observed: \basic, \contraposition, \abduction\ and \negFailure.
 Table~\ref{tab:cogprinc} provides an overview of the predicted conclusions according to these principles.\footnote{Note that any advanced principle also includes the logic program clauses specified for \textsf{basic}.}
@@
  by these principles? and
-(iii) Do participants learn, i.e.\ do they improve their logical ability by doing the same reasoning task again?
+(iii) Do participants learn, i.e.\ do they improve their logical ability (with respect to CL) by doing the same reasoning task again?
 We address these questions in the following.
+
+\subsection{On non-aggregated Data} \label{sect:nonag}
+
+A wide amount of cognitive theories for syllogistic reasoning have previously only be evaluated with respect to the aggregated data of psychological experiments.
+Usually, as assessment, typical statistical evaluation criteria with respect to the experimental data 
+ have been used (cf. root-mean-square error, Bayesian information criterion~\cite{schwarz1978}, Akaike information criterion~\cite{akaike1974new}).
+ These measures are helpful as a general guideline, however, their major drawback is that they only compute the quantitative overall performance.
+ In this section, we carry out a qualitative assessment of the suggested cognitive principles with respect to the individual differences
+  among the participants.
+% As already mentioned in the introduction, an issue with these evaluations is that these theories cannot account for individual differences. 
+% Therefore, in this section, we will investigate an approach that accounts for non-aggregated data of participants in syllogistic reasoning.
 
@@
 \clearpage
-\subsection{On non-aggregated Data} \label{sect:nonag}
-
-A wide amount of cognitive theories for syllogistic reasoning have previously only be evaluated with respect to the aggregated data of psychological experiments.
-Usually, as assessment, typical statistical evaluation criteria with respect to the experimental data 
- have been used (cf. root-mean-square error, Bayesian information criterion~\cite{schwarz1978}, Akaike information criterion~\cite{akaike1974new}).
- These measures are helpful as a general guideline, however, their major drawback is that they only compute the quantitative overall performance.
- In this section, we carry out a qualitative assessment of the suggested cognitive principles with respect to the individual differences
-  among the participants.
-% As already mentioned in the introduction, an issue with these evaluations is that these theories cannot account for individual differences. 
-% Therefore, in this section, we will investigate an approach that accounts for non-aggregated data of participants in syllogistic reasoning.
-
 \subsubsection{mReasoner}
 
-Johnson-Laird and Khemlani~\cite{khemlani:2016} proposed an approach to model the individual reasoning patterns of the participants, and 
+Johnson-Laird and Khemlani~\cite{khemlani:2016} proposed an approach to model individual reasoning patterns of participants, and 
 showed that different participants seem to derive different conclusions.
 Their approach is built on top of the mReasoner~\cite{khemlani:2013}, an extension of the mental model theory, first introduced by Johnson-Laird~\cite{johnsonlaird:1983}.
-Reasoning depends on the construction and manipulation of mental models, based on 3 fundamental principles: 
-(i) represent possibilities, (ii) principle of iconicity, and (iii) principle of dual processes.
+The reasoning procedure depends on the construction and manipulation of mental models, based on 3 fundamental principles: 
+(i) mental models represent possibilities, (ii) mental models are iconic (principle of iconicity), and (iii) reasoning procedure consists of two interacting processes, intuition and deliberation (principle of dual processes).
  The procedure of mReasoner consists of two steps:
@@
 predictions based on 2 datasets, week 1 and week 2, for 20 participants.
-The parameters for the construction of the mental models are (i) the 
-size of a mental model, (ii) the probability of drawing conclusions only from the most common mental models, or from the complete set,
-(iii) the probability to search for counter examples,
-and (iv) the probability to weaken the conclusion.
-The clustering with mReasoner was based on a particular clustering algorithm to determine the optimal number of clusters for their dataset.
-The K-means clustering algorithm was used to divide participants into clusters and then
-the mReaoner to simulated the clusters by choosing appropriate parameter settings.
+The parameters for the construction of the mental models are (i) 
+size of a mental model, (ii) probability of drawing conclusions only from the most common mental models, or from complete set of models,
+(iii) probability to search for counter examples,
+and (iv) probability to weaken the conclusion.
+The clustering with mReasoner was based on a particular clustering algorithm to determine the optimal number of clusters for their dataset
+% The K-means clustering algorithm was used to divide participants into clusters
+and then
+the mReaoner simulated the clusters by choosing appropriate parameter settings.
 Johnson-Laird and Khemlani~\cite{khemlani:2016} identified three different reasoning clusters: The intuitive cluster, the intermediate cluster and the deliberative cluster.
-The approach is novel in the sense that until now no other cognitive theory tried to rigorously account for individual differences in human syllogistic reasoning.
-However, their approach seems to have two drawbacks.
-First, some of their
-parameters were of probabilistic nature, i.e.\ there is no explaining factor behind them. Second, the authors did not distinguish between the participants' 
-responses in week 1 and week 2, but instead merged the responses. This might give a distorted picture, as participants could
-have learned from week 1 to week 2.
+The approach is novel as until now no other cognitive theory tried to rigorously account for individual differences in human syllogistic reasoning.
+However, their approach could be optimized in the sense that the authors did not distinguish between the participants'
+responses in week 1 and week 2, but merged their responses. This might give a distorted picture, as participants could have learned from week 1 to week 2.
+We will address  this issue in the following.
 
@@
 
-The non-aggregated data included a dataset of 20 participants,
-for which the task was carried out twice, with a break of one week~\cite{khemlani:2016}. On average, the basic principles explained the results about 10\;\% better than CL.
+Taking the non-aggregated data from~\cite{khemlani:2016} as benchmark,
+we investigated how well their responses could be predicted through the cognitive principles,
+and whether the individual participants' responses would consistently correspond to a certain set of principles.
+Recall, their dataset consisted of responses from 20 participants, with each two entries,
+as the task was carried out twice, with a break of one week.
+
+We observed the following patterns:
+On average, the basic principles explained the results about 10\;\% better than CL.
  Generally, the participants who were strong in CL/ contraposition, drew fewer conclusions according to abduction/ generalization, and vice versa.
- The changes within participant's answers from one to the other week was high, with 52/64 on average. 
- From week 1 to week 2, an increase with respect to CL (19/20, 6\;\%), basic (18/20, 6\;\%), contraposition (15/20, 5\;\%)
- and a decrease on generalization (15/20, 8\;\%) and no change on abduction (10/20, 0\;\%) answers could be observed.
-
-Table~\ref{tab:individual} provides an overview of the performance of of each participant (column 2), for week 1 and week 2 (column 1)
+ The changes within participant's responses from one to the other week was high, with 52/64 on average. 
+ From week 1 to week 2, an increase on conclusions corresponding to CL (19/20, 6\;\%), basic (18/20, 6\;\%), contraposition (15/20, 5\;\%)
+ and a decrease on conclusions corresponding to generalization (15/20, 8\;\%) and no change on
+ conclusions corresponding to abduction (10/20, 0\;\%) could be observed.
+
+Table~\ref{tab:individual} provides an overview of the performance  of each participant (column 2), for week 1 and week 2 (column 1)
 with respect to CL (column 4) and the cognitive principles (column 5, 6, 7, and 8).
 NVC (column 3) denotes how frequently participants responded \textit{no valid conclusion} and Changes (column 9) denotes how
-frequently participants changed their answers from week 1 to week 2. 
+frequently participants changed their responses from week 1 to week 2. 
 The symbols $\uparrow$, $\downarrow$, and $\leftrightarrow$ denote above, below, and average, respectively.
 
-We identified three different clusters: 
+We identified three clusters, using the same labels as in~\cite{khemlani:2016}: 
 The last two columns show the classification of the participants as deliberative, intuitive or intermediate. 
 % The participants in the deliberative cluster here corresponds exactly to~\cite{khemlani:2016}. 
-Our classification only differs wrt.\ Johnson-Laird and Khemlani's 
+Our classification only differs with respect to Johnson-Laird and Khemlani's 
 classification~\cite{khemlani:2016} for the intermediate and intuitive cluster with respect to three participants (1, 11 and 19).
@@
 
-Two of the three clusters can be specified through a set of cognitive principles, and for one, no distinguishable pattern could be identified:
+Two of the three clusters can be specified through a set of cognitive principles, but for one
+group of participants, no distinguishable pattern could be identified:
 % The summary is shown in Table~\ref{tab:clusters}.
 \begin{description}
- \item[Deliberative cluster] (6 out of 20 participants) is characterized by the participants whose answers corresponded to the predictions made by basic/ contraposition.
+ \item[Deliberative cluster] (6 out of 20 participants) is characterized by the participants whose responses corresponded to the predictions made by basic/ contraposition.
 Almost none of their conclusions corresponded to predictions made by abduction/ generalization.
-Theses participants did change their answers from week 1 to week 2
- lower than average and concluded higher than average NVC. Generally, their answers improved from week 1 to week 2 with respect to CL.
+Theses participants did change their responses from week 1 to week 2
+ lower than average and concluded higher than average NVC. Generally, their responses improved from week 1 to week 2 with respect to CL.
  \item[Intuitive cluster] (7 out of 20 participants)
-is characterized by the participants whose answers corresponded to the predictions given by abduction/ generalization.
-Almost none of their conclusions corresponded to the predictions made by contraposition (except of 2 participants in week 2).
-They showed a high change within their own answers from week 1 to week 2
- and concluded lower than average NVC. Generally, their answers improved from week 1 to week 2 with respect to CL, but they never performed \textit{as well as} the participants in 
+is characterized by the participants whose responses corresponded to the predictions given by abduction/ generalization.
+Almost none of their conclusions corresponded to the predictions made by contraposition (except for 2 participants in week 2).
+They showed a high change within their own responses from week 1 to week 2
+ and concluded lower than average NVC. Generally, their responses improved from week 1 to week~2 with respect to CL, but they never performed \textit{as well as} the participants in 
  the deliberative cluster.
- \item[Intermediate cluster] (7 out of 20 participants) is characterized by the participants whose answers did not corresponded to any particular principle.
-These participants had a high change within their own answers from week 1 to week 2
+ \item[Intermediate cluster] (7 out of 20 participants) is characterized by the participants whose responses did not corresponded to any particular set of principles.
+These participants had a high change within their own responses from week 1 to week~2
  and no particular pattern with respect to NVC conclusions.
-Generally they were not improving their answers from week 1 to week 2 with respect to CL, but normally compensated for answers valid in CL in some cases (that corresponded to contraposition) with answers not valid in CL in the other cases (that correspond to abduction/ generalization).
+Generally they were not improving their responses from week 1 to week 2 with respect to CL, but normally \textit{compensated} for responses valid in CL in some cases (that corresponded to contraposition) with responses not valid in CL in the other cases (that correspond to abduction/ generalization).
 \end{description}
@@
 
-We have investigated the cognitive adequacy of clustering by principles under WCS with respect to the individual answer patterns reported by Johnson-Laird and Khemlani~\cite{khemlani:2016}.
+We have investigated the cognitive adequacy of clustering by principles with respect to the individual response patterns reported by Johnson-Laird and Khemlani~\cite{khemlani:2016}.
 Their approach is the only one we are aware of that has evaluated predictions on human syllogistic reasoning
 with respect to the non-aggregated data.
-However in contrast to here, they did not distinguish between the participants' responses in week 1 and week 2.
-
-Our results are two-fold: First, we were able to show that the reasoning clusters as specified in~\cite{declare:2017}, could not be validated
-with respect to these individual reasoning patterns. However, and this leads us to the second more interesting result,
-we could show that participants' answers seemed to be explainable through a set of cognitive principles. In particular,
-some principles seemed to exclude each other by the individual reasoners throughout the experiment (contraposition and abduction), whereas others often appeared together (basic/ contraposition and abduction/ generalization).
-The individual participants' data allowed us to observe other reasoning patterns (high/ low frequency of NVC conclusions and high/ low frequency of changing answers between week 1 and week 2 answers).
-Interestingly, there was one group of participants for which we could not identify any pattern (intermediate)
-and might have changed their strategies within the task.
-Furthermore, it is quite likely that the set of cognitive principles presented in Section~\ref{sect:representing} is not complete.
+Here, we additionally took into account the differences between the participants' responses in week 1 and week 2. 
+Our results are two-fold: First, the reasoning clusters as specified in~\cite{declare:2017}, could not be validated
+with respect to a dataset on individual reasoners. However, we could show that the response patterns of the individual participants corresponded to new sets of cognitive principles. In particular,
+some principles seemed to exclude each other throughout the experiment (contraposition and abduction), whereas others often appeared together (basic with contraposition and abduction with generalization).
+Additionally, the provided data allowed us to observe other patterns (high/ low frequency of NVC and high/ low frequency of changing responses between week 1 and week 2). 
+Interestingly, there was one group of participants for which we could not identify any pattern (intermediate).
 
 % Static clusters for each participant might not be appropriate, rather, a sequential application of cognitive principles seems to be more likely
-For the future, we need to investigate whether the sequential application of the cognitive principles 
+For the future, we intend to investigate whether the sequential application of the cognitive principles 
 can give us more insights on the participants' reasoning pattern. In particular, our hypothesis is that these principles might correspond to some of the reasoning steps within the insight model
@@
 Similarly, a meta-analysis of the individual response patterns for the selection task has been suggested by Ragni, Kola and Johnson-Laird~\cite{ragni:2017}.
-This in turn might give us a new starting point to investigate whether the cognitive principles  specified here can characterize participants' conclusions in other reasoning episodes.
+This in turn might give us a new starting point to investigate whether the cognitive principles  specified here can characterize participants' responses in other reasoning episodes.
+Furthermore, it is quite likely that the set of cognitive principles presented in Section~\ref{sect:representing} is not complete, and thus other principles will be identified in the future.
 

### 2019-11-18 paper/paper-v3.tex
a few typos

@@
 
-iven this sobering outcome with respect to the non-aggregated evaluation, 
+Given this sobering outcome with respect to the non-aggregated evaluation, 
 we decided to ignore the clusters, and investigated whether parallels between the following cognitive principles could be observed: \basic, \contraposition, \abduction\ and \negFailure.

### 2019-11-18 paper/paper-v3.tex
rephrased the introduction

@@
 
+\usepackage[lf]{berenis}
 \renewcommand{\baselinestretch}{1.5}
@@
 Even though the main logical formalism of Classical Logic (CL) has successfully been applied to our scientific reasoning,
-various experimental studies have shown that human reasoning systematically deviates 
-from conclusions that are valid under CL (cf.~\cite{wason:68,byrne:89}).
+it seems that already in 1913, Kotarbi{\'n}ski~\cite{kotarbinski:1913} \textit{suggested
+the need to revise the two-valued logic as it seemed to interfere with the freedom of human thinking}~\cite[p.17]{malinowski:1993}.
 % Originally, one of the objectives in the field of AI and logic programming was to formalize human reasoning and commonsense reasoning~\cite{McCarthy59,McCarthy98}. 
@@
 % this expertise was not consulted, i.e. there was no communication among the relevant fields. 
+Since then, various cognitive theories have been proposed,
+ but it seems that none of them is generally accepted
+within the community. As various experimental studies have shown that human reasoning systematically deviates from conclusions that are valid under CL (cf.~\cite{wason:68,byrne:89}),
+cognitive scientists 
+seem at least to agree with Kotarbi{\'n}ski,
+ that classical (two-valued)
+logic would not be an adequate candidate for a cognitive theory.
+
 Cognitive theories that try to explain human reasoning need to be evaluated by their cognitive adequacy~\cite{strube:1992} and their performance should be compared to other theories.
@@
 	\end{minipage}
-    \begin{minipage}{0.34\textwidth}
+    \begin{minipage}{0.35\textwidth}
         \centering
@@
 \footnotetext{Thanks to the
- anonymous reviewer for the observation, that the $\cxt$ operator seems to be a special case of {\L}ukasiwicz's necessity operator~\cite[p. 20]{malinowski:1993}, restricted to literals.
+ anonymous reviewer for the observation, that the $\cxt$ operator seems to be a special case of {\L}ukasiewicz necessity operator~\cite[p. 20]{malinowski:1993}, restricted to literals.
  }

### 2019-11-18 paper/paper-v3.tex
small change first paragraph

@@
 % this expertise was not consulted, i.e. there was no communication among the relevant fields. 
-Since then, various cognitive theories have been proposed,
- but it seems that none of them is generally accepted
-within the community. As various experimental studies have shown that human reasoning systematically deviates from conclusions that are valid under CL (cf.~\cite{wason:68,byrne:89}),
-cognitive scientists 
-seem at least to agree with Kotarbi{\'n}ski,
- that classical (two-valued)
-logic would not be an adequate candidate for a cognitive theory.
+% but providing a unified, integrated  accepted theory still appears 
+% to be a major challenge. 
+In the last decades, cognitive scientists have repeatedly
+shown through experimental studies that human reasoning systematically deviates from conclusions that are valid under CL (cf.~\cite{wason:68,byrne:89}),
+and thus it seems that they agree with Kotarbi{\'n}ski,
+ that classical (two-valued) logic would not be an adequate candidate for a cognitive theory.
 

### 2019-11-19 paper/paper-v3.tex
optimized the first paragraph in intro

@@
 
-Even though the main logical formalism of Classical Logic (CL) has successfully been applied to our scientific reasoning,
-it seems that already in 1913, Kotarbi{\'n}ski~\cite{kotarbinski:1913} \textit{suggested
-the need to revise the two-valued logic as it seemed to interfere with the freedom of human thinking}~\cite[p.17]{malinowski:1993}.
+Even though the main logical formalism of Classical Logic (CL) has successfully been applied to our scientific reasoning, Kotarbi{\'n}ski~\cite{kotarbinski:1913} already stated in 1913, \textit{the need to revise the two-valued logic as it seemed to interfere with the freedom of human thinking}~\cite[p.17]{malinowski:1993}.
 % Originally, one of the objectives in the field of AI and logic programming was to formalize human reasoning and commonsense reasoning~\cite{McCarthy59,McCarthy98}. 
@@
 shown through experimental studies that human reasoning systematically deviates from conclusions that are valid under CL (cf.~\cite{wason:68,byrne:89}),
-and thus it seems that they agree with Kotarbi{\'n}ski,
- that classical (two-valued) logic would not be an adequate candidate for a cognitive theory.
+and thus they seem to agree with Kotarbi{\'n}ski: Classical (two-valued) logic is not an adequate candidate for a cognitive theory.
+ What should be the properties of such a candidate in the first place?
 

### 2019-11-21 paper/paper-v3.tex
final

@@
 
-Even though the main logical formalism of Classical Logic (CL) has successfully been applied to our scientific reasoning, Kotarbi{\'n}ski~\cite{kotarbinski:1913} already stated in 1913, \textit{the need to revise the two-valued logic as it seemed to interfere with the freedom of human thinking}~\cite[p.17]{malinowski:1993}.
+Even though the main logical formalism of Classical Logic (CL) has successfully been applied to our scientific reasoning, Kotarbi{\'n}ski~\cite{kotarbinski:1913} already stated in 1913 \textit{the need to revise the two-valued logic as it seemed to interfere with the freedom of human thinking}~\cite[p.17]{malinowski:1993}.
 % Originally, one of the objectives in the field of AI and logic programming was to formalize human reasoning and commonsense reasoning~\cite{McCarthy59,McCarthy98}. 
@@
 and thus they seem to agree with Kotarbi{\'n}ski: Classical (two-valued) logic is not an adequate candidate for a cognitive theory.
- What should be the properties of such a candidate in the first place?
+ But what should be the properties of such a candidate in the first place?
 
@@
 for the development of an adequate cognitive theory.
-Only recently approaches have been proposed that account for individual differences (or clusters) in syllogistic reasoning (cf.~\cite{khemlani:2016,declare:2017}).
+Only recently approaches have been proposed that account for individual differences in syllogistic reasoning (cf.~\cite{khemlani:2016,declare:2017}).
 
@@
 Here, we provide a qualitative assessment of the \textit{clustering by principles} approach proposed in~\cite{declare:2017},
-that performed well on aggregated data, and investigate whether these clusters are also adequate with respect to the non-aggregated dataset given by Johnson-Laird and Khemlani~\cite{khemlani:2016}.
+that performed well on aggregated data, and investigate whether these clusters are also adequate with respect to the non-aggregated dataset provided by Johnson-Laird and Khemlani~\cite{khemlani:2016}.
 In Section~\ref{sect:wcsclustering} and~\ref{sect:representing}, we introduce the underlying computational logic approach, the cognitive principles and their formalization.
 Section~\ref{sect:aggregateddata} illustrates an example and briefly discusses the performance of the approach presented in~\cite{declare:2017} on non-aggregated data.
-Section~\ref{sect:nonag} provides a more adequate characterization of the individuals' reasoning patterns.
+Section~\ref{sect:nonag} provides novel characterization of the individuals' reasoning patterns through cognitive principles.
 
@@
 \footnotetext{Thanks to the
- anonymous reviewer for the observation, that the $\cxt$ operator seems to be a special case of {\L}ukasiewicz necessity operator~\cite[p. 20]{malinowski:1993}, restricted to literals.
+ anonymous reviewer for the observation, that the $\cxt$ operator seems to be a special case of {\L}ukasiewicz necessity operator (cf.~\cite[p. 20]{malinowski:1993}), restricted to literals.
  }
@@
 
-\subsection{On aggregated Data}\label{sect:aggregateddata}
+\subsection{Clusters}\label{sect:aggregateddata}
 
@@
 
-\subsection{On non-aggregated Data} \label{sect:nonag}
+\subsection{Individual Reasoners} \label{sect:nonag}
 
@@
 
-\subsubsection{Individual Reasoning Patterns}
+\subsubsection{Characterization through Principles}
 
@@
 
-Taking the non-aggregated data from~\cite{khemlani:2016} as benchmark,
-we investigated how well their responses could be predicted through the cognitive principles,
-and whether the individual participants' responses would consistently correspond to a certain set of principles.
-Recall, their dataset consisted of responses from 20 participants, with each two entries,
+Taking the non-aggregated data from~\cite{khemlani:2016},
+we investigated how well the participants' responses could be explained through the cognitive principles,
+and whether their individual responses would be consistent with respect to a certain set of principles.
+Recall, Johnson-Laird and Khemlani's dataset consisted of responses from 20 participants, with each two entries,
 as the task was carried out twice, with a break of one week.
@@
 
-We identified three clusters, using the same labels as in~\cite{khemlani:2016}: 
+We identified three types of reasoners, characterizing them according to the terms already used in~\cite{khemlani:2016}: 
 The last two columns show the classification of the participants as deliberative, intuitive or intermediate. 
@@
 Our classification only differs with respect to Johnson-Laird and Khemlani's 
-classification~\cite{khemlani:2016} for the intermediate and intuitive cluster with respect to three participants (1, 11 and 19).
+classification~\cite{khemlani:2016} for the intermediate and intuitive reasoners with respect to three participants (1, 11 and 19).
 A possible explanation for this difference could be that they
 did not take into account the participants' changes between week 1 and week 2.
-
-Two of the three clusters can be specified through a set of cognitive principles, but for one
+Two of the three types of reasoners can be characterized through a set of cognitive principles, but for one
 group of participants, no distinguishable pattern could be identified:
@@
 \begin{description}
- \item[Deliberative cluster] (6 out of 20 participants) is characterized by the participants whose responses corresponded to the predictions made by basic/ contraposition.
+ \item[Deliberative reasoners] (6 out of 20 participants) is characterized by the participants whose responses corresponded to the predictions made by basic/ contraposition.
 Almost none of their conclusions corresponded to predictions made by abduction/ generalization.
@@
  lower than average and concluded higher than average NVC. Generally, their responses improved from week 1 to week 2 with respect to CL.
- \item[Intuitive cluster] (7 out of 20 participants)
+ \item[Intuitive reasoners] (7 out of 20 participants)
 is characterized by the participants whose responses corresponded to the predictions given by abduction/ generalization.
@@
  the deliberative cluster.
- \item[Intermediate cluster] (7 out of 20 participants) is characterized by the participants whose responses did not corresponded to any particular set of principles.
+ \item[Intermediate reasoners] (7 out of 20 participants) is characterized by the participants whose responses did not corresponded to any particular set of principles.
 These participants had a high change within their own responses from week 1 to week~2
